### 2016-11-29 paper_FMA_nov2016.tex
xav_v1

\documentclass[a4paper,12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{array}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage[top=1in,bottom=2in,left=1in,right=1in]{geometry}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{color}
\usepackage[normalem]{ulem}

\usepackage{booktabs} 
\usepackage{rotating,tabularx}

%\newcommand{\Mn}[1]{ \mathrm{M}_{#1}(\mathbb{R}) }
%\newcommand{\lex}{ \mathrm{lex} }
%\newcommand{\vect}{ \mathrm{vec} }
%\newcommand{\R}{\mathbb{R}}
%\newcommand{\x}{\mathbf{x}}
%\newcommand{\vv}{\mathbf{v}}
%\newcommand{\uu}{\mathbf{u}}
%\newcommand{\e}{\mathbf{e}}
%\newcommand{\rd}{\mathrm{d}}
%\newcommand{\re}{\mathrm{e}}
%\newcommand{\E}{\mathbb{E}}
%\newcommand{\Prob}{\mathbb{P}}
%\newcommand{\op}{\mathrm{op}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{example}[theorem]{Example}
%\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{example}[theorem]{Example}

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{tr}}

%\title{Free Music Archive (FMA): \\ A Dataset For Music Analysis}
\title{FMA: A Dataset For Music Analysis}

\author{Kirell Benzi, Micha\"el Defferrard, Pierre Vandergheynst, Xavier Bresson\\ \vspace{-0.5cm} \\  LTS2, EPFL, Lausanne, Switzerland}

\date{\today}

\begin{document}

%%%%%%%%%%%%%%%%%%%
\maketitle
%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%
\abstract{We present a new music dataset that can be used for several music analysis tasks. Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, availability and legality of the music data, or the lack of meta-data for artists analysis or song ratings for recommender systems. Existing datasets such as GTZAN, TagATune, and Million Song Dataset suffer from the previous limitations. It is however essential to establish such benchmark datasets to advance the field of music analysis, like the ImageNet dataset that made possible the large success of deep learning techniques in computer vision. In this paper, we introduce the Free Music Archive (FMA) dataset which size is 86,992 songs with raw audio tracks and meta-data including artist name, song title, music genre, and track counts. For research purposes, we define 2 additional datasets from the original one; a small genre-balanced dataset of 1,000 data and a medium genre-unbalanced dataset of 14,570 data, both with audio tracks and Echonest features. For all datasets, we provide a train-test splitting for future algorithms' comparisons. }
%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%
\section{Introduction and Motivation}
%%%%%%%%%%%%%%%%%%%

\noindent
{\bf Music Analysis Field.} Developing new mathematical models and algorithms to solve challenging real-world problems is obviously of first importance in any field of research. But such novel techniques must be evaluated and compared to the existing state-of-the-art techniques to be adopted as new standards by research communities. Evaluation and comparison require benchmark datasets to achieve their purpose, which are sometimes challenging to collect and share. In computer vision, the community has developed over the years well-established benchmark datasets such as MNIST \cite{pro:LeCunBottouBengioHaffner98MNIST} publicly accessible at \cite{art:MNIST}, CIFAR \cite{art:Krizhevsky09CIFAR} available at \cite{art:CIFAR}, or ImageNet \cite{art:FeiFei09ImageNet} provided at \cite{art:ImageNet}. All these image datasets are free, legal and easily available on Internet. Such datasets have proved essential to advance the field of computer vision. The most celebrated example is the ImageNet dataset and challenge in 2012. This unprecedented dataset of 1.5M image data allowed to demonstrate the power of deep learning techniques, which was able to win the competition by a large margin over the second best. In music analysis, such benchmark datasets are not easily available and they also lack of essential information. For examples, developing deep learning techniques for music applications require the use of raw audio data, and designing recommender systems need meta-data like artist features and song ratings. Besides, the most influential competition in the field of music analysis organized every year is the Music Information Retrieval Evaluation eXchange (MIREX) \cite{art:MIREX}. MIREX proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc. However, participants do not have access to the dataset; neither the test set, nor the important train set. They must upload their code that will be evaluated by the organizers. In other words, participants cannot train their analysis models on any part of the dataset. These existing shortcomings make essential the development of new music datasets to advance the field of music analysis algorithms.   \\

\noindent
{\bf Available music datasets and their limitations.} In music analysis, three datasets have been widely used by the international society for music information retrieval (ISMIR) \cite{art:ISMIR}: the GTZAN, TagATune, and MSD (Million Song Dataset) datasets. However each dataset come with some limitations: \\
$\bullet$ GTZAN \cite{art:TzanetakisCook02GTZAN} is a collection of 1000 songs with 10 music genres. Each song is represented by a 22050Hz Mono 16-bit wav audio file of 30sec. It is the most popular music dataset with 6,600 answers on Google search, and 600 on Google scholar on Nov 2016. It has been widely used for different music applications. A complete study of this dataset has been done in \cite{art:Sturm12GTZAN}, and the dataset is available at \cite{art:GTZAN}. The main limitations of GTZAN is the legality of the dataset, the small size with 1000 data, no complete meta-data regarding artist names and song titles, and no additional meta-data like ratings. \\
$\bullet$ TagATune \cite{art:Law07TagATune} was a popular dataset of 5,405 source songs of 230 artists. It was used for music tagging and music metric learning. Its website is \cite{art:TagATune}, but it is not anymore publicly available since 2009.\\
$\bullet$ The Million Song Dataset (MSD) \cite{art:Bertin11MSD} is a free and legal collection of 1 million music songs. Each song data is composed of high-level and medium-level audio features provided by  Echonest service and meta-data like artist name, song title, track genre, etc. It is available at \cite{art:MSD}. The very large scale of this music dataset would make it an ideal candidate for deep learning, which works best with large datasets. However, deep learning techniques work directly on the raw data (audio tracks in this case, like raw image pixels in computer vision) and not on pre-processed audio features given by Echonest. Unfortunately, this prevents the application of deep learning on this dataset. Finally, Echonest reassigned the indexes of the MDS songs, which today makes (almost) impossible the use of the most music recent features developed by Echonest. \\

\noindent
{\bf What would be the requirements for an ideal music dataset?}  \\
(1) Availability of raw audio data like 30sec or more\\
(2) Availability of meta-data such as genres, artist, title, year, lyrics, users' ratings, users' comments, etc\\
(3) Large-scale dataset to be able to sample the high-dimensional distribution of music genres\\
(4) Free and easy access of the dataset\\
(5) Legality of the dataset, no copyright issue with e.g. open access licence\\

%%%%%%%%%%%%%%%%%%%
\section{Introducing the FMA dataset}
%%%%%%%%%%%%%%%%%%%

\noindent
{\bf The website.} The FMA website \cite{art:FMA} is run by WFMU, the longest-running free-form radio station in the United States. The website provides a large catalogue of artists and high-quality mp3 songs. Each song is legally free to download as artists decided to work under open licence such as Creative Commons license. This is directly inspired by the open source movement, and the goal of the FMA site is to offer a legal and technological platform for artists and listeners to harness the full potential of online music sharing \cite{art:MossFMA}. FMA has been funded by the New York State Music Fund, the MacArthur Foundation, and the National Endowment for the Arts.\\

\noindent
{\bf Crawling the data.} The FMA organization provides an API \cite{art:FMA_API} to collect the meta-data associated to each song such as artist name, track title, track genre, and track listens. The collect of the mp3 audio files was done with a web scrapper. The collection of FMA was done in April 2016, and the size of the dataset at this time was 86,992 songs. It was recently announced on July 2016 that the website had reached the landmark of 100,000 songs. \\

\noindent
{\bf The raw FMA dataset.} The fully collected dataset on April 2016 contained 86,992 music data. Each music data is composed of an mp3 audio file and some meta-data provided on the FMA website for each song, and crawled by the API. All mp3 audio tracks are coded with sampling rate of 44,100Hz, bit rate 128kb/s, and in stereo. However, not all meta-data is available for each song, and we thus decided to remove the data which has not the following complete meta-information; artist name, track title, track genres, and track listens. This reduced the dataset to 82,113 data. Then, we created a top genre by simply picking the first genre in the original list of track genres. A few examples of data are given in Table \ref{tab1}. Figure \ref{fig2} also presents a part of the distribution of genre sizes of the 82,113 data, which obeys to a power law distribution as often found in natural data collection. We further cleaned the dataset by removing songs which belong to "non-standard" music genres like 'Noise', 'Garage', 'Sound Collage', 'Singer', 'Audio Collage', 'Glitch', 'Unclassifiable', 'Lo-Fi', 'Spoken', 'Poetry', 'Talk Radio', 'Avant-Garde', 'Experimental', 'Ambient', and 'Field Recording'. We also suppressed the songs with the title 'Untitled'. After this filtering, the remaining size is 37,741. Then, we decided to cut the long tail of the power law distribution by removing songs of genres with less than 300 songs, which gave us 30,109 data. Finally, we kept the songs that have the raw audio and all Echonest features. Echonest features are medium-level and high-level audio features provided by Echonest \cite{art:Echonest}, a music platform for developers and media companies which has a collection of 3.5M artists and 37.5M songs. Echonest features can be either computed by uploading the songs to the Echonest website, which are then analyzed and send them back to the user, or simply retrieved from the Echonest database if the songs were analyzed before. All Echonest operations are done through an API \cite{art:Echonest_API}. Our final filtering consists in keeping songs that have both raw audio file and complete Echonest features. The final size of the dataset is 14,570. Figure \ref{fig3} presents a part of the distribution of the genre sizes for this final dataset. \\

\noindent
{\bf Proposed benchmark datasets.} For research purposes, we propose 3 datasets:\\
(1) FMA\_small: This dataset contains 1000 songs with 10 genres equally balanced, that is 100 songs per class. For each song, we extracted 30sec precisely in the middle of the song. The idea is to create an alternative to the popular GTZAN dataset with respect to the size parameter. But unlike GTZAN, all 1000 songs have not only raw audio tracks but also meta-data composed of artist name, track title, track genres, track listens, and Echonest features. The GTZAN data may not have for example song title or artist name, and also no copyright permission. For reproducibility research, we also provide a split 75-25 of the dataset for future techniques' comparisons. Eventually, the genre distribution of this dataset is given in Figure \ref{fig3}.  \\
(2) FMA\_medium: This dataset has 14,570 songs of 30sec with 20 genres. Unlike FMA\_small, the size distribution of genres is not balanced, but all songs have raw audio files and all meta-data. We also provide a split 75-25 of this dataset. The genre distribution is plotted in Figure \ref{fig2}. \\
(3) FMA\_large: This is the original full dataset collected on April 2016. It has 86,992 data. No post-processing was applied. This is as real-world data as it can get. The genre distribution is given by Figure \ref{fig1}. \\

%%%%%%%%%%%%%%%%%%%%
%\section{Classification Experiment [MICHAEL SECTION LATER LATER]}
%%%%%%%%%%%%%%%%%%%%
%
%\noindent
%{\bf Baseline classification techniques.}  In this section, we consider perhaps the most popular music analysis problem, the music genre recognition (MGR) problem. We apply standard baseline classification techniques on 2 representations of audio data: \\
%(1) Echonest medium-level 256-dim features, \\
%(2) Spectrograms of audio data by concatenating 10 spectrograms of 3sec into a single long vector.\\
%Table xxx reports the classification accuracy on the train and test datasets.\\
%
%
%\noindent
%{\bf Deep learning.} One major motivation to construct the FMA dataset was to apply the powerful deep learning techniques to the music analysis problem. As a preliminary result, we simply apply a convolutional neural network on the 3sec-spectrograms, and we select the class by majority voting. \\
%Table xxx reports the classification accuracy on the train and test datasets.\\

%%%%%%%%%%%%%%%%%%%
\section{Conclusion and Perspectives}
%%%%%%%%%%%%%%%%%%%

\noindent
{\bf Availability and reproducible research.} For the sake of open science, the proposed 3 benchmarked datasets are made available for free at \href{https://lts2.epfl.ch/datasets}{https://lts2.epfl.ch/datasets}. In addition to the existing ones, we hope the music analysis community will benefit from this new dataset.   \\

\noindent
{\bf Potential applications.} We foresee multiple applications of this new dataset. The most obvious one is for music analysis with deep learning techniques. As raw music tracks are available, deep learning techniques like convolutional neural networks \cite{pro:LeCunBottouBengioHaffner98MNIST} and recurrent neural networks \cite{art:HochreiterSchmidhuber97LSTM} can be directly applied. Besides, standard data analysis tasks such as classification, clustering, regression, text analysis, visualization can also be applied to this new music dataset.\\

%%%%%%%%%%%%%%%%%%%
\vspace{1cm}
\bibliography{bib_paper}
\bibliographystyle{plain}
%%%%%%%%%%%%%%%%%%%

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width=13cm]{histo1.pdf}
\caption{Histograms of the top 15 music genres for the large FMA dataset of 82,113 songs.}
\label{fig1}
\end{figure}

%\begin{table}[h!]
%\centering
%\begin{tabular}{lr}
%\toprule
%{} &  count \\
%\midrule
%top\_genre     &        \\
%Electronic    &  15098 \\
%Avant-Garde   &   7715 \\
%Rock          &   5636 \\
%Pop           &   5211 \\
%Experimental  &   4769 \\
%Folk          &   3339 \\
%Hip-Hop       &   3273 \\
%Punk          &   2948 \\
%Lo-Fi         &   2619 \\
%Noise         &   2339 \\
%Soundtrack    &   1937 \\
%Jazz          &   1573 \\
%Classical     &   1487 \\
%Blues         &   1453 \\
%International &   1397 \\
%\bottomrule
%\end{tabular}
%\caption{xxx}
%\label{table2}
%\end{table}

\begin{figure}[h!]
\centering
\includegraphics[width=13cm]{histo2.pdf}
\caption{Histograms of the 20 music genres for the medium FMA dataset of 14,570 songs.}
\label{fig2}
\end{figure}

%\begin{table}[h!]
%\centering
%\begin{tabular}{lr}
%\toprule
%{} &  count \\
%\midrule
%top\_genre           &        \\
%Electronic          &   2962 \\
%Rock                &   2385 \\
%Pop                 &   1636 \\
%Hip-Hop             &   1059 \\
%Folk                &   1025 \\
%Punk                &    882 \\
%Indie-Rock          &    601 \\
%Jazz                &    588 \\
%Old-Time / Historic &    460 \\
%Psych-Rock          &    456 \\
%International       &    431 \\
%Classical           &    408 \\
%Chiptune            &    254 \\
%Blues               &    245 \\
%Psych-Folk          &    245 \\
%Post-Punk           &    222 \\
%Metal               &    191 \\
%Soundtrack          &    172 \\
%Trip-Hop            &    171 \\
%Post-Rock           &    118 \\
%\bottomrule
%\end{tabular}
%\caption{xxx}
%\label{table3}
%\end{table}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{histo3.pdf}
\caption{Histograms of the 10 music genres for the small FMA dataset of 1000 songs.}
\label{fig3}
\end{figure}

\newpage

%\begin{table}[h!]
\begin{sidewaystable}
\tiny{
\centering
\begin{tabular}{lrlrllll}
\toprule
 fma\_id &                                             artist &  play\_count &                                              title &                                             genres &            top\_genre  \\
    10 &                                          Kurt Vile &       42936 &                                            Freeway &                                              [Pop] &                  Pop  \\
    134 &                                               AWOL &         880 &                                       Street Music &                                          [Hip-Hop] &              Hip-Hop  \\
     141 &                    Alec K. Redfearn \& the Eyesores &         590 &                                               Ohio &                                             [Folk] &                 Folk  \\
     142 &                    Alec K. Redfearn \& the Eyesores &         670 &                               Punjabi Watery Grave &                                             [Folk] &                 Folk  \\
    144 &                                   Amoebic Ensemble &         901 &                                            Wire Up &                                             [Jazz] &                 Jazz  \\
   145 &                                   Amoebic Ensemble &         682 &                                         Amoebiasis &                                             [Jazz] &                 Jazz  \\
    188 &                                           Ed Askew &         973 &                                            Piano 1 &                                             [Folk] &                 Folk  \\
     206 &                                           Ed Askew &          73 &                                        Bella Crane &                                             [Folk] &                 Folk \\
     236 &                                       Banana Clipz &        6695 &                              Push Am (Left, Right) &                              [Electronic, African] &           Electronic  \\
     237 &                                          Barnacled &        1056 &                    Garbage and Fire &                                             [Jazz] &                 Jazz  \\
     238 &                                          Barnacled &         961 &                                     France Attacks &                                             [Jazz] &                 Jazz  \\
     461 &                              Cantonement Jazz Band &        3933 &                                           Bessemer &                               [Blues, Jazz: Vocal] &                Blues  \\
     462 &                              Cantonement Jazz Band &        3256 &                                     Has Been Blues &                               [Blues, Jazz: Vocal] &                Blues  \\
     463 &                              Cantonement Jazz Band &        3238 &                                       I'll Be Blue &                               [Blues, Jazz: Vocal] &                Blues  \\
    824 &                     Here Comes A Big Black Cloud!! &         391 &                                         Black Mold &          [Rock, Loud-Rock, Psych-Rock, Indie-Rock] &                 Rock  \\
     825 &                     Here Comes A Big Black Cloud!! &        1847 &                                        Death March &          [Rock, Loud-Rock, Psych-Rock, Indie-Rock] &                 Rock  \\
    837 &                                          Heroin UK &         146 &                                           DopeSick &                                             [Rock] &                 Rock \\
     889 &                                 Illusion of Safety &         163 &                                          Wasteland &                                       [Electronic] &           Electronic  \\
     896 &                                        Impediments &        1542 &                                               2012 &                                  [Punk, Power-Pop] &                 Punk  \\
          992 &                                      Jason Willett &         560 &                   Beautiful Song w/ kick drum solo &                                             [Rock] &                 Rock  \\
\bottomrule
\end{tabular}
\caption{A few song data in the FMA dataset.}
\label{tab1}
%\end{table}
}
\end{sidewaystable}

\end{document}

### 2016-12-06 paper_FMA_nov2016.tex
figures in own folder
xav_v2

@@
 \usepackage[top=1in,bottom=2in,left=1in,right=1in]{geometry}
-\usepackage{graphicx}
+\usepackage{graphicx} \graphicspath{{figures/}}
 \usepackage{subfigure}
@@
 \abstract{We present a new music dataset that can be used for several music analysis tasks.
-Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, availability and legality of the music data, or the lack of meta-data for artists analysis or song ratings for recommender systems.
-Existing datasets such as GTZAN, TagATune, and Million Song Dataset suffer from the previous limitations.
-It is however essential to establish such benchmark datasets to advance the field of music analysis, like the ImageNet dataset that made possible the large success of deep learning techniques in computer vision.
-In this paper, we introduce the Free Music Archive (FMA) dataset which size is 86,992 songs with raw audio tracks and meta-data including artist name, song title, music genre, and track counts.
-For research purposes, we define 2 additional datasets from the original one; a small genre-balanced dataset of 1,000 data and a medium genre-unbalanced dataset of 14,570 data, both with audio tracks and Echonest features.
+Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of meta-data for artists analysis or song ratings for recommender systems.
+Existing datasets such as GTZAN, TagATune, and Million Song suffer from the previous limitations.
+It is however essential to establish such benchmark datasets to advance the field of music analysis, like the ImageNet dataset which made possible the large success of deep learning techniques in computer vision.
+In this paper, we introduce the Free Music Archive (FMA) which contains 77,643 songs and 68 genres spanning 26.9 days of song listening and meta-data including artist name, song title, music genre, and track counts.
+For research purposes, we define two additional datasets from the original one: a small balanced-genre dataset of 4,000 song data and 10 genres compassing 33.3 hours of raw audio and a medium unbalanced-genre dataset of 14,511 data and 20 genres offering 5.1 days of track listening, both datasets come with meta-data and Echonest audio features.
 For all datasets, we provide a train-test splitting for future algorithms' comparisons.
-}
+\\\\
+{\it Keywords.} Music dataset, Free Music Archive (FMA), open science, reproducibility research, music analysis.} 
 %%%%%%%%%%%%%%%%%%%
@@
 MIREX proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc.
-However, participants do not have access to the dataset; neither the test set, nor the important train set.
-They must upload their code that will be evaluated by the organizers.
+However, participants do not have access to neither the test set, nor the important train set.
+They must upload their code to a website that will be evaluated by the organizers.
 In other words, participants cannot train their analysis models on any part of the dataset.
-These existing shortcomings make essential the development of new music datasets to advance the field of music analysis algorithms.
+These existing shortcomings make essential the development of new music datasets to advance the field of music analysis.
 \\
@@
 $\bullet$ GTZAN \cite{art:TzanetakisCook02GTZAN} is a collection of 1000 songs with 10 music genres.
-Each song is represented by a 22050Hz Mono 16-bit wav audio file of 30sec.
+Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
 It is the most popular music dataset with 6,600 answers on Google search, and 600 on Google scholar on Nov 2016.
@@
 A complete study of this dataset has been done in \cite{art:Sturm12GTZAN}, and the dataset is available at \cite{art:GTZAN}.
-The main limitations of GTZAN is the legality of the dataset, the small size with 1000 data, no complete meta-data regarding artist names and song titles, and no additional meta-data like ratings.
+The main limitations of GTZAN is the legality of the dataset, the small size, no complete meta-data regarding artist names and song titles, and no additional meta-data like ratings.
 \\
@@
 However, deep learning techniques work directly on the raw data (audio tracks in this case, like raw image pixels in computer vision) and not on pre-processed audio features given by Echonest.
-Unfortunately, this prevents the application of deep learning on this dataset.
-Finally, Echonest reassigned the indexes of the MDS songs, which today makes (almost) impossible the use of the most music recent features developed by Echonest.
+This prevents the application of deep learning on this dataset.
+Finally, Echonest reassigned the indexing of the MDS songs with its database, which from now makes (almost) impossible the use of the most music recent features developed by Echonest.
 \\
 
+\newpage
 \noindent
@@
 The collect of the mp3 audio files was done with a web scrapper.
-The collection of FMA was done in April 2016, and the size of the dataset at this time was 86,992 songs.
+The collection of FMA was done in April 2016, and the size of the dataset at this time was 89,912 songs.
 It was recently announced on July 2016 that the website had reached the landmark of 100,000 songs.
@@
 \noindent
-{\bf The raw FMA dataset.} The fully collected dataset on April 2016 contained 86,992 music data.
-Each music data is composed of an mp3 audio file and some meta-data provided on the FMA website for each song, and crawled by the API.
-All mp3 audio tracks are coded with sampling rate of 44,100Hz, bit rate 128kb/s, and in stereo.
-However, not all meta-data is available for each song, and we thus decided to remove the data which has not the following complete meta-information; artist name, track title, track genres, and track listens.
-This reduced the dataset to 82,113 data.
+{\bf The raw FMA dataset.} The dataset collected on April 2016 contains 89,912 music data.
+As a first filtering, we kept the data that has audio tracks, reducing the size to 86,886 music data.
+All audio tracks are mp3-encoded with sampling rate of 44,100Hz, bit rate 128kb/s, and in stereo.
+Then, we considered the meta-data provided by the FMA website and crawled with its API.
+The meta-information collected is artist name, track title, track genres, and track listens.
+We removed data without genre, giving 79,947 song data.
 Then, we created a top genre by simply picking the first genre in the original list of track genres.
 A few examples of data are given in Table \ref{tab1}.
-Figure \ref{fig2} also presents a part of the distribution of genre sizes of the 82,113 data, which obeys to a power law distribution as often found in natural data collection.
-We further cleaned the dataset by removing songs which belong to "non-standard" music genres like 'Noise', 'Garage', 'Sound Collage', 'Singer', 'Audio Collage', 'Glitch', 'Unclassifiable', 'Lo-Fi', 'Spoken', 'Poetry', 'Talk Radio', 'Avant-Garde', 'Experimental', 'Ambient', and 'Field Recording'.
+At this stage the number of genres is 138, but half have only a few samples.
+Thus we decided to retain the genres with more than 100 samples.
+This defines a dataset of 77,643 music data with 68 genres.
+Figure \ref{fig_large} presents the distribution of the top 20 genres, which obeys to a power law distribution as often encountered in natural data collection.
+We further cleaned the dataset by removing tracks which belong to ``non-standard'' music genres like 'Noise', 'Garage', 'Sound Collage', 'Singer', 'Audio Collage', 'Glitch', 'Unclassifiable', 'Lo-Fi', 'Spoken', 'Poetry', 'Talk Radio', 'Avant-Garde', 'Experimental', 'Ambient', and 'Field Recording'.
 We also suppressed the songs with the title 'Untitled'.
-After this filtering, the remaining size is 37,741.
-Then, we decided to cut the long tail of the power law distribution by removing songs of genres with less than 300 songs, which gave us 30,109 data.
+Besides, we decided to cut the long tail of the power law distribution by removing songs of genres with less than 300 songs.
 Finally, we kept the songs that have the raw audio and all Echonest features.
@@
 All Echonest operations are done through an API \cite{art:Echonest_API}.
-Our final filtering consists in keeping songs that have both raw audio file and complete Echonest features.
-The final size of the dataset is 14,570.
-Figure \ref{fig3} presents a part of the distribution of the genre sizes for this final dataset.
+Our final filtering consists in keeping songs that have both raw audio file and complete Echonest features already computed and stored in the Echonest database.
+The final size of the dataset is 14,511 music data and 20 genres.
+Figure \ref{fig_medium} presents the distribution of the genre sizes for this dataset.
 \\
@@
 {\bf Proposed benchmark datasets.} For research purposes, we propose 3 datasets:\\
-(1) FMA\_small: This dataset contains 1000 songs with 10 genres equally balanced, that is 100 songs per class.
+(1) FMA\_small: This dataset contains 4,000 songs with 10 genres equally balanced, that is 400 songs per class.
 For each song, we extracted 30sec precisely in the middle of the song.
-The idea is to create an alternative to the popular GTZAN dataset with respect to the size parameter.
-But unlike GTZAN, all 1000 songs have not only raw audio tracks but also meta-data composed of artist name, track title, track genres, track listens, and Echonest features.
+The total length is 33.3 hours of audio listening and the size is 3.4GB.
+The genre distribution is given in Figure \ref{fig_small}.
+The motivation for this dataset is to create an alternative to the popular GTZAN dataset with respect to the size and genre parameters.
+But unlike GTZAN, all songs have not only raw audio tracks but also meta-data composed of artist name, track title, track genres, track listens, and Echonest features.
 The GTZAN data may not have for example song title or artist name, and also no copyright permission.
-For reproducibility research, we also provide a split 75-25 of the dataset for future techniques' comparisons.
-Eventually, the genre distribution of this dataset is given in Figure \ref{fig3}.
+For reproducibility research, we also provide a split 80-20 of this dataset for future techniques' comparisons.
 \\
-(2) FMA\_medium: This dataset has 14,570 songs of 30sec with 20 genres.
-Unlike FMA\_small, the size distribution of genres is not balanced, but all songs have raw audio files and all meta-data.
-We also provide a split 75-25 of this dataset.
-The genre distribution is plotted in Figure \ref{fig2}.
+(2) FMA\_medium: This dataset has 20 genres with 14,511 songs of 30sec, also extracted from the middle of the track.
+The total length is 5.1 days of listening and the size is 12.2GB.
+Unlike FMA\_small, the size distribution of genres is not balanced, see Figure \ref{fig_medium}, but all songs have raw audio files and all meta-data i.e.
+artist name, track title, track genres, track listens, and Echonest features.
+We also provide a split 80-20 of this dataset.
 \\
-(3) FMA\_large: This is the original full dataset collected on April 2016.
-It has 86,992 data.
-No post-processing was applied.
-This is as real-world data as it can get.
-The genre distribution is given by Figure \ref{fig1}.
-\\
+(3) FMA\_large: This dataset has 77,643 data of 30sec, again taken from the track middle, and 68 genres.
+It provides 26.9 days of track listening for 79.8GB.
+Figure \ref{fig_large} presents the top classes in the genre distribution.
+All songs come with the following meta-data: artist name, track title, track genres, track listens.
+Eventually, we give a train-test split 80-20 of this dataset.\\
+
+\noindent
+Regarding the splitting train-test sets, we considered these constraints:\\
+(a) For each genre, we randomly split train and test songs with the same ratio 80-20.\\
+(b) We forced an artist to be either in the train set, or the test set (meaning the artist's songs cannot be in both train and test sets). \\
+(c) We filled alternatively the train set and test set in the descending order of the count of artist songs. This prevented the test set to be filled with artists who have only a few songs.\\
+
+\noindent
+The FMA\_large dataset of 77,643 data contains songs of 30sec length, and artist name, track title, track genres, track listens as meta-data.
+Our original motivation was to give the full audio tracks, not limited to 30sec, but the size of the whole music tracks is 806.8GB (for approximatively 256 days of music listening) makes the hosting a challenging issue.
 
@@
 \noindent
-{\bf Availability and reproducible research.} For the sake of open science, the proposed 3 benchmarked datasets are made available for free at \href{https://lts2.epfl.ch/datasets}{https://lts2.epfl.ch/datasets}.
-In addition to the existing ones, we hope the music analysis community will benefit from this new dataset.
-\\
+{\bf Availability and reproducible research.} The 3 proposed FMA datasets are available at\\
+
+ \url{https://lts2.epfl.ch/datasets/fma}\\
+
+ \noindent
+ We also provide IPython notebooks at this address\\
+
+ \url{https://github.com/mdeff/fma}\\
+
+ \noindent
+ to load the datasets and to start using them for research purposes.\\
 
@@
 \centering
-\includegraphics[width=13cm]{histo1.pdf}
-\caption{Histograms of the top 15 music genres for the large FMA dataset of 82,113 songs.}
-\label{fig1}
+\includegraphics[height=6cm]{histo_large.pdf}
+\vspace{-0.5cm}
+\caption{Top 20 music genres for the FMA\_large dataset of 77,643 songs.}
+\label{fig_large}
 \end{figure}
-
-%\begin{table}[h!]
-%\centering
-%\begin{tabular}{lr}
-%\toprule
-%{} &  count \\
-%\midrule
-%top\_genre     &        \\
-%Electronic    &  15098 \\
-%Avant-Garde   &   7715 \\
-%Rock          &   5636 \\
-%Pop           &   5211 \\
-%Experimental  &   4769 \\
-%Folk          &   3339 \\
-%Hip-Hop       &   3273 \\
-%Punk          &   2948 \\
-%Lo-Fi         &   2619 \\
-%Noise         &   2339 \\
-%Soundtrack    &   1937 \\
-%Jazz          &   1573 \\
-%Classical     &   1487 \\
-%Blues         &   1453 \\
-%International &   1397 \\
-%\bottomrule
-%\end{tabular}
-%\caption{xxx}
-%\label{table2}
-%\end{table}
 
@@
 \centering
-\includegraphics[width=13cm]{histo2.pdf}
-\caption{Histograms of the 20 music genres for the medium FMA dataset of 14,570 songs.}
-\label{fig2}
+\includegraphics[height=6cm]{histo_medium.pdf}
+\vspace{-0.5cm}
+\caption{The 20 music genres for the FMA\_medium dataset with 14,511 songs.}
+\label{fig_medium}
 \end{figure}
-
-%\begin{table}[h!]
-%\centering
-%\begin{tabular}{lr}
-%\toprule
-%{} &  count \\
-%\midrule
-%top\_genre           &        \\
-%Electronic          &   2962 \\
-%Rock                &   2385 \\
-%Pop                 &   1636 \\
-%Hip-Hop             &   1059 \\
-%Folk                &   1025 \\
-%Punk                &    882 \\
-%Indie-Rock          &    601 \\
-%Jazz                &    588 \\
-%Old-Time / Historic &    460 \\
-%Psych-Rock          &    456 \\
-%International       &    431 \\
-%Classical           &    408 \\
-%Chiptune            &    254 \\
-%Blues               &    245 \\
-%Psych-Folk          &    245 \\
-%Post-Punk           &    222 \\
-%Metal               &    191 \\
-%Soundtrack          &    172 \\
-%Trip-Hop            &    171 \\
-%Post-Rock           &    118 \\
-%\bottomrule
-%\end{tabular}
-%\caption{xxx}
-%\label{table3}
-%\end{table}
 
@@
 \centering
-\includegraphics[width=10cm]{histo3.pdf}
-\caption{Histograms of the 10 music genres for the small FMA dataset of 1000 songs.}
-\label{fig3}
+\includegraphics[height=5cm]{histo_small.pdf}
+\vspace{-0.5cm}
+\caption{The 10 music genres for the FMA\_small dataset with 4,000 songs.}
+\label{fig_small}
 \end{figure}
@@
  fma\_id &                                             artist &  play\_count &                                              title &                                             genres &            top\_genre  \\
+ \toprule
     10 &                                          Kurt Vile &       42936 &                                            Freeway &                                              [Pop] &                  Pop  \\

### 2016-12-06 paper_FMA_nov2016.tex
xav_arxiv_06dec16

@@
 In this paper, we introduce the Free Music Archive (FMA) which contains 77,643 songs and 68 genres spanning 26.9 days of song listening and meta-data including artist name, song title, music genre, and track counts.
-For research purposes, we define two additional datasets from the original one: a small balanced-genre dataset of 4,000 song data and 10 genres compassing 33.3 hours of raw audio and a medium unbalanced-genre dataset of 14,511 data and 20 genres offering 5.1 days of track listening, both datasets come with meta-data and Echonest audio features.
+For research purposes, we define two additional datasets from the original one: a small genre-balanced dataset of 4,000 song data and 10 genres compassing 33.3 hours of raw audio and a medium genre-unbalanced dataset of 14,511 data and 20 genres offering 5.1 days of track listening, both datasets come with meta-data and Echonest audio features.
 For all datasets, we provide a train-test splitting for future algorithms' comparisons.
@@
 It has been widely used for different music applications.
-A complete study of this dataset has been done in \cite{art:Sturm12GTZAN}, and the dataset is available at \cite{art:GTZAN}.
+A complete study of this dataset was done in \cite{art:Sturm12GTZAN}, and the dataset is available at \cite{art:GTZAN}.
 The main limitations of GTZAN is the legality of the dataset, the small size, no complete meta-data regarding artist names and song titles, and no additional meta-data like ratings.
@@
 This prevents the application of deep learning on this dataset.
-Finally, Echonest reassigned the indexing of the MDS songs with its database, which from now makes (almost) impossible the use of the most music recent features developed by Echonest.
+Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
 \\
@@
 {\bf What would be the requirements for an ideal music dataset?}  \\
-(1) Availability of raw audio data like 30sec or more\\
-(2) Availability of meta-data such as genres, artist, title, year, lyrics, users' ratings, users' comments, etc\\
+(1) Availability of raw audio tracks like 30sec or more,\\
+(2) Availability of meta-data such as genres, artist, title, year, lyrics, users' ratings, users' comments, etc,\\
 (3) Large-scale dataset to be able to sample the high-dimensional distribution of music genres\\
-(4) Free and easy access of the dataset\\
-(5) Legality of the dataset, no copyright issue with e.g. open access licence\\
+(4) Free and easy access of the dataset,\\
+(5) Legality of the dataset, no copyright issue with e.g. open access licence.\\
 
@@
 The meta-information collected is artist name, track title, track genres, and track listens.
-We removed data without genre, giving 79,947 song data.
-Then, we created a top genre by simply picking the first genre in the original list of track genres.
+We removed data without genre information, giving 79,947 song data.
+Then, we created a top genre by simply picking the first genre in the original list of track genres provided by the artists.
 A few examples of data are given in Table \ref{tab1}.
 At this stage the number of genres is 138, but half have only a few samples.
-Thus we decided to retain the genres with more than 100 samples.
+Thus we decided to only retain the genres with more than 100 samples.
 This defines a dataset of 77,643 music data with 68 genres.
@@
 The GTZAN data may not have for example song title or artist name, and also no copyright permission.
-For reproducibility research, we also provide a split 80-20 of this dataset for future techniques' comparisons.
+For reproducibility research, we also provide a train-test split 80-20 of this dataset for future techniques' comparisons.
 \\
@@
 All songs come with the following meta-data: artist name, track title, track genres, track listens.
-Eventually, we give a train-test split 80-20 of this dataset.\\
+Eventually, we give a split 80-20 of this dataset.\\
 
@@
 The FMA\_large dataset of 77,643 data contains songs of 30sec length, and artist name, track title, track genres, track listens as meta-data.
-Our original motivation was to give the full audio tracks, not limited to 30sec, but the size of the whole music tracks is 806.8GB (for approximatively 256 days of music listening) makes the hosting a challenging issue.
+Our original motivation was to give the full audio tracks, not limited to 30sec, but the size of the whole music tracks is 806.8GB (for approximatively 256 days of music listening) making the hosting a challenging issue.
 

### 2017-02-15 paper_FMA_nov2016.tex
Acknowledgments
update web links

@@
 \usepackage{fancyhdr}
-\usepackage{hyperref}
+\usepackage[colorlinks=true,allcolors=blue]{hyperref}
 \usepackage[top=1in,bottom=2in,left=1in,right=1in]{geometry}
@@
 
+% TODO
+% Problems
+% Genre classif / mutli-genre classif
+% Recommendation
+% Play count prediction
+
 %%%%%%%%%%%%%%%%%%%
@@
 
-\noindent
-{\bf Availability and reproducible research.} The 3 proposed FMA datasets are available at\\
+\paragraph{Availability and reproducible research.}
 
- \url{https://lts2.epfl.ch/datasets/fma}\\
-
- \noindent
- We also provide IPython notebooks at this address\\
+All the datasets as well as Jupyter notebooks used to produce them, the
+analysis, and baselines as are available online at
 
@@
 
- \noindent
- to load the datasets and to start using them for research purposes.\\
+ Code to support the use of the data is provided as well.
 
-\noindent
-{\bf Potential applications.} We foresee multiple applications of this new dataset.
+\paragraph{Potential applications.} We foresee multiple applications of this new dataset.
 The most obvious one is for music analysis with deep learning techniques.
@@
 Besides, standard data analysis tasks such as classification, clustering, regression, text analysis, visualization can also be applied to this new music dataset.\\
+
+\section*{Acknowledgments}
+
+We are grateful to SWITCH and EPFL for hosting the dataset within the context
+of the \href{https://projects.switch.ch/scale-up/}{SCALE-UP} project, funded in
+part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC
+P-2 program}.
 

### 2017-03-16 fma.tex
paper update

--- 2017-02-15 paper_FMA_nov2016.tex
@@
-\documentclass[a4paper,12pt]{article}
-\usepackage{amsfonts}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{array}
-\usepackage{fancyhdr}
-\usepackage[colorlinks=true,allcolors=blue]{hyperref}
-\usepackage[top=1in,bottom=2in,left=1in,right=1in]{geometry}
-\usepackage{graphicx} \graphicspath{{figures/}}
-\usepackage{subfigure}
+% -----------------------------------------------
+% Template for ISMIR Papers
+% 2017 version, based on previous ISMIR templates
+
+% Requirements :
+% * 6+n page length maximum
+% * 4MB maximum file size
+% * Copyright note must appear in the bottom left corner of first page
+% * Clearer statement about citing own work in anonymized submission
+% (see conference website for additional details)
+% -----------------------------------------------
+
+\documentclass{article}
+\usepackage{ismir,amsmath,cite,url}
+\usepackage{graphicx} \graphicspath{{figs/}}
 \usepackage{color}
-\usepackage[normalem]{ulem}
 
+\usepackage[utf8]{inputenc}
+\usepackage[bookmarks=false,colorlinks=true,allcolors=blue]{hyperref}
 \usepackage{booktabs}
+\usepackage{footnote} \makesavenoteenv{table}
 \usepackage{rotating,tabularx}
 
-%\newcommand{\Mn}[1]{ \mathrm{M}_{#1}(\mathbb{R}) }
-%\newcommand{\lex}{ \mathrm{lex} }
-%\newcommand{\vect}{ \mathrm{vec} }
-%\newcommand{\R}{\mathbb{R}}
-%\newcommand{\x}{\mathbf{x}}
-%\newcommand{\vv}{\mathbf{v}}
-%\newcommand{\uu}{\mathbf{u}}
-%\newcommand{\e}{\mathbf{e}}
-%\newcommand{\rd}{\mathrm{d}}
-%\newcommand{\re}{\mathrm{e}}
-%\newcommand{\E}{\mathbb{E}}
-%\newcommand{\Prob}{\mathbb{P}}
-%\newcommand{\op}{\mathrm{op}}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{example}[theorem]{Example}
-%\newtheorem{proposition}[theorem]{Proposition}
-%\newtheorem{corollary}[theorem]{Corollary}
-%\newtheorem{remark}[theorem]{Remark}
-%\newtheorem{definition}[theorem]{Definition}
-%\newtheorem{example}[theorem]{Example}
-
-\newcommand{\diag}{\mathrm{diag}}
-\newcommand{\tr}{\mathrm{tr}}
-
-%\title{Free Music Archive (FMA): \\ A Dataset For Music Analysis}
 \title{FMA: A Dataset For Music Analysis}
 
-\author{Kirell Benzi, Micha\"el Defferrard, Pierre Vandergheynst, Xavier Bresson\\ \vspace{-0.5cm} \\  LTS2, EPFL, Lausanne, Switzerland}
+\oneauthor
+{Kirell Benzi, \hspace{.2cm} Michaël Defferrard, \hspace{.2cm} Pierre Vandergheynst, \hspace{.2cm} Xavier Bresson}
+ {{\tt michael.defferrard@epfl.ch, kirell.benzi@epfl.ch} \\
+  {\tt pierre.vandergheynst@epfl.ch, xavier.bresson@epfl.ch} \\
+  LTS2, EPFL, Lausanne, Switzerland}
 
-\date{\today}
+%\threeauthors
+%  {First Author} {Affiliation1 \\ {\tt author1@ismir.edu}}
+%  {Second Author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
+%  {Third Author} {Affiliation3 \\ {\tt author3@ismir.edu}}
 
+%% To make customize author list in Creative Common license, uncomment and customize the next line
+\def\authorname{Kirell Benzi, Michaël Defferrard, Pierre Vandergheynst, Xavier Bresson}
+
+\sloppy % please retain sloppy command for improved formatting
 \begin{document}
+\maketitle
 
-%%%%%%%%%%%%%%%%%%%
-\maketitle
-%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%
-\abstract{We present a new music dataset that can be used for several music analysis tasks.
+% about 150-200 words.
+\begin{abstract}
+We present a new music dataset that can be used for several music analysis tasks.
 Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of meta-data for artists analysis or song ratings for recommender systems.
@@
 For all datasets, we provide a train-test splitting for future algorithms' comparisons.
-\\\\
-{\it Keywords.} Music dataset, Free Music Archive (FMA), open science, reproducibility research, music analysis.} 
-%%%%%%%%%%%%%%%%%%%
+\end{abstract}
+
+The released dataset is composed of 100,000 tracks, xx artists, xx albums, xx hours of audio (xx GiB), 164 genres, all under permissive Creative Commons licenses.
+
+The primary goal of this dataset is to help researchers close the semantic gap between the computed low-level features such as mfcc and the high-level targets such as genre.
+Hopefully dl can help, but needs lot of data.
 
@@
 Evaluation and comparison require benchmark datasets to achieve their purpose, which are sometimes challenging to collect and share.
-In computer vision, the community has developed over the years well-established benchmark datasets such as MNIST \cite{pro:LeCunBottouBengioHaffner98MNIST} publicly accessible at \cite{art:MNIST}, CIFAR \cite{art:Krizhevsky09CIFAR} available at \cite{art:CIFAR}, or ImageNet \cite{art:FeiFei09ImageNet} provided at \cite{art:ImageNet}.
+In computer vision, the community has developed over the years well-established benchmark datasets such as \href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist}, \href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR} \cite{cifar}, or \href{http://www.image-net.org}{ImageNet} \cite{imagenet}.
 All these image datasets are free, legal and easily available on Internet.
@@
 For examples, developing deep learning techniques for music applications require the use of raw audio data, and designing recommender systems need meta-data like artist features and song ratings.
-Besides, the most influential competition in the field of music analysis organized every year is the Music Information Retrieval Evaluation eXchange (MIREX) \cite{art:MIREX}.
-MIREX proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc.
+Besides, the most influential competition in the field of music analysis organized every year is the Music Information Retrieval Evaluation eXchange (MIREX).\footnote{\url{http://www.music-ir.org/mirex/wiki/}} MIREX proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc.
 However, participants do not have access to neither the test set, nor the important train set.
@@
 \noindent
-{\bf Available music datasets and their limitations.} In music analysis, three datasets have been widely used by the international society for music information retrieval (ISMIR) \cite{art:ISMIR}: the GTZAN, TagATune, and MSD (Million Song Dataset) datasets.
-However each dataset come with some limitations: \\
-$\bullet$ GTZAN \cite{art:TzanetakisCook02GTZAN} is a collection of 1000 songs with 10 music genres.
+{\bf Available music datasets and their limitations.} In music analysis, three datasets have been widely used by the international society for music information retrieval (ISMIR)\footnote{\url{http://www.ismir.net/}}: the GTZAN, TagATune, and MSD (Million Song Dataset) datasets.
+However each dataset come with some limitations:
+We discuss below some of the most related datasets, see \tabref{tab:size} for a more comprehensive list.
+\begin{description}
+	\item[GTZAN] \cite{gtzan} is a collection of 1000 songs with 10 music genres.
 Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
-It is the most popular music dataset with 6,600 answers on Google search, and 600 on Google scholar on Nov 2016.
+It is the most popular music dataset \cite{mgr_eval}, with 6,600 answers on Google search, and 600 on Google scholar on Nov 2016.
 It has been widely used for different music applications.
-A complete study of this dataset was done in \cite{art:Sturm12GTZAN}, and the dataset is available at \cite{art:GTZAN}.
+A complete study, and critic, of this dataset was done in \cite{gtzan_critic_1}, and the dataset is available online.
 The main limitations of GTZAN is the legality of the dataset, the small size, no complete meta-data regarding artist names and song titles, and no additional meta-data like ratings.
-\\
-$\bullet$ TagATune \cite{art:Law07TagATune} was a popular dataset of 5,405 source songs of 230 artists.
-It was used for music tagging and music metric learning.
-Its website is \cite{art:TagATune}, but it is not anymore publicly available since 2009.\\
-$\bullet$ The Million Song Dataset (MSD) \cite{art:Bertin11MSD} is a free and legal collection of 1 million music songs.
+	\item[MagnaTagATune] \cite{magnatagatune} is a popular dataset of 5,405 source songs from 230 artists, cut in 25,863 clips.
+The music was collected from the \href{https://magnatune.com/}{Magnatune} label and was tagged using the \href{http://tagatune.org/}{TagATune} game.
+The dataset includes meta-data, audio features and raw audio, although the 16 kHz, 32kbps, mono mp3 are of poor audio quality.
+	\item[The Million Song Dataset (MSD)] \cite{msd} is a free and legal collection of 1 million music songs.
 Each song data is composed of high-level and medium-level audio features provided by  Echonest service and meta-data like artist name, song title, track genre, etc.
-It is available at \cite{art:MSD}.
+It has its own dedicated website with related code and additional datasets.
 The very large scale of this music dataset would make it an ideal candidate for deep learning, which works best with large datasets.
@@
 Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
-\\
+		While researchers have been able to download sample audio from the online service \url{7signal.com} and extract features by themselves, see e.g. \cite{msd_features}, the process is tedious.
+	\item[AudioSet] \cite{audioset} is a dataset of 2.1 million annotated 10-second sound clips associated with 527 classes.
+The ontology is quite broad and covers sounds from music to cars, engines or animals.
+The raw audio is however not included and has to be downloaded from Youtube.
+It is not clear of what happens if a referenced video gets deleted.
+\end{description}
 
-\newpage
+\begin{table}
+	\centering
+	\begin{tabular}{lrrcc}
+		\toprule
+		dataset & \#clips & \#artists & year & audio \\
+		\midrule
+		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC} \cite{RWC} & 465 & - & 2001 & yes \\
+		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500} \cite{cal500} & 500 & 500 & 2007 & yes \\
+		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
+		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
+		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
+		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\  % Less common
+		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{GarageBand} \cite{garageband} & 1,886 & - & 2005 & yes \\  % Less common
+		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\ % Less common
+		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{Extended Ballroom} \cite{extballroom} & 4,180 & - & 2016 & yes\footnote{Audio has to be downloaded from \url{www.ballroomdancers.com}.} \\
+		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
+		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k / Swat10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\
+		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} & 25,863\footnote{The 25,863 clips come from 5,405 source mp3.} & 230 & 2009 & yes \\
+		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
+		FMA & 110,000 & - & 2007 & yes \\
+		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
+		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\footnote{Audio can be downloaded from \url{7signal.com}.} \\
+		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,100,000 & - & 2017 & yes\footnote{Audio has to be downloaded from \url{youtube.com}.} \\
+		\bottomrule
+	\end{tabular}
+	\caption{Size comparison with some other datasets.}
+	\label{tab:size}
+\end{table}
+
+% Lists of datasets:
+% * http://www.audiocontentanalysis.org/data-sets/
+% * A Survey of Evaluation in Music Genre Recognition
+
+% Todo year, licensing
+
+Our differentiating factors:
+\begin{description}
+	\item[High quality audio]: full-length original audio uploaded by the artist. Stereo mp3 with bitrates varying between x.
+	\item[Permissive licensing]: all songs are licensed under Creative Commons by the artists.
+	\item[Future proof]: all audio and meta-data have a DOI and are archived on \url{zenodo.org}, a long-term digital archive.
+No risk for a webservice to disappear or a website to be shut down.
+All audio can be downloaded as a single 900GB zip, no need to scrap the web yourself.
+	% Western / commercial music, how to qualify broadness?
+	\item[Reproducible]: we share all the code used to create the dataset, so that it can be reproduced or extended. A train / test split is provided to make research on the FMA reproducible.
+\end{description}
+
+% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
+% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
+
+% TODO: talk about related datasets
+% * SecondHandSongs dataset -> cover songs
+% * musiXmatch dataset -> lyrics
+% * Last.fm dataset -> song-level tags and similarity
+% * Taste Profile subset -> user data
+% * thisismyjam-to-MSD mapping -> more user data
+% * tagtraum genre annotations -> genre labels
+% * Top MAGD dataset -> more genre labels
+
+We see from \tabref{tab:size} that most of the smallest datasets come with audio while most of the largest don't.
+That state of affair is due to copyright laws which forbid the redistribution of audio.
+The solution to that problem is to aim for tracks released under permissive licenses, such as Creative Commons\footnote{\url{https://creativecommons.org/}}.
+
+Creative Commons: MagnaTagATune
+Proprietary, non-redistributable: MSD, AudioSet?
+
 \noindent
@@
 \noindent
-{\bf The website.} The FMA website \cite{art:FMA} is run by WFMU, the longest-running free-form radio station in the United States.
+{\bf The website.} The FMA website\footnote{\url{https://freemusicarchive.org/}} is run by WFMU, the longest-running free-form radio station in the United States.
 The website provides a large catalogue of artists and high-quality mp3 songs.
@@
 \noindent
-{\bf Crawling the data.} The FMA organization provides an API \cite{art:FMA_API} to collect the meta-data associated to each song such as artist name, track title, track genre, and track listens.
+{\bf Crawling the data.} The FMA organization provides an API\footnote{\url{https://freemusicarchive.org/api}} to collect the meta-data associated to each song such as artist name, track title, track genre, and track listens.
 The collect of the mp3 audio files was done with a web scrapper.
@@
 Finally, we kept the songs that have the raw audio and all Echonest features.
-Echonest features are medium-level and high-level audio features provided by Echonest \cite{art:Echonest}, a music platform for developers and media companies which has a collection of 3.5M artists and 37.5M songs.
+Echonest features are medium-level and high-level audio features provided by Echonest\footnote{\url{http://the.echonest.com/}}, a music platform for developers and media companies which has a collection of 3.5M artists and 37.5M songs.
 Echonest features can be either computed by uploading the songs to the Echonest website, which are then analyzed and send them back to the user, or simply retrieved from the Echonest database if the songs were analyzed before.
-All Echonest operations are done through an API \cite{art:Echonest_API}.
+All Echonest operations are done through an API\footnote{\url{http://developer.echonest.com}}.
 Our final filtering consists in keeping songs that have both raw audio file and complete Echonest features already computed and stored in the Echonest database.
@@
 
+Most tracks are released under a Creative Commons license\footnote{\url{https://creativecommons.org/}} which allows redistribution of the audio.
+
+\subsection{Features}
+
+Extracted with librosa. cite librosa
+
+\section{Proposed tasks} % usage, problems
+
+genre classification, mood classification, artist identification/recognition, instrument recognition, music annotation (tags),
+Artist Identification, Genre classification, Mood Classification, Instrument identification, Music Similarity, Autotagging, Automatic playlist generation
+
+Can be down: genre, artist, popularity prediction, tags, year prediction
+
+Music genre recognition (MGR) is one of the most researched areas of MIR.
+
+We propose three genre classification problems of varying difficulty.
+
+\begin{enumerate}
+	\item Easy, based on \texttt{fma\_small.zip}: 10 top-level genres, balanced with 400 tracks per genre, 1 genre per track
+	\item Medium, based on \texttt{fma\_medium.zip}: all 16 top-level genres, unbalanced with x-y tracks per genre, multiple genres per track
+	\item Hard, based on \texttt{fma\_large.zip}: all 164 genres, unbalanced with x-y tracks per genre, multiple genres per track
+\end{enumerate}
+
+Training / testing division.
+
+\section{Genre Recognition} % MGR, Baselines
 %%%%%%%%%%%%%%%%%%%%
@@
 
-\paragraph{Availability and reproducible research.}
-
+%\paragraph{Availability and reproducible research.}
+{\bf Availability and reproducible research.}
 All the datasets as well as Jupyter notebooks used to produce them, the
@@
 
- \url{https://github.com/mdeff/fma}\\
+\url{https://github.com/mdeff/fma}
 
@@
 
-\paragraph{Potential applications.} We foresee multiple applications of this new dataset.
+%\paragraph{Potential applications.}
+{\bf Potential applications.}
+We foresee multiple applications of this new dataset.
 The most obvious one is for music analysis with deep learning techniques.
-As raw music tracks are available, deep learning techniques like convolutional neural networks \cite{pro:LeCunBottouBengioHaffner98MNIST} and recurrent neural networks \cite{art:HochreiterSchmidhuber97LSTM} can be directly applied.
-Besides, standard data analysis tasks such as classification, clustering, regression, text analysis, visualization can also be applied to this new music dataset.\\
+As raw music tracks are available, deep learning techniques like convolutional neural networks \cite{mnist} and recurrent neural networks \cite{art:HochreiterSchmidhuber97LSTM} can be directly applied.
+Besides, standard data analysis tasks such as classification, clustering, regression, text analysis, visualization can also be applied to this new music dataset.
 
-\section*{Acknowledgments}
+\section{Acknowledgments}
 
@@
 
-%%%%%%%%%%%%%%%%%%%
-\vspace{1cm}
-\bibliography{bib_paper}
-\bibliographystyle{plain}
-%%%%%%%%%%%%%%%%%%%
+We are grateful to the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributors, be they artists or curators, for the fantastic content.
 
-\newpage
+\bibliography{refs}
 
@@
 \end{figure}
-
-\newpage
 

### 2017-03-17 fma.tex
paper update

@@
 
+% Anonymity: authors, GitHub link
+
 \documentclass{article}
@@
 \usepackage{booktabs}
-\usepackage{footnote} \makesavenoteenv{table}
+%\usepackage[para,online,flushleft]{threeparttable}
+\usepackage{threeparttable}
+%\usepackage{footnote} \makesavenoteenv{table}
 \usepackage{rotating,tabularx}
@@
 
-The primary goal of this dataset is to help researchers close the semantic gap between the computed low-level features such as mfcc and the high-level targets such as genre.
-Hopefully dl can help, but needs lot of data.
+The dataset comes with a collection of meta-data such as the song titles, artists and albums, together with full-length audio and a set of pre-computed features as well as a train/test split.
+
+Content-based music information retrieval tasks are typically solved with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
+Although that approach was dominant in the past, feature learning and end-to-end learning (from signal to label) have started to receive more attention from the MIR community in recent years. 
+The primary goal of this dataset is to enable researchers to train large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
+
+This paper describes the dataset and how it was created, proposes some tasks and evaluate some baselines for music genre recognition (MGR).
+All the code to reproduce it as well as links to download the data and usage examples are available at \url{https://github.com/mdeff/fma}.
 
 %%%%%%%%%%%%%%%%%%%
-\section{Introduction and Motivation}
+\section{Introduction} % and Motivation}
 %%%%%%%%%%%%%%%%%%%
@@
 
+In contrast with other modalities such as images or text for which a wealth of content is available, the lack of a large-scale and easily available dataset for MIR has hindered research on data-heavy models such as Deep Learning (DL).
+
 \noindent
@@
 However each dataset come with some limitations:
-We discuss below some of the most related datasets, see \tabref{tab:size} for a more comprehensive list.
+We discuss below some of the most related datasets, see \tabref{tab:datasets} for a more comprehensive list.
+\tabref{tab:datasets} shows the most common datasets used for audio-based / content-based MIR.
 \begin{description}
@@
 Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
-It is the most popular music dataset \cite{mgr_eval}, with 6,600 answers on Google search, and 600 on Google scholar on Nov 2016.
-It has been widely used for different music applications.
+It is the first benchmark MGR dataset to have been made publicly available, and as a result continues to be the most used public dataset for MGR \cite{mgr_eval}.
 A complete study, and critic, of this dataset was done in \cite{gtzan_critic_1}, and the dataset is available online.
@@
 Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
-		While researchers have been able to download sample audio from the online service \url{7signal.com} and extract features by themselves, see e.g. \cite{msd_features}, the process is tedious.
+		While researchers have been able to download sample audio from the online service \url{7signal.com} and extract features by themselves, the process is tedious \cite{msd_features}.
 	\item[AudioSet] \cite{audioset} is a dataset of 2.1 million annotated 10-second sound clips associated with 527 classes.
@@
 	\centering
+	\begin{threeparttable}
 	\begin{tabular}{lrrcc}
@@
 		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
+		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 729 & - & 2004 & yes \\
 		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
 		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
-		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\  % Less common
-		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{GarageBand} \cite{garageband} & 1,886 & - & 2005 & yes \\  % Less common
-		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\ % Less common
-		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{Extended Ballroom} \cite{extballroom} & 4,180 & - & 2016 & yes\footnote{Audio has to be downloaded from \url{www.ballroomdancers.com}.} \\
+		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
+		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{garageband} & 1,886 & - & 2005 & yes \\  % GarageBand
+		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} & 3,115 & - & 2010 & no \\
+		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} & 3,180 & 1,517 & 2010 & yes \\
+		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
+		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{1} \\
 		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
-		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k / Swat10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\
-		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} & 25,863\footnote{The 25,863 clips come from 5,405 source mp3.} & 230 & 2009 & yes \\
+		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
+		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{2} & 230 & 2009 & yes\tnote{3} \\ % MagnaTagATune
 		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
-		FMA & 110,000 & - & 2007 & yes \\
+		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf 110,000 & \bf - & \bf 2017 & \bf yes \\
 		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
-		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\footnote{Audio can be downloaded from \url{7signal.com}.} \\
-		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,100,000 & - & 2017 & yes\footnote{Audio has to be downloaded from \url{youtube.com}.} \\
+		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{1} \\
+		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,100,000 & - & 2017 & no\tnote{1} \\
 		\bottomrule
 	\end{tabular}
-	\caption{Size comparison with some other datasets.}
-	\label{tab:size}
+	\begin{tablenotes}
+		\item[1] Audio can be downloaded from a web service. % \url{www.ballroomdancers.com} \url{7signal.com} \url{youtube.com}
+		\item[2] The 25,863 clips come from 5,405 songs.
+		\item[3] Low quality 16 kHz, 32kbps, mono mp3.
+	\end{tablenotes}
+	\end{threeparttable}
+	\caption{Comparison with some other datasets.}
+	\label{tab:datasets}
 \end{table}
@@
 
-% Todo year, licensing
+% TODO licensing
+
+\begin{table}
+	\centering
+	\begin{threeparttable}
+	\begin{tabular}{lrrcc}
+		\toprule
+		dataset & \#clips & length\tnote{1} & \#days\tnote{2} & size\tnote{3} \\
+		\midrule
+		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863 & 29 & 8.7 & 3 \\ % MagnaTagATune
+		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,100,000 & 10 & 243 & - \\
+		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 30, 60\tnote{4} & 541 & 625 \\
+		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf 110,000 & \bf 10-180 & \bf ? & \bf 900 \\
+		\bottomrule
+	\end{tabular}
+	\begin{tablenotes}
+		\item[1] Length of available clips in seconds.
+		\item[2] Number of days to listen to the whole available audio.
+		\item[3] Size of the audio (accounts for duration and quality).
+		\item[4] Most excerpts from 7digital are 30 or 60s \cite{msd_features}.
+	\end{tablenotes}
+	\end{threeparttable}
+	\caption{Audio size comparison.}
+	\label{tab:size}
+\end{table}
+
+% table quality
 
@@
 \begin{description}
-	\item[High quality audio]: full-length original audio uploaded by the artist. Stereo mp3 with bitrates varying between x.
-	\item[Permissive licensing]: all songs are licensed under Creative Commons by the artists.
-	\item[Future proof]: all audio and meta-data have a DOI and are archived on \url{zenodo.org}, a long-term digital archive.
-No risk for a webservice to disappear or a website to be shut down.
-All audio can be downloaded as a single 900GB zip, no need to scrap the web yourself.
+	\item[Permissive licensing.] MIR research has historically suffered from the lack of publicly available benchmark datasets.
+Most of these issues stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
+All songs are licensed under Creative Commons by the artists and can thus be redistributed.
+* In contrast with images or text for which plenty of data is available online
+	\item[Audio.] most of the largest datasets are distributed with features, sometimes computed by commercial services like the Echonest.
+Researchers are wary of those proprietary features and have seek to compute features themselves, e.g.
+on the MSD \cite{msd_features}.
+In contrast, FMA provides audio for each song.
+	\item[High quality audio.] full-length original audio uploaded by the artist.
+Stereo mp3 with bitrates varying between x.
+Audio, if released or available to download, is usually clips of 10 to 30 seconds.
+To our knowledge, none of the released datasets up to this date come with full-length audio.
+	\item[Future proof]: all audio and meta-data can be downloaded as ZIP archives, no need to scrap the web yourself.
+They are checksummed, have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by CERN.
+We thus alleviate the risks of tracks to become unavailable, to change without notice or for online services to shut down.
 	% Western / commercial music, how to qualify broadness?
-	\item[Reproducible]: we share all the code used to create the dataset, so that it can be reproduced or extended. A train / test split is provided to make research on the FMA reproducible.
+	\item[Reproducible]: we share all the code used to (i) collect the dataset, (ii) analyze it, (iii) extract the sub-sets, (iv) compute the features and (v) the baselines, so that it can be reproduced or extended.
+As the original data and APIs are public, anybody can recreate the dataset.
+A train / test split is provided to make research on the FMA reproducible.
+	\item[Easy to work with]: working with the dataset only requires to download some archives.
+Moreover, we provide an helper Python package\footnote\url{https://pypi.python.org/pypi/freemusicarchive} and usage examples as Jupyter notebooks to quick start using the data.
 \end{description}
@@
 
-We see from \tabref{tab:size} that most of the smallest datasets come with audio while most of the largest don't.
+The sets in \tabref{tab:datasets} are either smaller, either provide audio through online services.
+
+From \tabref{tab:datasets}, it is obvious that the main problem is that larger datasets do not come with the audio.
+Either they contain only features derived from the audio, either they provide links for the user to download the audio from an online service.
+The problem with the former is that researchers are stuck with the set of chosen features, while the later make no assurance that the files or services are not gonna disappear or change without notice.
+We see from \tabref{tab:datasets} that most of the smallest datasets come with audio while most of the largest don't.
 That state of affair is due to copyright laws which forbid the redistribution of audio.
@@
 %%%%%%%%%%%%%%%%%%%
-\section{Introducing the FMA dataset}
+\section{The Dataset} %Introducing the FMA dataset}
 %%%%%%%%%%%%%%%%%%%
+
+% Describe what it is, what are the meta-data and features.
+
+% TODO audio properties: samplerate (% 44kHz), bitrate (% 128, VBR), channels (% mono, stereo)
+Most tracks are released under a Creative Commons license\footnote{\url{https://creativecommons.org/}} which allows redistribution of the audio.
+
+\subsection{Analysis}
+% Figures and tables from analysis.ipynb
+
+\subsection{Features} % Feature sets
+
+Extracted with librosa\footnote{\url{https://github.com/librosa/librosa}}.
+
+In addition, a number of descriptive features extracted with the services from The Echo Nest are provided.
+These features include tempo, loudness, timings of fade-in and fade-out, and MFCC-like features for a number of segments.
+
+\subsection{Collection} % Creation, Generation
+% How it was collected.
 
@@
 
-Most tracks are released under a Creative Commons license\footnote{\url{https://creativecommons.org/}} which allows redistribution of the audio.
-
-\subsection{Features}
-
-Extracted with librosa. cite librosa
-
 \section{Proposed tasks} % usage, problems
+% List some MIR tasks and propose some who can be done with that dataset
 
@@
 
-Music genre recognition (MGR) is one of the most researched areas of MIR.
+% TODO
+% Problems
+% Genre classif / mutli-genre classif
+% Recommendation
+% Play count prediction
+% Covers: generate covers per genre / artist
+
+Meta-data analysis: which genres are artists contributing to, drift in time, curators.
+It would also be interesting to study how good does it represent mainstream music.
+
+As it happened with e.g. the MSD, people may find and link other sources of data with the dataset, such as lyrics or tags associated to the songs, e.g. from last.fm.
 
@@
 \section{Genre Recognition} % MGR, Baselines
-%%%%%%%%%%%%%%%%%%%%
-%\section{Classification Experiment [MICHAEL SECTION LATER LATER]}
-%%%%%%%%%%%%%%%%%%%%
-%
+% One of the most common task, we give some baselines so that people have a reference about how hard it is.
+
+Music genre recognition (MGR) is one of the most researched areas of MIR.
+
 %\noindent
@@
 
-% TODO
-% Problems
-% Genre classif / mutli-genre classif
-% Recommendation
-% Play count prediction
-
 %%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%
+
+The FMA enables researchers to test algorithms on a large-scale collection, thus allowing to test them on more real-world like environments.
+By providing audio, we do not limit the benchmarking to pre-computed features and allow scientists to develop and test new feature sets, learn features, or learn mappings directly from the audio.
 
@@
 
+While this data release is pretty large, it is still two orders of magnitude behind commercial services who have access to tens of millions of songs.\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}}
+
+% last.fm: 45 million http://www.skilledtests.com/wiki/Last.fm_statistics
+
 \section{Acknowledgments}
@@
 
-We are grateful to the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributors, be they artists or curators, for the fantastic content.
+We are grateful to the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
 

### 2017-03-17 fma.tex
size comparison between datasets

@@
 \usepackage{rotating,tabularx}
+
+\usepackage{array}
+
+\usepackage{color}
+\newcommand{\todo}[1]{{\color{red} #1 }}
 
@@
 The primary goal of this dataset is to enable researchers to train large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
+Hopefully, the release of such amount of easily accessible data will foster research which will ultimately surpass the performance ceiling observed in e.g.
+genre recognition, as it as been the case in computer vision with the release of ImageNet.
 
@@
 
-%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction} % and Motivation}
-%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
-		While researchers have been able to download sample audio from the online service \url{7signal.com} and extract features by themselves, the process is tedious \cite{msd_features}.
+		While researchers have been able to download sample audio from the online service \href{https://www.7digital.com}{7digital.com} and extract features by themselves, the process is tedious \cite{msd_features}.
 	\item[AudioSet] \cite{audioset} is a dataset of 2.1 million annotated 10-second sound clips associated with 527 classes.
@@
 	\begin{tablenotes}
-		\item[1] Audio can be downloaded from a web service. % \url{www.ballroomdancers.com} \url{7signal.com} \url{youtube.com}
+		\item[1] Audio can be downloaded from a web service. % \url{www.ballroomdancers.com} \url{7digital.com} \url{youtube.com}
 		\item[2] The 25,863 clips come from 5,405 songs.
@@
 
-\begin{table}
+\begin{table*}[t]
 	\centering
 	\begin{threeparttable}
-	\begin{tabular}{lrrcc}
+		\begin{tabular}{lr
+			>{$}r<{$}
+			@{}>{${}={}$}c@{}
+			>{$}l<{$}
+			>{$}c<{$}
+			cc}
 		\toprule
-		dataset & \#clips & length\tnote{1} & \#days\tnote{2} & size\tnote{3} \\
+		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \text{scale} & size [GB] & size [days] \\
 		\midrule
-		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863 & 29 & 8.7 & 3 \\ % MagnaTagATune
-		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,100,000 & 10 & 243 & - \\
-		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 30, 60\tnote{4} & 541 & 625 \\
-		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf 110,000 & \bf 10-180 & \bf ? & \bf 900 \\
+		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} &
+			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\ % MagnaTagATune
+		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
+			2,100,000 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & - & 243 \\
+		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
+			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 541 \\
+		\href{https://github.com/mdeff/fma/}{FMA} \tnote{3} &
+			110,000 & 180\cdot44\cdot10^3\cdot2 && 1.6\cdot10^7 & 1.7\cdot10^{12} & 900 & 229 \\
+		\midrule
+		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
+			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
+		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
+			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
+		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{4} &
+			14,000,000 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.4\cdot10^{12} & 1000 & - \\
 		\bottomrule
@@
 	\begin{tablenotes}
-		\item[1] Length of available clips in seconds.
-		\item[2] Number of days to listen to the whole available audio.
-		\item[3] Size of the audio (accounts for duration and quality).
-		\item[4] Most excerpts from 7digital are 30 or 60s \cite{msd_features}.
+		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos.
+		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
+		\item[3] Average duration and sample rate. See \todo{ad} for details.
+		\item[4] Average resolution. Although most applications resize or crop to $256\cdot256$.
 	\end{tablenotes}
 	\end{threeparttable}
-	\caption{Audio size comparison.}
+	\caption{Size comparison.
+Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images.
+Scale is the number of samples times the dimensionality.
+Size is for a (zipped) archive of all mp3 or jpeg in GB, which is an indication of the quantity of information, e.g.
+duration or quality.
+The last column is the number of days to listen to the whole available audio.}
 	\label{tab:size}
-\end{table}
+\end{table*}
 
@@
 \begin{description}
+	\item[Large scale.] Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with genre.
+While FMA is not the largest in term of number of clips, every other dataset with readily available quality audio are two orders of magnitude smaller.
+A possibly more important size metric here is the total amount of audio, i.e.
+the sum of the lengths of all clips, in which FMA is in a much better position (see \tabref{tab:size}).
+For comparison, ImageNet \cite{imagenet}, the largest benchmark dataset in computer vision features 14 million images.
+Despite a much larger collection, the amount of data is comparable to that of MSD or FMA because of the smaller dimensionality, as shown in \tabref{tab:size}.
+For statistical learning algorithms, the scale of FMA and MSD is thus comparable to ImageNet.
 	\item[Permissive licensing.] MIR research has historically suffered from the lack of publicly available benchmark datasets.
@@
 
-\noindent
-{\bf What would be the requirements for an ideal music dataset?}  \\
-(1) Availability of raw audio tracks like 30sec or more,\\
-(2) Availability of meta-data such as genres, artist, title, year, lyrics, users' ratings, users' comments, etc,\\
-(3) Large-scale dataset to be able to sample the high-dimensional distribution of music genres\\
-(4) Free and easy access of the dataset,\\
-(5) Legality of the dataset, no copyright issue with e.g. open access licence.\\
+% Do not present like this because we might want representativeness. Which would necessarily include commercial music.
 
-%%%%%%%%%%%%%%%%%%%
+%{\bf What would be the requirements for an ideal music dataset?} \\
+%(1) Availability of raw audio tracks like 30sec or more,\\
+%(2) Availability of meta-data such as genres, artist, title, year, lyrics, users' ratings, users' comments, etc,\\
+%(3) Large-scale dataset to be able to sample the high-dimensional distribution of music genres\\
+%(4) Free and easy access of the dataset,\\
+%(5) Legality of the dataset, no copyright issue with e.g. open access licence.\\
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{The Dataset} %Introducing the FMA dataset}
-%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 % Figures and tables from analysis.ipynb
+
+\subsubsection{Genres}
+
+While it is difficult to construct a hierarchy of genres, we follow that of our source.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Proposed tasks} % usage, problems
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % List some MIR tasks and propose some who can be done with that dataset
@@
 Can be down: genre, artist, popularity prediction, tags, year prediction
+% Cannot: mood, style, annotations
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Genre Recognition} % MGR, Baselines
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % One of the most common task, we give some baselines so that people have a reference about how hard it is.
+% ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
 
-Music genre recognition (MGR) is one of the most researched areas of MIR.
+% genre not defined by audio content only
+While the utility of music genre recognition (MGR) has been debated, mostly because of its ambiguity and cultural definition, it is widely used and understood by end-users who find it useful to discuss musical categories \cite{mgr_why}.
+Moreover, it is one of the most researched areas of MIR.
+
+While there certainly is noise in our labels, we trust the artists to have selected the most appropriate set of genres for their own songs.
 
@@
 
-%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion and Perspectives}
-%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Acknowledgments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliography{refs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2017-03-17 fma.tex
paper update

@@
 
-% Anonymity: authors, GitHub link
+% Anonymity: authors, GitHub links, acknowledgments
 
@@
 \usepackage{graphicx} \graphicspath{{figs/}}
-\usepackage{color}
 
-\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc} \usepackage[utf8]{inputenc}
 \usepackage[bookmarks=false,colorlinks=true,allcolors=blue]{hyperref}
-\usepackage{booktabs}
-%\usepackage[para,online,flushleft]{threeparttable}
-\usepackage{threeparttable}
-%\usepackage{footnote} \makesavenoteenv{table}
-\usepackage{rotating,tabularx}
-
-\usepackage{array}
+\usepackage{booktabs, threeparttable, tabularx}
+\usepackage{enumitem} \setlist{itemsep=0pt,topsep=0pt,parsep=0pt,partopsep=0pt}
 
@@
 For examples, developing deep learning techniques for music applications require the use of raw audio data, and designing recommender systems need meta-data like artist features and song ratings.
-Besides, the most influential competition in the field of music analysis organized every year is the Music Information Retrieval Evaluation eXchange (MIREX).\footnote{\url{http://www.music-ir.org/mirex/wiki/}} MIREX proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc.
+Besides, the most influential competition in the field of music analysis organized every year is the Music Information Retrieval Evaluation eXchange (MIREX).\href{http://www.music-ir.org/mirex/wiki/}{MIREX} proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc.
 However, participants do not have access to neither the test set, nor the important train set.
@@
 \noindent
-{\bf Available music datasets and their limitations.} In music analysis, three datasets have been widely used by the international society for music information retrieval (ISMIR)\footnote{\url{http://www.ismir.net/}}: the GTZAN, TagATune, and MSD (Million Song Dataset) datasets.
+{\bf Available music datasets and their limitations.} In music analysis, three datasets have been widely used by the international society for music information retrieval (\href{http://www.ismir.net/}{ISMIR}): the GTZAN, TagATune, and MSD (Million Song Dataset) datasets.
 However each dataset come with some limitations:
@@
 		\toprule
-		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \text{scale} & size [GB] & size [days] \\
+		Dataset & \#samples & \multicolumn{3}{c}{Dimensionality} & \multicolumn{3}{c}{Size} \\
+		\cmidrule{6-8}
+				&           & \multicolumn{3}{c}{}               & \text{scale} & [GB] & [days] \\
 		\midrule
@@
 All songs are licensed under Creative Commons by the artists and can thus be redistributed.
+The meta-data produced by ourselves are released under \todo{Creative Commons ??
+v4} and the code under an MIT license.
 * In contrast with images or text for which plenty of data is available online
@@
 As the original data and APIs are public, anybody can recreate the dataset.
-A train / test split is provided to make research on the FMA reproducible.
 	\item[Easy to work with]: working with the dataset only requires to download some archives.
-Moreover, we provide an helper Python package\footnote\url{https://pypi.python.org/pypi/freemusicarchive} and usage examples as Jupyter notebooks to quick start using the data.
+Moreover, we provide an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and some usage examples in the \texttt{usage.ipynb} Jupyter notebook to quick start using the data.
 \end{description}
@@
 That state of affair is due to copyright laws which forbid the redistribution of audio.
-The solution to that problem is to aim for tracks released under permissive licenses, such as Creative Commons\footnote{\url{https://creativecommons.org/}}.
+The solution to that problem is to aim for tracks released under permissive licenses, such as \href{https://creativecommons.org/}{Creative Commons}.
 
@@
 
+%\subsection{The Free Music Archive}
+
+The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive} (FMA), an interactive library of high-quality, legal audio downloads directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
+The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
+The website provides a large catalog of artists and high-quality mp3-encoded songs, hand-picked by established audio curators.
+Each song is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio.
+Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 songs.
+Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing.
+While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis \cite{art:MossFMA}.
+
+\subsection{Creation} % Collection, Creation, Generation
+% How it was collected.
+
+\begin{table}
+	\centering
+	\begin{tabular}{|ll|}
+		\hline
+		track\_id & track\_title \\
+		album\_title & artist\_name \\
+		\hline
+	\end{tabular}
+	\caption{List of per-song meta-data obtained from the FMA.}
+	\label{tab:metadata}
+\end{table}
+
+Song meta-data, such as the title, album, name of the artist, genres as well as some usage statistics, was crawled using the available \href{https://freemusicarchive.org/api}{API}.
+See \tabref{tab:metadata} for a list and \texttt{webapi.ipynb} for how to query the API with our helpers.
+The collected meta-information is stored in the \texttt{tracks.json} table, which \tabref{tab:tracks} shows an excerpt.
+Additionally, the used genre hierarchy, shown in \tabref{tab:genres}, was collected and stored in \texttt{genres.json}.
+Audio was then downloaded over HTTP.
+The collection of FMA was done in \todo{April 2016}, and the size of the dataset at this time was 89,912 songs.
+
+\todo{Top-level genres.} Then, we created a top genre by simply picking the first genre in the original list of track genres provided by the artists.
+
+%\subsection{Data Cleaning}
+
+\todo{Adjust when redone.}
+As a first filtering, we kept the data that has audio tracks, reducing the size to 86,886 music data. 
+We removed data without genre information, giving 79,947 song data.
+At this stage the number of genres is 138, but half have only a few samples.
+Thus we decided to only retain the genres with more than 100 samples.
+This defines a dataset of 77,643 music data with 68 genres.
+We further cleaned the dataset by removing tracks which belong to ``non-standard'' music genres like 'Noise', 'Garage', 'Sound Collage', 'Singer', 'Audio Collage', 'Glitch', 'Unclassifiable', 'Lo-Fi', 'Spoken', 'Poetry', 'Talk Radio', 'Avant-Garde', 'Experimental', 'Ambient', and 'Field Recording'.
+We also suppressed the songs with the title 'Untitled'.
+Besides, we decided to cut the long tail of the power law distribution by removing songs of genres with less than 300 songs.
+
+The FMA\_large dataset of 77,643 data contains songs of 30sec length, and artist name, track title, track genres, track listens as meta-data.
+Our original motivation was to give the full audio tracks, not limited to 30sec, but the size of the whole music tracks is 806.8GB (for approximatively 256 days of music listening) making the hosting a challenging issue.
+
+See \texttt{generation.ipynb} for full details.
+
+\begin{table*}
+	\tiny
+	\centering
+	\begin{tabular}{lrlrllll}
+	\toprule
+	 fma\_id &                                             artist &  play\_count &                                              title &                                             genres &            top\_genre  \\
+	 \toprule
+		10 &                                          Kurt Vile &       42936 &                                            Freeway &                                              [Pop] &                  Pop  \\
+		134 &                                               AWOL &         880 &                                       Street Music &                                          [Hip-Hop] &              Hip-Hop  \\
+		 142 &                    Alec K.
+Redfearn \& the Eyesores &         670 &                               Punjabi Watery Grave &                                             [Folk] &                 Folk  \\
+		 236 &                                       Banana Clipz &        6695 &                              Push Am (Left, Right) &                              [Electronic, African] &           Electronic  \\
+		 237 &                                          Barnacled &        1056 &                    Garbage and Fire &                                             [Jazz] &                 Jazz  \\
+		 461 &                              Cantonement Jazz Band &        3933 &                                           Bessemer &                               [Blues, Jazz: Vocal] &                Blues  \\
+		824 &                     Here Comes A Big Black Cloud!!
+&         391 &                                         Black Mold &          [Rock, Loud-Rock, Psych-Rock, Indie-Rock] &                 Rock  \\
+		 825 &                     Here Comes A Big Black Cloud!!
+&        1847 &                                        Death March &          [Rock, Loud-Rock, Psych-Rock, Indie-Rock] &                 Rock  \\
+		837 &                                          Heroin UK &         146 &                                           DopeSick &                                             [Rock] &                 Rock \\
+		 889 &                                 Illusion of Safety &         163 &                                          Wasteland &                                       [Electronic] &           Electronic  \\
+		 896 &                                        Impediments &        1542 &                                               2012 &                                  [Punk, Power-Pop] &                 Punk  \\
+			  992 &                                      Jason Willett &         560 &                   Beautiful Song w/ kick drum solo &                                             [Rock] &                 Rock  \\
+	\bottomrule
+	\end{tabular}
+	\caption{Some rows and columns of the \texttt{tracks.json} table. \todo{representative and varied}}
+	\label{tab:tracks}
+\end{table*}
+
+\subsection{Content} \label{sec:content}
+% Figures and tables from analysis.ipynb
 % TODO audio properties: samplerate (% 44kHz), bitrate (% 128, VBR), channels (% mono, stereo)
-Most tracks are released under a Creative Commons license\footnote{\url{https://creativecommons.org/}} which allows redistribution of the audio.
 
-\subsection{Analysis}
-% Figures and tables from analysis.ipynb
+All tracks come with the meta-data listed in \tabref{tab:metadata}, although some values may be missing.
+All audio tracks are mp3-encoded with sampling rate of 44,100Hz, bit rate 128kb/s, and in stereo.
 
-\subsubsection{Genres}
+%\subsubsection{Genres}
 
 While it is difficult to construct a hierarchy of genres, we follow that of our source.
+\tabref{tab:genres} and \figref{fig:genres} show the distribution of the \todo{??} most common genres.
+As often encountered with natural categories, it obeys a power law.
+See \texttt{analysis.ipynb} for more details.
+
+\begin{table}
+	\centering
+	\begin{tabular}{cclc}
+		\toprule
+		ID & Parent & Genre & \#songs \\
+		\midrule
+		4 & None & Electronic & 50,000 \\
+		4 & None & Electronic & 50,000 \\
+		4 & None & Electronic & 50,000 \\
+		\midrule
+		45 & 2 & Psych-Rock & 500 \\
+		45 & 2 & Psych-Rock & 500 \\
+		45 & 2 & Psych-Rock & 500 \\
+		\bottomrule
+	\end{tabular}
+	\caption{Number of songs per genre. Top part is the 16 top-level genres, bottom part is the second- and third-level genres.}
+	\label{tab:genres}
+\end{table}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{histo_large.pdf}
+	\caption{Power law distribution of the top 20 genres.}
+	\label{fig:genres}
+\end{figure}
 
@@
 
-Extracted with librosa\footnote{\url{https://github.com/librosa/librosa}}.
+We extracted a wide range of audio features from the audio, using the \href{https://github.com/librosa/librosa}{librosa} Python library. An overview
+on these features is given in \tabref{tab:features}.
+In addition, a number of descriptive medium-level and high-level audio features extracted by \href{http://the.echonest.com/}{The Echo Nest}, a music platform for developers and media companies, are provided for some tracks.
+These features include tempo, loudness, timings of fade-in and fade-out, and MFCC-like features for a number of segments.
+Those features can be obtained either by uploading the songs to their website, which are then analyzed and sent back to the user, or simply retrieved from their database if the songs are known to them, i.e.
+they were analyzed before.
+All of which can be done through an \href{http://developer.echonest.com}{API}.
+In our case, we only retrieved features from songs which were already in their database.
+See \texttt{features.ipynb} for more details.
 
-In addition, a number of descriptive features extracted with the services from The Echo Nest are provided.
-These features include tempo, loudness, timings of fade-in and fade-out, and MFCC-like features for a number of segments.
+\begin{table}
+	\centering
+	\begin{tabular}{llr}
+		\toprule
+		Feature set & Extractor & Dimensionality \\
+		\midrule
+		MFCCs \cite{mfcc} & librosa & 26 \\
+		MFCCs \cite{mfcc} & librosa & 26 \\
+		MFCCs \cite{mfcc} & librosa & 26 \\
+		\bottomrule
+	\end{tabular}
+	\caption{Overview of features extracted from the FMA samples.}
+	\label{tab:features}
+\end{table}
 
-\subsection{Collection} % Creation, Generation
-% How it was collected.
+\subsection{Splits}
 
-\noindent
-{\bf The website.} The FMA website\footnote{\url{https://freemusicarchive.org/}} is run by WFMU, the longest-running free-form radio station in the United States.
-The website provides a large catalogue of artists and high-quality mp3 songs.
-Each song is legally free to download as artists decided to work under open licence such as Creative Commons license.
-This is directly inspired by the open source movement, and the goal of the FMA site is to offer a legal and technological platform for artists and listeners to harness the full potential of online music sharing \cite{art:MossFMA}.
-FMA has been funded by the New York State Music Fund, the MacArthur Foundation, and the National Endowment for the Arts.\\
+A train / test split is proposed to make research on the FMA reproducible.
+Below are the constraints we followed:
+\begin{enumerate}
+	\item For each genre, we randomly split train and test songs with the same ratio 80-20\%.
+	\item We forced an artist to be either in the train set, or the test set (meaning the artist's songs cannot be in both train and test sets).
+	\item We filled alternatively the train set and test set in the descending order of the count of artist songs. This forces the two sets to have the same distribution of songs per artist.
+\end{enumerate}
+A track is assigned in train or split for all the subsets and the above constraints are satisfied for all sets. See \texttt{generation.ipynb} for full details.
 
-\noindent
-{\bf Crawling the data.} The FMA organization provides an API\footnote{\url{https://freemusicarchive.org/api}} to collect the meta-data associated to each song such as artist name, track title, track genre, and track listens.
-The collect of the mp3 audio files was done with a web scrapper.
-The collection of FMA was done in April 2016, and the size of the dataset at this time was 89,912 songs.
-It was recently announced on July 2016 that the website had reached the landmark of 100,000 songs.
-\\
+\subsection{Subsets} \label{sec:subsets}
 
-\noindent
-{\bf The raw FMA dataset.} The dataset collected on April 2016 contains 89,912 music data.
-As a first filtering, we kept the data that has audio tracks, reducing the size to 86,886 music data.
-All audio tracks are mp3-encoded with sampling rate of 44,100Hz, bit rate 128kb/s, and in stereo.
-Then, we considered the meta-data provided by the FMA website and crawled with its API.
-The meta-information collected is artist name, track title, track genres, and track listens.
-We removed data without genre information, giving 79,947 song data.
-Then, we created a top genre by simply picking the first genre in the original list of track genres provided by the artists.
-A few examples of data are given in Table \ref{tab1}.
-At this stage the number of genres is 138, but half have only a few samples.
-Thus we decided to only retain the genres with more than 100 samples.
-This defines a dataset of 77,643 music data with 68 genres.
-Figure \ref{fig_large} presents the distribution of the top 20 genres, which obeys to a power law distribution as often encountered in natural data collection.
-We further cleaned the dataset by removing tracks which belong to ``non-standard'' music genres like 'Noise', 'Garage', 'Sound Collage', 'Singer', 'Audio Collage', 'Glitch', 'Unclassifiable', 'Lo-Fi', 'Spoken', 'Poetry', 'Talk Radio', 'Avant-Garde', 'Experimental', 'Ambient', and 'Field Recording'.
-We also suppressed the songs with the title 'Untitled'.
-Besides, we decided to cut the long tail of the power law distribution by removing songs of genres with less than 300 songs.
-Finally, we kept the songs that have the raw audio and all Echonest features.
-Echonest features are medium-level and high-level audio features provided by Echonest\footnote{\url{http://the.echonest.com/}}, a music platform for developers and media companies which has a collection of 3.5M artists and 37.5M songs.
-Echonest features can be either computed by uploading the songs to the Echonest website, which are then analyzed and send them back to the user, or simply retrieved from the Echonest database if the songs were analyzed before.
-All Echonest operations are done through an API\footnote{\url{http://developer.echonest.com}}.
-Our final filtering consists in keeping songs that have both raw audio file and complete Echonest features already computed and stored in the Echonest database.
-The final size of the dataset is 14,511 music data and 20 genres.
-Figure \ref{fig_medium} presents the distribution of the genre sizes for this dataset.
-\\
+\begin{table}
+	\centering
+	\begin{tabular}{lrrccc}
+		\toprule
+		dataset & clips & genres & length & \multicolumn{2}{c}{Size} \\
+		\cmidrule{5-6}
+		        &         &          &  [s]   & [GB] & [days] \\
+		\midrule
+		small  &   4,000 & 10  & 30    & 3.4  & 1.4  \\
+		medium &  14,511 & 20  & 30    & 12.2 & 5.1  \\
+		large  & 110,000 & 164 & 30    & 90   & 26.9 \\
+		huge   & 110,000 & 164 & <1800 & 900  & 229  \\
+		\bottomrule
+	\end{tabular}
+	\caption{Proposed subsets of the FMA.}
+	\label{tab:subsets}
+\end{table}
 
-\noindent
-{\bf Proposed benchmark datasets.} For research purposes, we propose 3 datasets:\\
-(1) FMA\_small: This dataset contains 4,000 songs with 10 genres equally balanced, that is 400 songs per class.
-For each song, we extracted 30sec precisely in the middle of the song.
-The total length is 33.3 hours of audio listening and the size is 3.4GB.
-The genre distribution is given in Figure \ref{fig_small}.
-The motivation for this dataset is to create an alternative to the popular GTZAN dataset with respect to the size and genre parameters.
-But unlike GTZAN, all songs have not only raw audio tracks but also meta-data composed of artist name, track title, track genres, track listens, and Echonest features.
-The GTZAN data may not have for example song title or artist name, and also no copyright permission.
-For reproducibility research, we also provide a train-test split 80-20 of this dataset for future techniques' comparisons.
-\\
-(2) FMA\_medium: This dataset has 20 genres with 14,511 songs of 30sec, also extracted from the middle of the track.
-The total length is 5.1 days of listening and the size is 12.2GB.
-Unlike FMA\_small, the size distribution of genres is not balanced, see Figure \ref{fig_medium}, but all songs have raw audio files and all meta-data i.e.
-artist name, track title, track genres, track listens, and Echonest features.
-We also provide a split 80-20 of this dataset.
-\\
-(3) FMA\_large: This dataset has 77,643 data of 30sec, again taken from the track middle, and 68 genres.
-It provides 26.9 days of track listening for 79.8GB.
-Figure \ref{fig_large} presents the top classes in the genre distribution.
-All songs come with the following meta-data: artist name, track title, track genres, track listens.
-Eventually, we give a split 80-20 of this dataset.\\
-
-\noindent
-Regarding the splitting train-test sets, we considered these constraints:\\
-(a) For each genre, we randomly split train and test songs with the same ratio 80-20.\\
-(b) We forced an artist to be either in the train set, or the test set (meaning the artist's songs cannot be in both train and test sets). \\
-(c) We filled alternatively the train set and test set in the descending order of the count of artist songs. This prevented the test set to be filled with artists who have only a few songs.\\
-
-\noindent
-The FMA\_large dataset of 77,643 data contains songs of 30sec length, and artist name, track title, track genres, track listens as meta-data.
-Our original motivation was to give the full audio tracks, not limited to 30sec, but the size of the whole music tracks is 806.8GB (for approximatively 256 days of music listening) making the hosting a challenging issue.
+Finally, for the dataset to be useful as a development set or for people with lower computational resources, we propose the following subsets, each of which is a subset or the larger set:
+\begin{enumerate}
+	\item \textbf{Large}: this is the full dataset except audio is 30 seconds clips extracted from the middle of the tracks.
+	\item \textbf{Medium}: all tracks which have Echonest features.
+	\item \textbf{Small}: tracks from 10 top-level genres, balanced with 400 tracks per genre, 1 genre per track.
+This subset is similar to the very popular GTZAN in terms of size and genre distribution with the added benefits of the FMA, that is meta-data, pre-computed features and copyright-free audio.
+\end{enumerate}
+\tabref{tab:subsets} highlights the main differentiating factors between the proposed subsets.
+See \texttt{analysis.ipynb} for an analysis of the subsets' content similar to that shown in \secref{sec:content}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Proposed tasks} % usage, problems
+\section{Proposed usage} % tasks, usage, problems
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % List some MIR tasks and propose some who can be done with that dataset
+
+% MSD: metadata analysis, artist recognition, automatic music tagging, recommendation, cover song recognition, lyrics, limitations
 
@@
 
+% MSD: Year Prediction
+
 % genre not defined by audio content only
@@
 
+\subsection{Data}
+
 While there certainly is noise in our labels, we trust the artists to have selected the most appropriate set of genres for their own songs.
+
+\subsection{Methods}
+
+\subsection{Evaluation and results}
+
+The performance of various classifiers on various features are shown in \tabref{tab:mgr} as baselines for future work.
+See \texttt{baselines.ipynb} for more details.
+
+\begin{table}
+	\centering
+	\begin{tabular}{lccccc}
+		\toprule
+		Features & NB & SVM & kNN & DT & RF \\
+		\midrule
+		MFCCs  & 30.12 & 30.12 & 30.12 & 30.12 \\
+		MFCCs  & 30.12 & 30.12 & 30.12 & 30.12 \\
+		MFCCs  & 30.12 & 30.12 & 30.12 & 30.12 \\
+		Echonest  & 30.12 & 30.12 & 30.12 & 30.12 \\
+		\bottomrule
+	\end{tabular}
+	\caption{Genre recognition results for various features and classifiers.}
+	\label{tab:mgr}
+\end{table}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% MSD: the future of the dataset, visibility for MIR (to DL people!)
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliography{refs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[h!]
-\centering
-\includegraphics[height=6cm]{histo_large.pdf}
-\vspace{-0.5cm}
-\caption{Top 20 music genres for the FMA\_large dataset of 77,643 songs.}
-\label{fig_large}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[height=6cm]{histo_medium.pdf}
-\vspace{-0.5cm}
-\caption{The 20 music genres for the FMA\_medium dataset with 14,511 songs.}
-\label{fig_medium}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[height=5cm]{histo_small.pdf}
-\vspace{-0.5cm}
-\caption{The 10 music genres for the FMA\_small dataset with 4,000 songs.}
-\label{fig_small}
-\end{figure}
-
-%\begin{table}[h!]
-\begin{sidewaystable}
-\tiny{
-\centering
-\begin{tabular}{lrlrllll}
-\toprule
- fma\_id &                                             artist &  play\_count &                                              title &                                             genres &            top\_genre  \\
- \toprule
-    10 &                                          Kurt Vile &       42936 &                                            Freeway &                                              [Pop] &                  Pop  \\
-    134 &                                               AWOL &         880 &                                       Street Music &                                          [Hip-Hop] &              Hip-Hop  \\
-     141 &                    Alec K.
-Redfearn \& the Eyesores &         590 &                                               Ohio &                                             [Folk] &                 Folk  \\
-     142 &                    Alec K.
-Redfearn \& the Eyesores &         670 &                               Punjabi Watery Grave &                                             [Folk] &                 Folk  \\
-    144 &                                   Amoebic Ensemble &         901 &                                            Wire Up &                                             [Jazz] &                 Jazz  \\
-   145 &                                   Amoebic Ensemble &         682 &                                         Amoebiasis &                                             [Jazz] &                 Jazz  \\
-    188 &                                           Ed Askew &         973 &                                            Piano 1 &                                             [Folk] &                 Folk  \\
-     206 &                                           Ed Askew &          73 &                                        Bella Crane &                                             [Folk] &                 Folk \\
-     236 &                                       Banana Clipz &        6695 &                              Push Am (Left, Right) &                              [Electronic, African] &           Electronic  \\
-     237 &                                          Barnacled &        1056 &                    Garbage and Fire &                                             [Jazz] &                 Jazz  \\
-     238 &                                          Barnacled &         961 &                                     France Attacks &                                             [Jazz] &                 Jazz  \\
-     461 &                              Cantonement Jazz Band &        3933 &                                           Bessemer &                               [Blues, Jazz: Vocal] &                Blues  \\
-     462 &                              Cantonement Jazz Band &        3256 &                                     Has Been Blues &                               [Blues, Jazz: Vocal] &                Blues  \\
-     463 &                              Cantonement Jazz Band &        3238 &                                       I'll Be Blue &                               [Blues, Jazz: Vocal] &                Blues  \\
-    824 &                     Here Comes A Big Black Cloud!!
-&         391 &                                         Black Mold &          [Rock, Loud-Rock, Psych-Rock, Indie-Rock] &                 Rock  \\
-     825 &                     Here Comes A Big Black Cloud!!
-&        1847 &                                        Death March &          [Rock, Loud-Rock, Psych-Rock, Indie-Rock] &                 Rock  \\
-    837 &                                          Heroin UK &         146 &                                           DopeSick &                                             [Rock] &                 Rock \\
-     889 &                                 Illusion of Safety &         163 &                                          Wasteland &                                       [Electronic] &           Electronic  \\
-     896 &                                        Impediments &        1542 &                                               2012 &                                  [Punk, Power-Pop] &                 Punk  \\
-          992 &                                      Jason Willett &         560 &                   Beautiful Song w/ kick drum solo &                                             [Rock] &                 Rock  \\
-\bottomrule
-\end{tabular}
-\caption{A few song data in the FMA dataset.}
-\label{tab1}
-%\end{table}
-}
-\end{sidewaystable}
-
 \end{document}

### 2017-03-17 fma.tex
update

@@
 Our differentiating factors:
-\begin{description}
-	\item[Large scale.] Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with genre.
+
+%\begin{itemize}
+\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with genre.
 While FMA is not the largest in term of number of clips, every other dataset with readily available quality audio are two orders of magnitude smaller.
@@
 For statistical learning algorithms, the scale of FMA and MSD is thus comparable to ImageNet.
-	\item[Permissive licensing.] MIR research has historically suffered from the lack of publicly available benchmark datasets.
+
+\textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets.
 Most of these issues stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
-All songs are licensed under Creative Commons by the artists and can thus be redistributed.
-The meta-data produced by ourselves are released under \todo{Creative Commons ??
-v4} and the code under an MIT license.
+All songs, i.e.
+all \texttt{.mp3} in the \texttt{.zip} archives, are licensed under Creative Commons by the artists which allows redistribution.
+The meta-data produced by ourselves, i.e.
+all the \texttt{.json} files, is released under the terms of the \href{https://creativecommons.org/licenses/by/4.0)}{Creative Commons Attribution 4.0 International License (CC BY 4.0)} and the code, i.e.
+all the files in the git repository, under the MIT license.
+
 * In contrast with images or text for which plenty of data is available online
-	\item[Audio.] most of the largest datasets are distributed with features, sometimes computed by commercial services like the Echonest.
+
+\textbf{Audio.} most of the largest datasets are distributed with features, sometimes computed by commercial services like the Echonest.
 Researchers are wary of those proprietary features and have seek to compute features themselves, e.g.
@@
 In contrast, FMA provides audio for each song.
-	\item[High quality audio.] full-length original audio uploaded by the artist.
+
+\textbf{High quality audio.} full-length original audio uploaded by the artist.
 Stereo mp3 with bitrates varying between x.
@@
 To our knowledge, none of the released datasets up to this date come with full-length audio.
-	\item[Future proof]: all audio and meta-data can be downloaded as ZIP archives, no need to scrap the web yourself.
+
+\textbf{Future proof}: all audio and meta-data can be downloaded as ZIP archives, no need to scrap the web yourself.
 They are checksummed, have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by CERN.
 We thus alleviate the risks of tracks to become unavailable, to change without notice or for online services to shut down.
+
 % Western / commercial music, how to qualify broadness?
-	\item[Reproducible]: we share all the code used to (i) collect the dataset, (ii) analyze it, (iii) extract the sub-sets, (iv) compute the features and (v) the baselines, so that it can be reproduced or extended.
+\textbf{Reproducible}: we share all the code used to (i) collect the dataset, (ii) analyze it, (iii) extract the sub-sets, (iv) compute the features and (v) the baselines, so that it can be reproduced or extended.
 As the original data and APIs are public, anybody can recreate the dataset.
-	\item[Easy to work with]: working with the dataset only requires to download some archives.
+
+\textbf{Easy to work with}: working with the dataset only requires to download some archives.
 Moreover, we provide an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and some usage examples in the \texttt{usage.ipynb} Jupyter notebook to quick start using the data.
-\end{description}
+%\end{itemize}
 
@@
 
-See \texttt{generation.ipynb} for full details.
+See \texttt{creation.ipynb} for full details.
 
@@
 	\end{tabular}
-	\caption{Some rows and columns of the \texttt{tracks.json} table. \todo{representative and varied}}
+	\caption{Some rows and columns of the table stored in \texttt{tracks.json}. \todo{representative and varied}}
 	\label{tab:tracks}
@@
 	\end{tabular}
-	\caption{Number of songs per genre. Top part is the 16 top-level genres, bottom part is the second- and third-level genres.}
+	\caption{Number of songs per genre. Top part is the 16 top-level genres, bottom part is the second- and third-level genres. The genre hierarchy is stored in \texttt{genres.json}.}
 	\label{tab:genres}
@@
 \end{enumerate}
-A track is assigned in train or split for all the subsets and the above constraints are satisfied for all sets. See \texttt{generation.ipynb} for full details.
+A track is assigned in train or split for all the subsets and the above constraints are satisfied for all sets. See \texttt{creation.ipynb} for full details.
 
@@
 		large  & 110,000 & 164 & 30    & 90   & 26.9 \\
-		huge   & 110,000 & 164 & <1800 & 900  & 229  \\
+		full   & 110,000 & 164 & <1800 & 900  & 229  \\
 		\bottomrule
@@
 \begin{enumerate}
+	\item \textbf{Full}: this is the complete dataset as described in \secref{sec:content}.
 	\item \textbf{Large}: this is the full dataset except audio is 30 seconds clips extracted from the middle of the tracks.
@@
 We are grateful to SWITCH and EPFL for hosting the dataset within the context
-of the \href{https://projects.switch.ch/scale-up/}{SCALE-UP} project, funded in
+of the \href{https://projects.switch.ch/scale-up}{SCALE-UP} project, funded in
 part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC
@@
 
+We also want to thank the team behind \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
+
 \bibliography{refs}

### 2017-03-21 fma.tex
update

@@
 \documentclass{article}
-\usepackage{ismir,amsmath,cite,url}
+\usepackage[T1]{fontenc} \usepackage[utf8]{inputenc}
+\usepackage[svgnames]{xcolor}
+\usepackage[bookmarks=false,colorlinks=true,allcolors=DarkBlue]{hyperref} % MidnightBlue, NavyBlue, DarkBlue, Maroon
+\usepackage{ismir,amsmath}
 \usepackage{graphicx} \graphicspath{{figs/}}
-
-\usepackage[T1]{fontenc} \usepackage[utf8]{inputenc}
-\usepackage[bookmarks=false,colorlinks=true,allcolors=blue]{hyperref}
 \usepackage{booktabs, threeparttable, tabularx}
@@
 
-\usepackage{color}
 \newcommand{\todo}[1]{{\color{red} #1 }}
@@
 \begin{abstract}
-We present a new music dataset that can be used for several music analysis tasks.
-Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of meta-data for artists analysis or song ratings for recommender systems.
+
+The Free Music Archive (FMA) is an open music dataset that can be used to evaluate several tasks in music information retrieval (MIR).
+The released data is composed of 110,000 tracks, xx artists, xx albums, 164 genres, for a total of 229 days of audio for 900 GiB, all under permissive Creative Commons licenses.
+It features a collection of meta-data such as the song title, artist name, album or genres, together with user data such as play counts, full-length and high-quality audio, as well as a set of pre-computed features.
+We propose a train / test split and three subsets: a small genre-balanced dataset of 4,000 songs and 10 genres, a medium genre-unbalanced dataset of 14,511 songs and 20 genres, and a 90 GiB version of the full set, with clips trimmed to 30s.
+This paper describes the dataset and how it was created, proposes some tasks and evaluate some baselines for music genre recognition (MGR).
+All the code to reproduce it as well as links to download the data and usage examples are available at \url{https://github.com/mdeff/fma}.
+
+% Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of meta-data for artists analysis or song ratings for recommender systems.
 Existing datasets such as GTZAN, TagATune, and Million Song suffer from the previous limitations.
 It is however essential to establish such benchmark datasets to advance the field of music analysis, like the ImageNet dataset which made possible the large success of deep learning techniques in computer vision.
-In this paper, we introduce the Free Music Archive (FMA) which contains 77,643 songs and 68 genres spanning 26.9 days of song listening and meta-data including artist name, song title, music genre, and track counts.
-For research purposes, we define two additional datasets from the original one: a small genre-balanced dataset of 4,000 song data and 10 genres compassing 33.3 hours of raw audio and a medium genre-unbalanced dataset of 14,511 data and 20 genres offering 5.1 days of track listening, both datasets come with meta-data and Echonest audio features.
-For all datasets, we provide a train-test splitting for future algorithms' comparisons.
+
+% which contains 77,643 songs and 68 genres spanning 26.9 days of song listening
 \end{abstract}
-
-The released dataset is composed of 100,000 tracks, xx artists, xx albums, xx hours of audio (xx GiB), 164 genres, all under permissive Creative Commons licenses.
-
-The dataset comes with a collection of meta-data such as the song titles, artists and albums, together with full-length audio and a set of pre-computed features as well as a train/test split.
-
-Content-based music information retrieval tasks are typically solved with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
-Although that approach was dominant in the past, feature learning and end-to-end learning (from signal to label) have started to receive more attention from the MIR community in recent years. 
-The primary goal of this dataset is to enable researchers to train large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
-Hopefully, the release of such amount of easily accessible data will foster research which will ultimately surpass the performance ceiling observed in e.g.
-genre recognition, as it as been the case in computer vision with the release of ImageNet.
-
-This paper describes the dataset and how it was created, proposes some tasks and evaluate some baselines for music genre recognition (MGR).
-All the code to reproduce it as well as links to download the data and usage examples are available at \url{https://github.com/mdeff/fma}.
 
@@
 
-\noindent
-{\bf Music Analysis Field.} Developing new mathematical models and algorithms to solve challenging real-world problems is obviously of first importance in any field of research.
-But such novel techniques must be evaluated and compared to the existing state-of-the-art techniques to be adopted as new standards by research communities.
-Evaluation and comparison require benchmark datasets to achieve their purpose, which are sometimes challenging to collect and share.
-In computer vision, the community has developed over the years well-established benchmark datasets such as \href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist}, \href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR} \cite{cifar}, or \href{http://www.image-net.org}{ImageNet} \cite{imagenet}.
-All these image datasets are free, legal and easily available on Internet.
-Such datasets have proved essential to advance the field of computer vision.
-The most celebrated example is the ImageNet dataset and challenge in 2012.
-This unprecedented dataset of 1.5M image data allowed to demonstrate the power of deep learning techniques, which was able to win the competition by a large margin over the second best.
-In music analysis, such benchmark datasets are not easily available and they also lack of essential information.
-For examples, developing deep learning techniques for music applications require the use of raw audio data, and designing recommender systems need meta-data like artist features and song ratings.
-Besides, the most influential competition in the field of music analysis organized every year is the Music Information Retrieval Evaluation eXchange (MIREX).\href{http://www.music-ir.org/mirex/wiki/}{MIREX} proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc.
-However, participants do not have access to neither the test set, nor the important train set.
-They must upload their code to a website that will be evaluated by the organizers.
-In other words, participants cannot train their analysis models on any part of the dataset.
-These existing shortcomings make essential the development of new music datasets to advance the field of music analysis.
-\\
+% hook
+While the development of new mathematical models and algorithms to solve challenging real-world problems is obviously of first importance to any field of research, evaluation and comparison to the existing state-of-the-art is necessary for a technique to be widely adopted by research communities.
+Such tasks require benchmark datasets.
+% to achieve their purpose, which are sometimes challenging to collect and share.
+In computer vision, the community has developed over the years established benchmark datasets such as \href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist}, \href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR} \cite{cifar}, or \href{http://www.image-net.org}{ImageNet} \cite{imagenet}.
+These datasets are free, legal, easily available, and have proved essential to advance the field.
+The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.5M images, allowed to demonstrate the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best \cite{convnet_imagenet}.
 
-In contrast with other modalities such as images or text for which a wealth of content is available, the lack of a large-scale and easily available dataset for MIR has hindered research on data-heavy models such as Deep Learning (DL).
-
-\noindent
-{\bf Available music datasets and their limitations.} In music analysis, three datasets have been widely used by the international society for music information retrieval (\href{http://www.ismir.net/}{ISMIR}): the GTZAN, TagATune, and MSD (Million Song Dataset) datasets.
-However each dataset come with some limitations:
-We discuss below some of the most related datasets, see \tabref{tab:datasets} for a more comprehensive list.
-\tabref{tab:datasets} shows the most common datasets used for audio-based / content-based MIR.
-\begin{description}
-	\item[GTZAN] \cite{gtzan} is a collection of 1000 songs with 10 music genres.
-Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
-It is the first benchmark MGR dataset to have been made publicly available, and as a result continues to be the most used public dataset for MGR \cite{mgr_eval}.
-A complete study, and critic, of this dataset was done in \cite{gtzan_critic_1}, and the dataset is available online.
-The main limitations of GTZAN is the legality of the dataset, the small size, no complete meta-data regarding artist names and song titles, and no additional meta-data like ratings.
-	\item[MagnaTagATune] \cite{magnatagatune} is a popular dataset of 5,405 source songs from 230 artists, cut in 25,863 clips.
-The music was collected from the \href{https://magnatune.com/}{Magnatune} label and was tagged using the \href{http://tagatune.org/}{TagATune} game.
-The dataset includes meta-data, audio features and raw audio, although the 16 kHz, 32kbps, mono mp3 are of poor audio quality.
-	\item[The Million Song Dataset (MSD)] \cite{msd} is a free and legal collection of 1 million music songs.
-Each song data is composed of high-level and medium-level audio features provided by  Echonest service and meta-data like artist name, song title, track genre, etc.
-It has its own dedicated website with related code and additional datasets.
-The very large scale of this music dataset would make it an ideal candidate for deep learning, which works best with large datasets.
-However, deep learning techniques work directly on the raw data (audio tracks in this case, like raw image pixels in computer vision) and not on pre-processed audio features given by Echonest.
-This prevents the application of deep learning on this dataset.
-Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
-		While researchers have been able to download sample audio from the online service \href{https://www.7digital.com}{7digital.com} and extract features by themselves, the process is tedious \cite{msd_features}.
-	\item[AudioSet] \cite{audioset} is a dataset of 2.1 million annotated 10-second sound clips associated with 527 classes.
-The ontology is quite broad and covers sounds from music to cars, engines or animals.
-The raw audio is however not included and has to be downloaded from Youtube.
-It is not clear of what happens if a referenced video gets deleted.
-\end{description}
-
-\begin{table}
+\begin{table}[t]
 	\centering
@@
 		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
-		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{garageband} & 1,886 & - & 2005 & yes \\  % GarageBand
-		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} & 3,115 & - & 2010 & no \\
-		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} & 3,180 & 1,517 & 2010 & yes \\
+		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
+		103-Artists \cite{103artists} & 2,445 & 103 & 2005 & yes \\
+		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} \cite{unique} & 3,115 & 3,115 & 2010 & yes \\
+		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} \cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
 		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
@@
 		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
-		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf 110,000 & \bf - & \bf 2017 & \bf yes \\
+		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf 110,000 & \bf \todo{?} & \bf 2017 & \bf yes \\
 		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
 		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{1} \\
-		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,100,000 & - & 2017 & no\tnote{1} \\
+		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{1} \\
 		\bottomrule
@@
 	\begin{tablenotes}
-		\item[1] Audio can be downloaded from a web service. % \url{www.ballroomdancers.com} \url{7digital.com} \url{youtube.com}
-		\item[2] The 25,863 clips come from 5,405 songs.
+		% 103-Artists is also known as DB-L, Extended Ballroom
+		\item[1] Audio can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
+		\item[2] The 25,863 clips are cut from 5,405 songs.
 		\item[3] Low quality 16 kHz, 32kbps, mono mp3.
@@
 	\end{threeparttable}
-	\caption{Comparison with some other datasets.}
+	\caption{Comparison with some other datasets.} % Genre classification datasets.
 	\label{tab:datasets}
 \end{table}
-
 % Lists of datasets:
@@
 % * A Survey of Evaluation in Music Genre Recognition
-
-% TODO licensing
 
@@
 \end{table*}
+% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
+% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
 
-% table quality
+% Besides, the most influential competition in the field of music analysis organized every year is the Music Information Retrieval Evaluation eXchange (MIREX).\href{http://www.music-ir.org/mirex/wiki/}{MIREX} proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc.
+However, participants do not have access to neither the test set, nor the important train set.
+They must upload their code to a website that will be evaluated by the organizers.
+In other words, participants cannot train their analysis models on any part of the dataset.
 
-Our differentiating factors:
+% {\bf Available music datasets and their limitations.}
+Unlike other modalities such as images or text for which a wealth of content is available, the lack of a large-scale, complete and easily available dataset for MIR has hindered research on data-heavy models such as DL.
+\tabref{tab:datasets} lists the most common datasets used for content-based MIR.
+GTZAN \cite{gtzan}, a collection of 1000 songs from 10 genres, was the first publicly available benchmark MGR dataset.
+As a result, despites its flaws \cite{gtzan_critic_1}, it continues to be the most used dataset for MGR \cite{mgr_eval}.
+The main limitations of GTZAN is its illegality, small size, and missing meta-data (no user ratings or even artist name and song title).
+% Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
+Looking at \tabref{tab:datasets}, the well-known MagnaTagATune and the Million Song Dataset (MSD) as well as the newer AudioSet appear as contenders for a large-scale reference dataset.
+% They however each have limitations.
+MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes meta-data, audio features and raw audio.
+The poor audio quality and limited number of songs does however limit its usage.
+MSD and AudioSet, while legal and very large, force researchers to download audio excerpts from online services.
+% Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
+% Each song data is composed of high-level and medium-level audio features provided by Echonest service and meta-data like artist name, song title, track genre, etc.
+It has its own dedicated website with related code and additional datasets.
+% 2.1 million annotated 10-second sound clips associated with 527 classes. The ontology is quite broad and covers sounds from music to cars, engines or animals. 
+On the other hand, the proposed dataset offers the following qualities, which in our view are essential for a reference benchmark.
 
-%\begin{itemize}
+% Value proposition.
 \textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with genre.
+Large collections are also more diverse and allows to average out annotation noise.
 While FMA is not the largest in term of number of clips, every other dataset with readily available quality audio are two orders of magnitude smaller.
@@
 Most of these issues stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
+The only solution to that problem is to aim for tracks released under permissive licenses.
 All songs, i.e.
-all \texttt{.mp3} in the \texttt{.zip} archives, are licensed under Creative Commons by the artists which allows redistribution.
+all \texttt{.mp3} in the \texttt{.zip} archives, are licensed under Creative Commons by the artists, which allows redistribution.
 The meta-data produced by ourselves, i.e.
@@
 all the files in the git repository, under the MIT license.
+% TODO licensing
+% Creative Commons: MagnaTagATune
+% Proprietary, non-redistributable: MSD, AudioSet?
 
-* In contrast with images or text for which plenty of data is available online
+\textbf{High-quality audio.}
+\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the largest datasets don't.
+Either they only contain features derived from the audio, either they provide links for researchers to download the audio from an online service.
+The problem with the former is that researchers are stuck with the set of chosen features and prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later makes no assurance that the files or services won't disappear or change without notice.
+Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{The Echo Nest} and have tried to extract features by themselves by downloading sample audio for the MSD from \href{https://www.7digital.com}{7digital.com}, which is tedious \cite{msd_features}.
+Finally, released or available to download audio is usually clips of 10 to 30 seconds, sometimes of low quality.
+FMA is distributed with legally redistributable full-length and high-quality original audio, as uploaded by the artists. \todo{See Table for a quantitative analysis of audio quality.}
+% TODO table quality
 
-\textbf{Audio.} most of the largest datasets are distributed with features, sometimes computed by commercial services like the Echonest.
-Researchers are wary of those proprietary features and have seek to compute features themselves, e.g.
-on the MSD \cite{msd_features}.
-In contrast, FMA provides audio for each song.
+\textbf{Complete.} In addition to audio, the dataset comes with a wealth of meta-data (see \tabref{tab:metadata}), including user data such as play counts, which are necessary to evaluate content-based recommender systems.
+% All meta-data from the online archive.
 
-\textbf{High quality audio.} full-length original audio uploaded by the artist.
-Stereo mp3 with bitrates varying between x.
-Audio, if released or available to download, is usually clips of 10 to 30 seconds.
-To our knowledge, none of the released datasets up to this date come with full-length audio.
+\textbf{Easy to work with.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.json} meta-data and \texttt{.mp3} audio, no need to scrap the web yourself.
+Moreover, we provide an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and some usage examples in the \texttt{usage.ipynb} Jupyter notebook to quick start using the data.
 
-\textbf{Future proof}: all audio and meta-data can be downloaded as ZIP archives, no need to scrap the web yourself.
-They are checksummed, have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by CERN.
+\textbf{Future proof and reproducible.} Archives are checksummed, have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by CERN.
 We thus alleviate the risks of tracks to become unavailable, to change without notice or for online services to shut down.
+Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) extract the sub-sets, (iv) compute the features and (v) the baselines, so that it can be reproduced or extended.
+As the songs and APIs are public, anybody can recreate or extend the collection.
 
 % Western / commercial music, how to qualify broadness?
-\textbf{Reproducible}: we share all the code used to (i) collect the dataset, (ii) analyze it, (iii) extract the sub-sets, (iv) compute the features and (v) the baselines, so that it can be reproduced or extended.
-As the original data and APIs are public, anybody can recreate the dataset.
-
-\textbf{Easy to work with}: working with the dataset only requires to download some archives.
-Moreover, we provide an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and some usage examples in the \texttt{usage.ipynb} Jupyter notebook to quick start using the data.
-%\end{itemize}
-
-% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
-% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
-
-% TODO: talk about related datasets
-% * SecondHandSongs dataset -> cover songs
-% * musiXmatch dataset -> lyrics
-% * Last.fm dataset -> song-level tags and similarity
-% * Taste Profile subset -> user data
-% * thisismyjam-to-MSD mapping -> more user data
-% * tagtraum genre annotations -> genre labels
-% * Top MAGD dataset -> more genre labels
-
-The sets in \tabref{tab:datasets} are either smaller, either provide audio through online services.
-
-From \tabref{tab:datasets}, it is obvious that the main problem is that larger datasets do not come with the audio.
-Either they contain only features derived from the audio, either they provide links for the user to download the audio from an online service.
-The problem with the former is that researchers are stuck with the set of chosen features, while the later make no assurance that the files or services are not gonna disappear or change without notice.
-We see from \tabref{tab:datasets} that most of the smallest datasets come with audio while most of the largest don't.
-That state of affair is due to copyright laws which forbid the redistribution of audio.
-The solution to that problem is to aim for tracks released under permissive licenses, such as \href{https://creativecommons.org/}{Creative Commons}.
-
-Creative Commons: MagnaTagATune
-Proprietary, non-redistributable: MSD, AudioSet?
 
 % Do not present like this because we might want representativeness. Which would necessarily include commercial music.
-
 %{\bf What would be the requirements for an ideal music dataset?} \\
@@
 \section{The Dataset} %Introducing the FMA dataset}
+% Describe what it is, what are the meta-data and features.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Describe what it is, what are the meta-data and features.
 
@@
 
+% not strictly genres (e.g. one of them is romantic dinner music)
+% one non-music category
+
 \begin{table}
@@
 	\label{fig:genres}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\linewidth]{genre_hierarchy.pdf}
+	\caption{Genre hierarchy example from the top-level Soul-RnB genre, \texttt{genre\_id}=14.}
+	\label{fig:genre_hierarchy}
 \end{figure}
@@
 
+% MSD website typical MIR tasks:
+% Segmentation
+% Automatic tagging
+% Year recognition
+% Imputation of missing data
+% Artist name, release, song title analysis
+% Preview audio
+% Yahoo ratings dataset
+% Visualization
+% Artist recognition
+% Cover song recognition
+% Lyrics
+
 genre classification, mood classification, artist identification/recognition, instrument recognition, music annotation (tags),
@@
 % Cannot: mood, style, annotations
+
+% Use: unsupervised learning for categorization, does genre emerge how ?
+% Task: classify 3secs excerpt. Much more samples, voting is often used.
 
@@
 Meta-data analysis: which genres are artists contributing to, drift in time, curators.
-It would also be interesting to study how good does it represent mainstream music.
+While this collection does obviously not contain hits from well-known artists, it would be interesting to study to what extent it is representative of mainstream music.
+%is representative of the commercial music that most users are interested in.
 
@@
 Training / testing division.
+
+\subsection{Limitations}
+
+\todo{what cannot be done currently}
+
+We can however expect people to cross-reference the dataset with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
 
@@
 
+% Spectrograms of audio data by concatenating 10 spectrograms of 3sec into a single long vector.
+% DL on spectrograms
+% DL on audio (should work better)
+
 \subsection{Methods}
+
+To assess the difficulty of the task we tested various standard classifiers on mainstream features.
+
+A major motivation to construct the dataset was to apply the powerful Deep Learning techniques to music analysis.
+As a preliminary result, we trained a convolutional neural network on 3sec-spectrograms, and we select the class by majority voting.
 
@@
 See \texttt{baselines.ipynb} for more details.
+
+%Table xxx reports the classification accuracy on the train and test datasets.\\
 
@@
 
-%\noindent
-%{\bf Baseline classification techniques.}  In this section, we consider perhaps the most popular music analysis problem, the music genre recognition (MGR) problem.
-We apply standard baseline classification techniques on 2 representations of audio data: \\
-%(1) Echonest medium-level 256-dim features, \\
-%(2) Spectrograms of audio data by concatenating 10 spectrograms of 3sec into a single long vector.\\
-%Table xxx reports the classification accuracy on the train and test datasets.\\
-%
-%
-%\noindent
-%{\bf Deep learning.} One major motivation to construct the FMA dataset was to apply the powerful deep learning techniques to the music analysis problem.
-As a preliminary result, we simply apply a convolutional neural network on the 3sec-spectrograms, and we select the class by majority voting.
-\\
-%Table xxx reports the classification accuracy on the train and test datasets.\\
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+% importance of open data
+
+Content-based music information retrieval tasks are typically solved with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
+Although that approach was dominant in the past, feature learning and end-to-end learning (from signal to label) have started to receive more attention from the MIR community in recent years. 
+The primary goal of this dataset is to enable researchers to train large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
+Hopefully, the release of such amount of easily accessible data will foster research which will ultimately surpass the performance ceiling observed in e.g.
+genre recognition, as it as been the case in computer vision with the release of ImageNet.
+
 The FMA enables researchers to test algorithms on a large-scale collection, thus allowing to test them on more real-world like environments.
 By providing audio, we do not limit the benchmarking to pre-computed features and allow scientists to develop and test new feature sets, learn features, or learn mappings directly from the audio.
+While MGR has been MIR's flagship application for a while, prediction accuracy reached a glass ceiling \cite{mgr_why}.
+We hope that the advent of a massive open dataset for that task will have the same effect as the release of ImageNet, i.e.
+a massive improvement in accuracy thanks to the advent of specialized deep learning architectures with testable performances which can be compared.
+Thanks to the availability of audio, deep learning architectures such as convolutional neural networks and recurrent neural networks can be directly applied to the waveform, bypassing any feature engineering.
+Such exploration may hopefully provide alternatives to solving MIR challenges \cite{mir_dl_feature_learning}.
+However, while this release is pretty large, we shall keep in mind that it is still two orders of magnitude behind commercial services who have access to tens of millions of songs.\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}}
 
-%\paragraph{Availability and reproducible research.}
-{\bf Availability and reproducible research.}
-All the datasets as well as Jupyter notebooks used to produce them, the
-analysis, and baselines as are available online at
-
-\url{https://github.com/mdeff/fma}
-
-Code to support the use of the data is provided as well.
-
-%\paragraph{Potential applications.}
-{\bf Potential applications.}
-We foresee multiple applications of this new dataset.
-The most obvious one is for music analysis with deep learning techniques.
-As raw music tracks are available, deep learning techniques like convolutional neural networks \cite{mnist} and recurrent neural networks \cite{art:HochreiterSchmidhuber97LSTM} can be directly applied.
-Besides, standard data analysis tasks such as classification, clustering, regression, text analysis, visualization can also be applied to this new music dataset.
-
-While this data release is pretty large, it is still two orders of magnitude behind commercial services who have access to tens of millions of songs.\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}}
-
-% last.fm: 45 million http://www.skilledtests.com/wiki/Last.fm_statistics
+% Once that problem is well advanced, researchers will be able to tackle more difficult tasks such as tagging or annotation.
+% Besides, standard data analysis tasks such as classification, clustering, regression, text analysis, visualization can also be applied to this new music dataset.
 

### 2017-03-21 fma.tex
update

@@
 \begin{abstract}
-
-The Free Music Archive (FMA) is an open music dataset that can be used to evaluate several tasks in music information retrieval (MIR).
-The released data is composed of 110,000 tracks, xx artists, xx albums, 164 genres, for a total of 229 days of audio for 900 GiB, all under permissive Creative Commons licenses.
-It features a collection of meta-data such as the song title, artist name, album or genres, together with user data such as play counts, full-length and high-quality audio, as well as a set of pre-computed features.
+The Free Music Archive (FMA) is an open and easily accessible music dataset that can be used to evaluate several tasks in music information retrieval (MIR).
+% TODO: which tasks
+While content-based MIR tasks have typically been solved with feature engineering,
+%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
+feature learning and end-to-end learning %, i.e. learning from signals to labels
+are receiving more attention from the community. 
+% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
+Those models are however data-hungry and this dataset is meant to foster research which will ultimately surpass the performance ceiling observed in e.g. genre recognition (MGR).
+% as it happened in computer vision with the release of ImageNet.
+The released data is composed of \todo{110,000 tracks, xx artists, xx albums}, 164 genres, for a total of 229 days of audio and 900 GiB, all under permissive Creative Commons licenses.
+It features a collection of meta-data such as the song title, artist name, album or genres, together with full-length and high-quality audio, user data such as play counts, as well as a set of pre-computed features.
 We propose a train / test split and three subsets: a small genre-balanced dataset of 4,000 songs and 10 genres, a medium genre-unbalanced dataset of 14,511 songs and 20 genres, and a 90 GiB version of the full set, with clips trimmed to 30s.
-This paper describes the dataset and how it was created, proposes some tasks and evaluate some baselines for music genre recognition (MGR).
+This paper describes the dataset and how it was created, proposes some tasks and evaluate some baselines for MGR.
 All the code to reproduce it as well as links to download the data and usage examples are available at \url{https://github.com/mdeff/fma}.
@@
 These datasets are free, legal, easily available, and have proved essential to advance the field.
-The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.5M images, allowed to demonstrate the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best \cite{convnet_imagenet}.
+The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.3M images \cite{imagenet_challenge}, allowed to demonstrate the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best \cite{convnet_imagenet}.
 
@@
 	\begin{threeparttable}
-	\begin{tabular}{lrrcc}
+	\begin{tabular}{l@{ }rrcc}
 		\toprule
@@
 	\begin{threeparttable}
-		\begin{tabular}{lr
-			>{$}r<{$}
+		\begin{tabular}{l@{\hspace{2.7em}}r
+			@{\hspace{2.7em}}>{$}r<{$}
 			@{}>{${}={}$}c@{}
 			>{$}l<{$}
-			>{$}c<{$}
-			cc}
+			@{\hspace{2.7em}}>{$}c<{$}
+			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
 		\toprule
@@
 		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
-			2,100,000 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & - & 243 \\
+			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & - & 241 \\
 		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
 			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 541 \\
-		\href{https://github.com/mdeff/fma/}{FMA} \tnote{3} &
+		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
 			110,000 & 180\cdot44\cdot10^3\cdot2 && 1.6\cdot10^7 & 1.7\cdot10^{12} & 900 & 229 \\
@@
 			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
-		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{4} &
-			14,000,000 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.4\cdot10^{12} & 1000 & - \\
+		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
+			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
+		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
+			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
 		\bottomrule
@@
 		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos.
-		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
-		\item[3] Average duration and sample rate. See \todo{ad} for details.
-		\item[4] Average resolution. Although most applications resize or crop to $256\cdot256$.
+		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}. See \todo{ad} for FMA.
+		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$, for $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
+		\item[4] Object detection / classification task. It is the most used dataset for image classification.
 	\end{tablenotes}
 	\end{threeparttable}
-	\caption{Size comparison.
+	\caption{Comparison between the largest public audio and image datasets.
 Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images.
 Scale is the number of samples times the dimensionality.
-Size is for a (zipped) archive of all mp3 or jpeg in GB, which is an indication of the quantity of information, e.g.
-duration or quality.
-The last column is the number of days to listen to the whole available audio.}
+Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GB, an indication of the quantity of information affected by the size and diversity of the data.
+The last column is the number of days necessary to listen to the whole available audio.}
 	\label{tab:size}
@@
 \textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with genre.
-Large collections are also more diverse and allows to average out annotation noise.
-While FMA is not the largest in term of number of clips, every other dataset with readily available quality audio are two orders of magnitude smaller.
-A possibly more important size metric here is the total amount of audio, i.e.
-the sum of the lengths of all clips, in which FMA is in a much better position (see \tabref{tab:size}).
-For comparison, ImageNet \cite{imagenet}, the largest benchmark dataset in computer vision features 14 million images.
-Despite a much larger collection, the amount of data is comparable to that of MSD or FMA because of the smaller dimensionality, as shown in \tabref{tab:size}.
-For statistical learning algorithms, the scale of FMA and MSD is thus comparable to ImageNet.
+They are also more diverse and allows to average out annotation noise.
+While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (see \tabref{tab:datasets}).
+For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (see \tabref{tab:size}) and shall thus be sufficient for the foreseable future.
+Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.
 
-\textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets.
-Most of these issues stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
-The only solution to that problem is to aim for tracks released under permissive licenses.
-All songs, i.e.
-all \texttt{.mp3} in the \texttt{.zip} archives, are licensed under Creative Commons by the artists, which allows redistribution.
-The meta-data produced by ourselves, i.e.
-all the \texttt{.json} files, is released under the terms of the \href{https://creativecommons.org/licenses/by/4.0)}{Creative Commons Attribution 4.0 International License (CC BY 4.0)} and the code, i.e.
-all the files in the git repository, under the MIT license.
+\textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets, which stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
+The sole solution to that problem is to aim for songs which license permits redistribution.
+Therfore all songs chosen for the FMA are licensed under Creative Commons by the artists.
+All meta-data and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{
+%Creative Commons Attribution 4.0 International License (CC BY 4.0)
+CC BY 4.0} and MIT licenses.
+% TODO paper CC BY 4.0 if published at ISMIR
 % TODO licensing
@@
 \textbf{High-quality audio.}
-\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the largest datasets don't.
-Either they only contain features derived from the audio, either they provide links for researchers to download the audio from an online service.
+\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger don't.
+They either contain features derived from the audio or provide links to download the audio from an online service.
 The problem with the former is that researchers are stuck with the set of chosen features and prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later makes no assurance that the files or services won't disappear or change without notice.
-Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{The Echo Nest} and have tried to extract features by themselves by downloading sample audio for the MSD from \href{https://www.7digital.com}{7digital.com}, which is tedious \cite{msd_features}.
+Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{The Echo Nest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
 Finally, released or available to download audio is usually clips of 10 to 30 seconds, sometimes of low quality.
-FMA is distributed with legally redistributable full-length and high-quality original audio, as uploaded by the artists. \todo{See Table for a quantitative analysis of audio quality.}
+In comparison, FMA is distributed with legally redistributable full-length, high-quality and non-modified audio. \todo{See Table ? for a quantitative analysis of quality.}
 % TODO table quality
 
-\textbf{Complete.} In addition to audio, the dataset comes with a wealth of meta-data (see \tabref{tab:metadata}), including user data such as play counts, which are necessary to evaluate content-based recommender systems.
+\textbf{Complete.} In addition to audio, the dataset comes with a wealth of meta-data, shown in \tabref{tab:metadata}.
+This includes per-song genre or user data such as play counts, which are necessary to evaluate content-based recommender systems.
 % All meta-data from the online archive.
 
-\textbf{Easy to work with.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.json} meta-data and \texttt{.mp3} audio, no need to scrap the web yourself.
+\textbf{Easily accessible.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.json} meta-data and \texttt{.mp3} audio, no need to scrap the web yourself.
 Moreover, we provide an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and some usage examples in the \texttt{usage.ipynb} Jupyter notebook to quick start using the data.
 
-\textbf{Future proof and reproducible.} Archives are checksummed, have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by CERN.
-We thus alleviate the risks of tracks to become unavailable, to change without notice or for online services to shut down.
+\textbf{Future proof and reproducible.} Archives are checksummed, have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}.
+Doing so alleviates the risks of tracks to become unavailable, to change without notice or for online services to shut down.
 Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) extract the sub-sets, (iv) compute the features and (v) the baselines, so that it can be reproduced or extended.
-As the songs and APIs are public, anybody can recreate or extend the collection.
+Moreover, anybody can recreate or extend the collection, thanks to public songs and APIs.
 
@@
 
-The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive} (FMA), an interactive library of high-quality, legal audio downloads directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
+The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library of high-quality, legal audio downloads directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
 The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
@@
 	\end{tabular}
-	\caption{List of per-song meta-data obtained from the FMA.}
+	\caption{List of per-song meta-data.}
 	\label{tab:metadata}
@@
 
+% Assess ground truth quality or augment genres by independent annotators, e.g. mechanical turk or crowdflowers
+% metadata such as genre, mood, groove, composer, performer, lyrics, meter, chord progressions,  instruments  present
+
 We can however expect people to cross-reference the dataset with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
@@
 
-Content-based music information retrieval tasks are typically solved with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
-Although that approach was dominant in the past, feature learning and end-to-end learning (from signal to label) have started to receive more attention from the MIR community in recent years. 
-The primary goal of this dataset is to enable researchers to train large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
-Hopefully, the release of such amount of easily accessible data will foster research which will ultimately surpass the performance ceiling observed in e.g.
-genre recognition, as it as been the case in computer vision with the release of ImageNet.
-
-The FMA enables researchers to test algorithms on a large-scale collection, thus allowing to test them on more real-world like environments.
+The FMA enables researchers to test algorithms on a large-scale collection, closer to real-world environments.
+Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of songs,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset (see \tabref{tab:size}), which is also far behind the amount of tagged photos \href{https://www.facebook.com}{Facebook} has access to for example.
 By providing audio, we do not limit the benchmarking to pre-computed features and allow scientists to develop and test new feature sets, learn features, or learn mappings directly from the audio.
-While MGR has been MIR's flagship application for a while, prediction accuracy reached a glass ceiling \cite{mgr_why}.
-We hope that the advent of a massive open dataset for that task will have the same effect as the release of ImageNet, i.e.
-a massive improvement in accuracy thanks to the advent of specialized deep learning architectures with testable performances which can be compared.
-Thanks to the availability of audio, deep learning architectures such as convolutional neural networks and recurrent neural networks can be directly applied to the waveform, bypassing any feature engineering.
-Such exploration may hopefully provide alternatives to solving MIR challenges \cite{mir_dl_feature_learning}.
-However, while this release is pretty large, we shall keep in mind that it is still two orders of magnitude behind commercial services who have access to tens of millions of songs.\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}}
+While genre recognition has been MIR's flagship application for a while, prediction accuracy reached a glass ceiling \cite{mgr_why}.
+We hope that the advent of a large open dataset for that task will have the same effect as the release of ImageNet, i.e.
+a massive improvement in accuracy thanks to the advent of specialized DL architectures with testable performances which can be compared.
+With availability of audio, DL architectures such as convolutional neural networks and recurrent neural networks can be applied to the waveform, treating it as a 1D vector along a time axis, to avoid any feature engineering.
+While those approaches have fallen behind learning from higher-level representations such as spectrograms \cite{dieleman_endtoend}, a greater exploration of the design space will hopefully provide alternatives to solving MIR challenges \cite{mir_dl_feature_learning}.
 
@@
 
+We are grateful to the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
 We are grateful to SWITCH and EPFL for hosting the dataset within the context
@@
 P-2 program}.
-
-We are grateful to the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
-
-We also want to thank the team behind \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
+% We also want to thank the team behind \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
 

### 2017-03-22 fma.tex
authorship

@@
 \oneauthor
-{Kirell Benzi, \hspace{.2cm} Michaël Defferrard, \hspace{.2cm} Pierre Vandergheynst, \hspace{.2cm} Xavier Bresson}
+{Michaël Defferrard, \hspace{.2cm} Kirell Benzi, \hspace{.2cm} Pierre Vandergheynst, \hspace{.2cm} Xavier Bresson}
  {{\tt michael.defferrard@epfl.ch, kirell.benzi@epfl.ch} \\

### 2017-03-23 fma.tex
edit after Kirell's feedback

@@
 \begin{abstract}
-The Free Music Archive (FMA) is an open and easily accessible music dataset that can be used to evaluate several tasks in music information retrieval (MIR).
+The Free Music Archive (FMA) is an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR).
 % TODO: which tasks
@@
 %with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
-feature learning and end-to-end learning %, i.e. learning from signals to labels
-are receiving more attention from the community. 
+the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
+That endeavor is however restrained by the limited availability of large audio datasets.
 % The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
-Those models are however data-hungry and this dataset is meant to foster research which will ultimately surpass the performance ceiling observed in e.g. genre recognition (MGR).
+By releasing the FMA we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g. genre recognition (MGR).
 % as it happened in computer vision with the release of ImageNet.
@@
 While the development of new mathematical models and algorithms to solve challenging real-world problems is obviously of first importance to any field of research, evaluation and comparison to the existing state-of-the-art is necessary for a technique to be widely adopted by research communities.
-Such tasks require benchmark datasets.
+Such tasks require open benchmark datasets to be reproducible.
 % to achieve their purpose, which are sometimes challenging to collect and share.
@@
 These datasets are free, legal, easily available, and have proved essential to advance the field.
-The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.3M images \cite{imagenet_challenge}, allowed to demonstrate the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best \cite{convnet_imagenet}.
+The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.3M images \cite{imagenet_challenge}, demonstrated the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best \cite{convnet_imagenet}, and unlocked incredible achievements in both fields \cite{dl}.
 
@@
 
+% Musicology
+
 % Use: unsupervised learning for categorization, does genre emerge how ?
@@
 
-Training / testing division.
+\subsection{Recommendation} % Recommender Systems}
+% Cite Kirell's paper
 
@@
 While the utility of music genre recognition (MGR) has been debated, mostly because of its ambiguity and cultural definition, it is widely used and understood by end-users who find it useful to discuss musical categories \cite{mgr_why}.
-Moreover, it is one of the most researched areas of MIR.
 
@@
 
+% Moreover, it is one of the most researched areas of MIR.
+While MGR has been MIR's flagship application for a while, prediction accuracy reached a glass ceiling \cite{mgr_why}.
+%thanks to the advent of specialized DL architectures with testable performances which can be compared.
+With availability of audio, DL architectures such as convolutional neural networks and recurrent neural networks can be applied to the waveform to avoid any feature engineering.
+While those approaches have fallen behind learning from higher-level representations such as spectrograms \cite{dieleman_endtoend}, a greater exploration of the design space will hopefully provide alternatives to solving MIR challenges \cite{mir_dl_feature_learning}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The FMA enables researchers to test algorithms on a large-scale collection, closer to real-world environments.
-Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of songs,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset (see \tabref{tab:size}), which is also far behind the amount of tagged photos \href{https://www.facebook.com}{Facebook} has access to for example.
+Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of songs,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset (see \tabref{tab:size}) which opened the door to dramatic performance improvements for many tasks in computer vision.
+%is also far behind the amount of tagged photos \href{https://www.facebook.com}{Facebook} has access to for example.
 By providing audio, we do not limit the benchmarking to pre-computed features and allow scientists to develop and test new feature sets, learn features, or learn mappings directly from the audio.
-While genre recognition has been MIR's flagship application for a while, prediction accuracy reached a glass ceiling \cite{mgr_why}.
-We hope that the advent of a large open dataset for that task will have the same effect as the release of ImageNet, i.e.
-a massive improvement in accuracy thanks to the advent of specialized DL architectures with testable performances which can be compared.
-With availability of audio, DL architectures such as convolutional neural networks and recurrent neural networks can be applied to the waveform, treating it as a 1D vector along a time axis, to avoid any feature engineering.
-While those approaches have fallen behind learning from higher-level representations such as spectrograms \cite{dieleman_endtoend}, a greater exploration of the design space will hopefully provide alternatives to solving MIR challenges \cite{mir_dl_feature_learning}.
+% talk about meta-data ?
+% Kirell: less DL oriented, some people don't like it. Insist on available audio.
 

### 2017-03-27 fma.tex
update

@@
 		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}. See \todo{ad} for FMA.
-		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$, for $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
+		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
 		\item[4] Object detection / classification task. It is the most used dataset for image classification.
@@
 \begin{enumerate}
-	\item \textbf{Full}: this is the complete dataset as described in \secref{sec:content}.
-	\item \textbf{Large}: this is the full dataset except audio is 30 seconds clips extracted from the middle of the tracks.
+	\item \textbf{Full}: the complete dataset, described in \secref{sec:content}.
+	\item \textbf{Large}: the full dataset with audio limited to 30 seconds clips extracted from the middle of the tracks. That reduces the size of the data by a factor 10.
 	\item \textbf{Medium}: all tracks which have Echonest features.
@@
 Can be down: genre, artist, popularity prediction, tags, year prediction
-% Cannot: mood, style, annotations
 
@@
 
+\subsection{Meta-data}
+
 Meta-data analysis: which genres are artists contributing to, drift in time, curators.
@@
 
-As it happened with e.g. the MSD, people may find and link other sources of data with the dataset, such as lyrics or tags associated to the songs, e.g. from last.fm.
+\subsection{Genre Recognition}
 
-We propose three genre classification problems of varying difficulty.
+We propose three genre classification problems of varying difficulty. We'll establish some baselines for that specific problem in the next section.
 
@@
 \todo{what cannot be done currently}
+% Cannot: mood, style, annotations
 
 % Assess ground truth quality or augment genres by independent annotators, e.g. mechanical turk or crowdflowers
+% Cross-verify on allmusic.com
 % metadata such as genre, mood, groove, composer, performer, lyrics, meter, chord progressions,  instruments  present
+% tags annotation
 
@@
 As a preliminary result, we trained a convolutional neural network on 3sec-spectrograms, and we select the class by majority voting.
+
+% Graph ConvNet
+% Would be good to popularize the method and make the connection with my core research. Will have to be compared to std ConvNets.
+% Node classification: each node is a track
+%	* powerful in a semi-supervised setting (traditional use of graphs in ML)
+%	* do this with Chibu or as a master thesis
+%	* artist relations --> disconnected graph
+%	* user relations (users playlists form communities)
+% Kirell [ICASSP recommender avec Vassilis]:
+%	* généralement c'est la similarité musicale
+%	* mix: metadata + audio features + social
+%	* on apprend une métrique qui pondère bien le tout par rapport au genre par exemple
+% Signal classification would mean finding a representation which can be supported by a graph. Notes on spectrograms ?
 
@@
 P-2 program}.
-% We also want to thank the team behind \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
+% We are grateful to \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
 

### 2017-03-28 fma.tex
change CRLF to LF
edit after Xavier's feedback

@@
 
-\oneauthor
-{Michaël Defferrard, \hspace{.2cm} Kirell Benzi, \hspace{.2cm} Pierre Vandergheynst, \hspace{.2cm} Xavier Bresson}
- {{\tt michael.defferrard@epfl.ch, kirell.benzi@epfl.ch} \\
-  {\tt pierre.vandergheynst@epfl.ch, xavier.bresson@epfl.ch} \\
-  LTS2, EPFL, Lausanne, Switzerland}
+\multauthor
+{Michaël Defferrard \hspace{.6cm} Kirell Benzi \hspace{.6cm} Pierre Vandergheynst \hspace{.6cm} Xavier Bresson$^1$}
+{
+	LTS2, EPFL, Lausanne, Switzerland \hspace{1cm}
+	{\small$^1$XB is now at NTU, Singapore. Work conducted when he was at EPFL.}\\
+	{\tt\small \{michael.defferrard,kirell.benzi,pierre.vandergheynst\}@epfl.ch, xbresson@ntu.edu.sg}
+  }
+
+%\thanks{Xavier Bresson is now with NTU, Singapore. This work was conducted when he was at EPFL.}
+
+%\multauthor
+%{$^1$ \hspace{1cm} Second author$^1$ \hspace{1cm} Third author$^2$} { \bfseries{Fourth author$^3$ \hspace{1cm} Fifth author$^2$ \hspace{1cm} Sixth author$^1$}\\
+%  $^1$ Department of Computer Science, University , Country\\
+%$^2$ International Laboratories, City, Country\\
+%$^3$  Company, Address\\
+%{\tt\small CorrespondenceAuthor@ismir.edu, PossibleOtherAuthor@ismir.edu}
+%}
+%\def\authorname{First author, Second author, Third author, Fourth author, Fifth author, Sixth author}
 
@@
 It features a collection of meta-data such as the song title, artist name, album or genres, together with full-length and high-quality audio, user data such as play counts, as well as a set of pre-computed features.
-We propose a train / test split and three subsets: a small genre-balanced dataset of 4,000 songs and 10 genres, a medium genre-unbalanced dataset of 14,511 songs and 20 genres, and a 90 GiB version of the full set, with clips trimmed to 30s.
-This paper describes the dataset and how it was created, proposes some tasks and evaluate some baselines for MGR.
-All the code to reproduce it as well as links to download the data and usage examples are available at \url{https://github.com/mdeff/fma}.
+We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 4,000 songs and 10 genres, a medium genre-unbalanced dataset of 14,511 songs and 20 genres, and a 90 GiB version of the full set, with clips trimmed to 30s.
+This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
+% Xavier: tasks, such as ...
+Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.
+%All the code to create,  as well as links to download the data and usage examples are available at \url{https://github.com/mdeff/fma}.
 
@@
 		\toprule
-		dataset & \#clips & \#artists & year & audio \\
+		dataset\tnote{1} & \#clips & \#artists & year & audio \\
 		\midrule
@@
 		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
-		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{1} \\
+		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
 		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
 		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
-		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{2} & 230 & 2009 & yes\tnote{3} \\ % MagnaTagATune
+		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
 		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
@@
 		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
-		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{1} \\
-		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{1} \\
+		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
+		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
 		\bottomrule
@@
 		% 103-Artists is also known as DB-L, Extended Ballroom
-		\item[1] Audio can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
-		\item[2] The 25,863 clips are cut from 5,405 songs.
-		\item[3] Low quality 16 kHz, 32kbps, mono mp3.
+		\item[1] Names are links to datasets' homepage.
+		\item[2] Audio can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
+		\item[3] The 25,863 clips are cut from 5,405 songs.
+		\item[4] Low quality 16 kHz, 32kbps, mono mp3.
 	\end{tablenotes}
@@
 % Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
+
+% Xavier: De meme pour Table 2, c'est interessant bien sur, mais encore une fois, les datasets de MSD et autres n'ont pas disponible (directement et facilement) les raw audio data.
+Cela rend confusing le main message.
+Aussi, je ne parlerais pas des MNIST, CIFAR et autres image datasets, ce n'est pas le point de ce papier, targeted music, et cela peut rendre notre message confusing.
 
@@
 
-%\subsection{The Free Music Archive}
+\subsection{The Free Music Archive}
 
@@
 	\centering
-	\begin{tabular}{|ll|}
+	\begin{tabular}{|llll|}
 		\hline
-		track\_id & track\_title \\
-		album\_title & artist\_name \\
+		track\_id & track\_title & album\_title & artist\_name \\
+		genres & play\_count & & \\
 		\hline
@@
 
-A train / test split is proposed to make research on the FMA reproducible.
+A split into training, validation and test sets is proposed to make research on the FMA reproducible. The training and validation sets should be merged if cross-validation is used instead.
 Below are the constraints we followed:
 \begin{enumerate}
-	\item For each genre, we randomly split train and test songs with the same ratio 80-20\%.
-	\item We forced an artist to be either in the train set, or the test set (meaning the artist's songs cannot be in both train and test sets).
-	\item We filled alternatively the train set and test set in the descending order of the count of artist songs. This forces the two sets to have the same distribution of songs per artist.
+	\item The songs were split with the same ratio 80/10/10\% across all genres, i.e. a ratio of about 6 between training and validation/test sets.
+	% PTB 19/3/3, MNIST 60/10, ImageNet 1200/50/100 --> ratio ~6-12
+	\item We forced an artist to appear in one set only, i.e. all songs from a given artist either appear in test, validation or train.
+	\item We alternatively filled the three sets in descending order of the count of artist songs. It forces the three sets to have the same distribution of songs per artist.
 \end{enumerate}
@@
 
-We are grateful to the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
-We are grateful to SWITCH and EPFL for hosting the dataset within the context
-of the \href{https://projects.switch.ch/scale-up}{SCALE-UP} project, funded in
-part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC
-P-2 program}.
+We want to thank the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
+In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{blog post about this dataset}, Cheyenne Hohman, the Director at the Archive, wrote that "by embracing the some rights reserved philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications"
+We are grateful to SWITCH and EPFL for hosting the dataset within the context of the \href{https://projects.switch.ch/scale-up}{SCALE-UP} project, funded in part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC P-2 program}.
+Xavier Bresson is supported by NRF Fellowship NRFF2017-10.
 % We are grateful to \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.

### 2017-03-29 fma.tex
update

@@
 %\subsection{Data Cleaning}
+% classes with less than 1% of the number of tracks of the biggest class Pop/Rock are removed
 
@@
 
-We extracted a wide range of audio features from the audio, using the \href{https://github.com/librosa/librosa}{librosa} Python library. An overview
-on these features is given in \tabref{tab:features}.
-In addition, a number of descriptive medium-level and high-level audio features extracted by \href{http://the.echonest.com/}{The Echo Nest}, a music platform for developers and media companies, are provided for some tracks.
-These features include tempo, loudness, timings of fade-in and fade-out, and MFCC-like features for a number of segments.
-Those features can be obtained either by uploading the songs to their website, which are then analyzed and sent back to the user, or simply retrieved from their database if the songs are known to them, i.e.
+% good features according to \cite{msd_features}: Statistical Spectrum Descriptors from rp_extract
+
+The upper part of \tabref{tab:features} lists the (mostly spectral) features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0.
+Each feature set is extracted on windows of 2048 samples spaced by 512 samples.
+We then computed 7 statistics over all windows: the mean, standard deviation, skew, kurtosis, median, minimum and maximum.
+All those pre-computed features are distributed in \texttt{features.csv} for all tracks.\footnote{The code to compute them can be found in \texttt{features.ipynb}}
+
+In addition,
+% a number of descriptive medium-level and high-level audio
+features extracted by \href{http://the.echonest.com/}{The Echo Nest}, a music platform, are provided for some tracks.
+These features include high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and MFCC-like temporal features for a number of segments \cite{echonest_temporal}.
+They are distributed in \texttt{echonest.csv}.
+%Echonest audio features are described in Kirell's paper
+%Those features can be obtained either by uploading the songs to their website, which are then analyzed and sent back to the user, or simply retrieved from their database if the songs are known to them, i.e.
 they were analyzed before.
@@
 In our case, we only retrieved features from songs which were already in their database.
-See \texttt{features.ipynb} for more details.
-
-\begin{table}
-	\centering
-	\begin{tabular}{llr}
-		\toprule
-		Feature set & Extractor & Dimensionality \\
-		\midrule
-		MFCCs \cite{mfcc} & librosa & 26 \\
-		MFCCs \cite{mfcc} & librosa & 26 \\
-		MFCCs \cite{mfcc} & librosa & 26 \\
-		\bottomrule
-	\end{tabular}
-	\caption{Overview of features extracted from the FMA samples.}
-	\label{tab:features}
-\end{table}
 
@@
 
-A split into training, validation and test sets is proposed to make research on the FMA reproducible. The training and validation sets should be merged if cross-validation is used instead.
+% PTB 19/3/3, MNIST 60/10, ImageNet 1200/50/100 --> ratio ~6-12
+A 80/10/10\% split into training, validation and test sets is proposed to make research on the FMA reproducible. The training and validation sets should be merged if cross-validation is used instead.
 Below are the constraints we followed:
+% Data was stratified according to the following criteria: ground truth, artist and times filters
 \begin{enumerate}
-	\item The songs were split with the same ratio 80/10/10\% across all genres, i.e. a ratio of about 6 between training and validation/test sets.
-	% PTB 19/3/3, MNIST 60/10, ImageNet 1200/50/100 --> ratio ~6-12
-	\item We forced an artist to appear in one set only, i.e. all songs from a given artist either appear in test, validation or train.
-	\item We alternatively filled the three sets in descending order of the count of artist songs. It forces the three sets to have the same distribution of songs per artist.
+	\item Stratified sampling was used to preserve the percentage of samples for each class. This is important for minority classes.
+	\item Stratified sampling with an artist filter, i.e. forcing an artist to appear in one set only. All songs from a given artist either appear in train, validation or test, never in two of them.
+	\item Stratified sampling with a time filter, i.e. for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
+	\item Finally, we alternatively filled the three sets in descending order of the count of artist songs.
+This forces the three sets to have the same distribution of songs per artist (in addition to songs per genre).
 \end{enumerate}
@@
 % MSD: metadata analysis, artist recognition, automatic music tagging, recommendation, cover song recognition, lyrics, limitations
+% MSD: Year Prediction
 
@@
 \subsection{Genre Recognition}
+% ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
 
@@
 \subsection{Recommendation} % Recommender Systems}
-% Cite Kirell's paper
+% Cite Kirell's paper: Song recommendation with non-negative matrix factorization and graph total variation
 
@@
 
+% Allmusic: The Web page offers a wide range of music information, including album reviews, artist biographies, discographies well as classification of albums according to genres, styles, moods and themes.
+Information is provided and curated by experts.
+% Genre information is very coarse, provided as a single tag for each album.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% One of the most common task, we give some baselines so that people have a reference about how hard it is.
-% ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
 
-% MSD: Year Prediction
+Music genres are categories that have arisen through a complex interplay of cultures, artists, and market forces to characterize similarities between compositions and organize music collections.
+Yet, the boundaries between genres still remain fuzzy, making the problem of music genre recognition (MGR) a nontrivial task.
+% genre not defined by audio content only
+While its utility has been debated, mostly because of its ambiguity and cultural definition, it is widely used and understood by end-users who find it useful to discuss musical categories \cite{mgr_why}.
+As such, it is one of the most researched areas of MIR.
+%\subsection{Data}
+The FMA is especially suited for MGR as it features fine genre information, i.e.
+multiple genres associated to individual songs, has a built-in genre hierarchy (\tabref{tab:genres}), and is annotated by the artists themselves.
+% While there certainly is noise in our labels, we trust the artists to have selected the most appropriate set of genres for their own songs.
 
-% genre not defined by audio content only
-While the utility of music genre recognition (MGR) has been debated, mostly because of its ambiguity and cultural definition, it is widely used and understood by end-users who find it useful to discuss musical categories \cite{mgr_why}.
+%\subsection{Methods}
 
-\subsection{Data}
+%\subsection{Evaluation and results}
 
-While there certainly is noise in our labels, we trust the artists to have selected the most appropriate set of genres for their own songs.
+\tabref{tab:mgr} reports classification accuracies obtained on the \todo{small} subset with \todo{14} different mainstream feature sets and 5 standard classifiers using scikit-learn \cite{scikit-learn}, version 0.18.1.
+Specifically, we employed \todo{Naive Bayes, Support Vector Machines (polynomial kernel with exponent 1), k-nearest Neighbours with $k=1$, a J48 Decision Tree, and Random Forests}.
+All classifiers were tested with mostly default settings.\footnote{See \texttt{baselines.ipynb} for all details.} Reported performances should not be taken as the state-art-of-the-art but rather as a lower-bound and an indication of the task's difficulty.
+Moreover, the produced code can serve as a reference and is easily modified to accommodate other features and classifiers.
 
-% Spectrograms of audio data by concatenating 10 spectrograms of 3sec into a single long vector.
-% DL on spectrograms
-% DL on audio (should work better)
+% As a preliminary result, we trained a convolutional neural network on 3sec-spectrograms, and we select the class by majority voting.
 
-\subsection{Methods}
+While MGR has been MIR's flagship application for a while, prediction accuracy reached a glass ceiling \cite{mgr_why}.
+A major motivation to construct this dataset was to enable the use of the powerful deep learning set of techniques to music analysis.
+%thanks to the advent of specialized DL architectures with testable performances which can be compared.
+With availability of audio, DL architectures such as convolutional neural networks and recurrent neural networks can be applied to the waveform to avoid any feature engineering.
+While those approaches have fallen behind learning from higher-level representations such as spectrograms \cite{dieleman_endtoend}, a greater exploration of the design space will hopefully provide alternatives to solving MIR challenges \cite{mir_dl_feature_learning}.
 
-To assess the difficulty of the task we tested various standard classifiers on mainstream features.
-
-A major motivation to construct the dataset was to apply the powerful Deep Learning techniques to music analysis.
-As a preliminary result, we trained a convolutional neural network on 3sec-spectrograms, and we select the class by majority voting.
-
-% Graph ConvNet
-% Would be good to popularize the method and make the connection with my core research. Will have to be compared to std ConvNets.
-% Node classification: each node is a track
-%	* powerful in a semi-supervised setting (traditional use of graphs in ML)
-%	* do this with Chibu or as a master thesis
-%	* artist relations --> disconnected graph
-%	* user relations (users playlists form communities)
-% Kirell [ICASSP recommender avec Vassilis]:
-%	* généralement c'est la similarité musicale
-%	* mix: metadata + audio features + social
-%	* on apprend une métrique qui pondère bien le tout par rapport au genre par exemple
-% Signal classification would mean finding a representation which can be supported by a graph. Notes on spectrograms ?
-
-\subsection{Evaluation and results}
-
-The performance of various classifiers on various features are shown in \tabref{tab:mgr} as baselines for future work.
-See \texttt{baselines.ipynb} for more details.
-
-%Table xxx reports the classification accuracy on the train and test datasets.\\
+%Some very recent advances are indeed promising Chuo
 
@@
 	\centering
-	\begin{tabular}{lccccc}
+	\begin{tabular}{lrrrrrr}
 		\toprule
-		Features & NB & SVM & kNN & DT & RF \\
+		Feature set & Dim. & NB & SVM & kNN & DT & RF \\
 		\midrule
-		MFCCs  & 30.12 & 30.12 & 30.12 & 30.12 \\
-		MFCCs  & 30.12 & 30.12 & 30.12 & 30.12 \\
-		MFCCs  & 30.12 & 30.12 & 30.12 & 30.12 \\
-		Echonest  & 30.12 & 30.12 & 30.12 & 30.12 \\
+		Chroma \cite{chroma} & 84 \\
+		Tonnetz \cite{tonnetz} & 42 \\
+		MFCC \cite{mfcc} & 140 \\
+		Spec. centroid & 7 \\
+		Spec. bandwidth & 7 \\
+		Spec. contrast \cite{contrast} & 49 \\
+		Spec. rolloff & 7 \\
+		RMSE & 7 \\ % Root-mean-square energy
+		Zero-crossing rate & 7 \\
+		\midrule
+		Ech. audio & 8 \\
+		Ech. social & 5 \\
+		Ech. temporal \cite{echonest_temporal} & 224 \\
+		\midrule
+		% Number and use them for aggregates
+		MFCC+Chroma & x \\
+		All & 150 \\
 		\bottomrule
 	\end{tabular}
-	\caption{Genre recognition results for various features and classifiers.}
+	%\caption{Genre recognition results for various features and classifiers.}
+	\caption{MGR results for various features and classifiers.}
+	% Todo small / medium / large / full
 	\label{tab:mgr}
+	\label{tab:features}
 \end{table}
-
-% Moreover, it is one of the most researched areas of MIR.
-While MGR has been MIR's flagship application for a while, prediction accuracy reached a glass ceiling \cite{mgr_why}.
-%thanks to the advent of specialized DL architectures with testable performances which can be compared.
-With availability of audio, DL architectures such as convolutional neural networks and recurrent neural networks can be applied to the waveform to avoid any feature engineering.
-While those approaches have fallen behind learning from higher-level representations such as spectrograms \cite{dieleman_endtoend}, a greater exploration of the design space will hopefully provide alternatives to solving MIR challenges \cite{mir_dl_feature_learning}.
 
@@
 
+Benchmarking is an important aspect in experimental sciences -- results reported by individual research groups need to be comparable.
+Important aspects of these are datasets that can be easily shared among researchers, together with a set of defined tasks and splits.
 The FMA enables researchers to test algorithms on a large-scale collection, closer to real-world environments.
@@
 % Besides, standard data analysis tasks such as classification, clustering, regression, text analysis, visualization can also be applied to this new music dataset.
+
+% Maybe in some future utopia, those kind of collaborations could shape a future where sharing is first and researchers feed open platforms with their algorithms while they feed us with data.
+That would especially be useful for recommendation systems.
 

### 2017-04-13 fma.tex
update

@@
 
-\newcommand{\todo}[1]{{\color{red} #1 }}
+\newcommand{\todo}[1]{{\color{red} #1}}
 
@@
 % as it happened in computer vision with the release of ImageNet.
-The released data is composed of \todo{110,000 tracks, xx artists, xx albums}, 164 genres, for a total of 229 days of audio and 900 GiB, all under permissive Creative Commons licenses.
-It features a collection of meta-data such as the song title, artist name, album or genres, together with full-length and high-quality audio, user data such as play counts, as well as a set of pre-computed features.
+The released data is composed of 106,930 tracks, 16,401 artists, 14,873 albums, 161 genres, for a total of 344 days of audio and 900 GiB, all under permissive Creative Commons licenses.
+It features a collection of metadata such as the song title, artist name, album or genres, together with full-length and high-quality audio, user data such as play counts, as well as a set of pre-computed features.
 We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 4,000 songs and 10 genres, a medium genre-unbalanced dataset of 14,511 songs and 20 genres, and a 90 GiB version of the full set, with clips trimmed to 30s.
@@
 
-% Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of meta-data for artists analysis or song ratings for recommender systems.
+% Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of metadata for artists analysis or song ratings for recommender systems.
 Existing datasets such as GTZAN, TagATune, and Million Song suffer from the previous limitations.
@@
 		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
-		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf 110,000 & \bf \todo{?} & \bf 2017 & \bf yes \\
+		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf 106,930 & \bf 16,401 & \bf 2017 & \bf yes \\
 		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
@@
 		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
-			110,000 & 180\cdot44\cdot10^3\cdot2 && 1.6\cdot10^7 & 1.7\cdot10^{12} & 900 & 229 \\
+			106,930 & 278\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & 900 & 343 \\
 		\midrule
@@
 As a result, despites its flaws \cite{gtzan_critic_1}, it continues to be the most used dataset for MGR \cite{mgr_eval}.
-The main limitations of GTZAN is its illegality, small size, and missing meta-data (no user ratings or even artist name and song title).
+The main limitations of GTZAN is its illegality, small size, and missing metadata (no user ratings or even artist name and song title).
 % Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
@@
 % They however each have limitations.
-MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes meta-data, audio features and raw audio.
+MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes metadata, audio features and raw audio.
 The poor audio quality and limited number of songs does however limit its usage.
@@
 % Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
-% Each song data is composed of high-level and medium-level audio features provided by Echonest service and meta-data like artist name, song title, track genre, etc.
+% Each song data is composed of high-level and medium-level audio features provided by Echonest service and metadata like artist name, song title, track genre, etc.
 It has its own dedicated website with related code and additional datasets.
@@
 Therfore all songs chosen for the FMA are licensed under Creative Commons by the artists.
-All meta-data and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{
+All metadata and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{
 %Creative Commons Attribution 4.0 International License (CC BY 4.0)
@@
 The problem with the former is that researchers are stuck with the set of chosen features and prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later  makes no assurance that the files or services won't disappear or change without notice.
-Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{The Echo Nest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
+Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
 Finally, released or available to download audio is usually clips of 10 to 30 seconds, sometimes of low quality.
@@
 
-\textbf{Complete.} In addition to audio, the dataset comes with a wealth of meta-data, shown in \tabref{tab:metadata}.
-This includes per-song genre or user data such as play counts, which are necessary to evaluate content-based recommender systems.
-% All meta-data from the online archive.
+\textbf{Complete.} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}.
+Basically all the metadata available through the API has been archived.
+In addition to the basics like title, album or artist names, this includes per-song genres and user data such as per-track/album/artist favorites, play counts and comments.
+% which are necessary to evaluate content-based recommender systems.
+There is also free-form text such as per-track/album/artist tags, album description and artist biography.
+% social such as artist location or website
+% images for tracks, albums and artists
+% tags are scarce.
 
-\textbf{Easily accessible.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.json} meta-data and \texttt{.mp3} audio, no need to scrap the web yourself.
-Moreover, we provide an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and some usage examples in the \texttt{usage.ipynb} Jupyter notebook to quick start using the data.
+\textbf{Easily accessible.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.csv} metadata and \texttt{.mp3} audio.
+No need to crawl the web and circumvent rate limits or access restrictions.
+Besides, we provide
+% Todo an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and
+some usage examples in the \texttt{usage.ipynb} Jupyter notebook to quick start using the data.
 
-\textbf{Future proof and reproducible.} Archives are checksummed, have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}.
+\textbf{Future proof and reproducible.} Archives are checksummed, \todo{have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}}.
 Doing so alleviates the risks of tracks to become unavailable, to change without notice or for online services to shut down.
-Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) extract the sub-sets, (iv) compute the features and (v) the baselines, so that it can be reproduced or extended.
-Moreover, anybody can recreate or extend the collection, thanks to public songs and APIs.
+Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) generate the subsets and splits, (iv) compute the features and (v) test the baselines, so that it can be reproduced or extended.
+Finally, anybody can recreate or extend the collection, thanks to public songs and APIs.
 
@@
 %(1) Availability of raw audio tracks like 30sec or more,\\
-%(2) Availability of meta-data such as genres, artist, title, year, lyrics, users' ratings, users' comments, etc,\\
+%(2) Availability of metadata such as genres, artist, title, year, lyrics, users' ratings, users' comments, etc,\\
 %(3) Large-scale dataset to be able to sample the high-dimensional distribution of music genres\\
@@
 \section{The Dataset} %Introducing the FMA dataset}
-% Describe what it is, what are the meta-data and features.
+% Describe what it is, what are the metadata and features.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 		\hline
-		track\_id & track\_title & album\_title & artist\_name \\
+		track\_id & album\_id & artist\_id & title \\
+		duration & bitrate & samplerate \\
+		\#listens & interest & \#comments & \#favorites \\
 		genres & play\_count & & \\
 		\hline
+		album\_id & title & \#tracks \\
+		date\_created & date\_released \\
+		\#listens & \#comments & \#favorites \\
+		\hline
+		artist\_id & name & location \\
+		\#comments & \#favorites \\
+		\hline
 	\end{tabular}
-	\caption{List of per-song meta-data.}
+	\caption{List of per-track, per-album and per-artist metadata contained in \texttt{tracks.csv}, \texttt{albums.csv} and \texttt{artists.csv}.}
+	% genres.csv ?
 	\label{tab:metadata}
@@
 
-Song meta-data, such as the title, album, name of the artist, genres as well as some usage statistics, was crawled using the available \href{https://freemusicarchive.org/api}{API}.
+Song metadata, such as the title, album, name of the artist, genres as well as some usage statistics, was crawled using the available \href{https://freemusicarchive.org/api}{API}.
 See \tabref{tab:metadata} for a list and \texttt{webapi.ipynb} for how to query the API with our helpers.
-The collected meta-information is stored in the \texttt{tracks.json} table, which \tabref{tab:tracks} shows an excerpt.
-Additionally, the used genre hierarchy, shown in \tabref{tab:genres}, was collected and stored in \texttt{genres.json}.
-Audio was then downloaded over HTTP.
-The collection of FMA was done in \todo{April 2016}, and the size of the dataset at this time was 89,912 songs.
+The collected meta-information is stored in the \texttt{tracks.csv} table, which \tabref{tab:tracks} shows an excerpt.
+As of April 1st 2017, the date of collection, the online archive largest track id was 155,321, of which 109,727 were valid (probably some tracks went deleted).
+Additionally, the used genre hierarchy, shown in \tabref{tab:genres}, was collected and stored in \texttt{genres.csv}.
+Audio was then downloaded over HTTPS.
 
@@
 
-The FMA\_large dataset of 77,643 data contains songs of 30sec length, and artist name, track title, track genres, track listens as meta-data.
+The FMA\_large dataset of 77,643 data contains songs of 30sec length, and artist name, track title, track genres, track listens as metadata.
 Our original motivation was to give the full audio tracks, not limited to 30sec, but the size of the whole music tracks is 806.8GB (for approximatively 256 days of music listening) making the hosting a challenging issue.
@@
 	\end{tabular}
-	\caption{Some rows and columns of the table stored in \texttt{tracks.json}. \todo{representative and varied}}
+	\caption{Some rows and columns of the table stored in \texttt{tracks.csv}. \todo{representative and varied}}
 	\label{tab:tracks}
@@
 
-All tracks come with the meta-data listed in \tabref{tab:metadata}, although some values may be missing.
+All tracks come with the metadata listed in \tabref{tab:metadata}, although some values may be missing.
 All audio tracks are mp3-encoded with sampling rate of 44,100Hz, bit rate 128kb/s, and in stereo.
+
+\todo{describe the metadata}
 
@@
 	\end{tabular}
-	\caption{Number of songs per genre. Top part is the 16 top-level genres, bottom part is the second- and third-level genres. The genre hierarchy is stored in \texttt{genres.json}.}
+	\caption{Number of songs per genre. Top part is the 16 top-level genres, bottom part is the second- and third-level genres. The genre hierarchy is stored in \texttt{genres.csv}.}
 	\label{tab:genres}
@@
 % a number of descriptive medium-level and high-level audio
-features extracted by \href{http://the.echonest.com/}{The Echo Nest}, a music platform, are provided for some tracks.
+features extracted by \href{http://the.echonest.com/}{Echonest}, a music platform, are provided for some tracks.
 These features include high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and MFCC-like temporal features for a number of segments \cite{echonest_temporal}.
@@
 In our case, we only retrieved features from songs which were already in their database.
+
+\subsection{Subsets} \label{sec:subsets}
+
+\begin{table}
+	\centering
+	\begin{tabular}{lrrccc}
+		\toprule
+		dataset & clips & genres & length & \multicolumn{2}{c}{Size} \\
+		\cmidrule{5-6}
+		        &         &          &  [s]   & [GB] & [days] \\
+		\midrule
+		small  &   8,000 &   8 &  30 & 3.4  & 2.8  \\
+		medium &  25,000 &  16 &  30 & 12.2 & 8.7  \\
+		large  & 104,469 & 161 &  30 & 90   & 36.3 \\
+		full   & 106,930 & 161 & 278 & 900  & 344  \\
+		\bottomrule
+	\end{tabular}
+	\caption{Proposed subsets of the FMA.}
+	\label{tab:subsets}
+\end{table}
+
+Finally, for the dataset to be useful as a development set or for people with lower computational resources, we propose the following subsets, each of which is a subset or the larger set:
+\begin{enumerate}
+	\item \textbf{Full}: the complete dataset, described in \secref{sec:content}.
+	\item \textbf{Large}: the full dataset with audio limited to 30 seconds clips extracted from the middle of the tracks (or entire song if shorter than 30 seconds).
+That reduces the size of the data by a factor 10.
+	\item \textbf{Medium}: all tracks which have Echonest features.
+	\item \textbf{Small}: tracks from 10 top-level genres, balanced with 400 tracks per genre, 1 genre per track.
+This subset is similar to the very popular GTZAN in terms of size and genre distribution with the added benefits of the FMA, that is metadata, pre-computed features and copyright-free audio.
+\end{enumerate}
+\tabref{tab:subsets} highlights the main differentiating factors between the proposed subsets.
+See \texttt{analysis.ipynb} for an analysis of the subsets' content similar to that shown in \secref{sec:content}.
 
@@
 \end{enumerate}
-A track is assigned in train or split for all the subsets and the above constraints are satisfied for all sets. See \texttt{creation.ipynb} for full details.
-
-\subsection{Subsets} \label{sec:subsets}
-
-\begin{table}
-	\centering
-	\begin{tabular}{lrrccc}
-		\toprule
-		dataset & clips & genres & length & \multicolumn{2}{c}{Size} \\
-		\cmidrule{5-6}
-		        &         &          &  [s]   & [GB] & [days] \\
-		\midrule
-		small  &   4,000 & 10  & 30    & 3.4  & 1.4  \\
-		medium &  14,511 & 20  & 30    & 12.2 & 5.1  \\
-		large  & 110,000 & 164 & 30    & 90   & 26.9 \\
-		full   & 110,000 & 164 & <1800 & 900  & 229  \\
-		\bottomrule
-	\end{tabular}
-	\caption{Proposed subsets of the FMA.}
-	\label{tab:subsets}
-\end{table}
-
-Finally, for the dataset to be useful as a development set or for people with lower computational resources, we propose the following subsets, each of which is a subset or the larger set:
-\begin{enumerate}
-	\item \textbf{Full}: the complete dataset, described in \secref{sec:content}.
-	\item \textbf{Large}: the full dataset with audio limited to 30 seconds clips extracted from the middle of the tracks. That reduces the size of the data by a factor 10.
-	\item \textbf{Medium}: all tracks which have Echonest features.
-	\item \textbf{Small}: tracks from 10 top-level genres, balanced with 400 tracks per genre, 1 genre per track.
-This subset is similar to the very popular GTZAN in terms of size and genre distribution with the added benefits of the FMA, that is meta-data, pre-computed features and copyright-free audio.
-\end{enumerate}
-\tabref{tab:subsets} highlights the main differentiating factors between the proposed subsets.
-See \texttt{analysis.ipynb} for an analysis of the subsets' content similar to that shown in \secref{sec:content}.
+A track is assigned in the same split across all the subsets. The above constraints are satisfied for all subsets. See \texttt{creation.ipynb} for full details.
 
@@
 
-\subsection{Meta-data}
+\subsection{Metadata}
 
@@
 %is representative of the commercial music that most users are interested in.
+
+% How similar to commercially available music / other datasets / common sets
 
@@
 
+* Single top\_genres prediction on medium and small (16 top-level genres, 1 top-level genre per track)
+* Multiple top\_genres on large and full (16 top-level genres, 0 to 8 top-level genres per track)
+* Multiple genres on all subsets (161 genres, 0 to 25 genres per track)
+
 \subsection{Recommendation} % Recommender Systems}
 % Cite Kirell's paper: Song recommendation with non-negative matrix factorization and graph total variation
+
+The availability of user data makes the FMA suitable to evaluate content-based recommender systems.
 
@@
 % tags annotation
+
+% musicbrainz, playme, 7digital
 
@@
 
-\tabref{tab:mgr} reports classification accuracies obtained on the \todo{small} subset with \todo{14} different mainstream feature sets and 5 standard classifiers using scikit-learn \cite{scikit-learn}, version 0.18.1.
+\tabref{tab:mgr} reports classification accuracies obtained on the \todo{small} subset with \todo{15} different mainstream feature sets and 5 standard classifiers using scikit-learn \cite{scikit-learn}, version 0.18.1.
 Specifically, we employed \todo{Naive Bayes, Support Vector Machines (polynomial kernel with exponent 1), k-nearest Neighbours with $k=1$, a J48 Decision Tree, and Random Forests}.
@@
 		\midrule
-		Chroma \cite{chroma} & 84 \\
-		Tonnetz \cite{tonnetz} & 42 \\
-		MFCC \cite{mfcc} & 140 \\
-		Spec. centroid & 7 \\
-		Spec. bandwidth & 7 \\
-		Spec. contrast \cite{contrast} & 49 \\
-		Spec. rolloff & 7 \\
-		RMSE & 7 \\ % Root-mean-square energy
-		Zero-crossing rate & 7 \\
+		1 Chroma \cite{chroma} & 84 \\
+		2 Tonnetz \cite{tonnetz} & 42 \\
+		3 MFCC \cite{mfcc} & 140 \\
+		4 Spec. centroid & 7 \\
+		5 Spec. bandwidth & 7 \\
+		6 Spec. contrast \cite{contrast} & 49 \\
+		7 Spec. rolloff & 7 \\
+		8 RMSE & 7 \\ % Root-mean-square energy
+		9 Zero-crossing rate & 7 \\
 		\midrule
@@
 		% Number and use them for aggregates
-		MFCC+Chroma & x \\
+		3 + 6 & 189 \\
+		3 + 6 + 1 & 273 \\
 		All & 150 \\
@@
 	\end{tabular}
-	%\caption{Genre recognition results for various features and classifiers.}
-	\caption{MGR results for various features and classifiers.}
-	% Todo small / medium / large / full
+	\caption{Genre recognition results for various features and classifiers on the FMA \todo{small} subset.}
 	\label{tab:mgr}
@@
 By providing audio, we do not limit the benchmarking to pre-computed features and allow scientists to develop and test new feature sets, learn features, or learn mappings directly from the audio.
-% talk about meta-data ?
+% talk about metadata ?
 % Kirell: less DL oriented, some people don't like it. Insist on available audio.
@@
 We want to thank the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
-In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{blog post about this dataset}, Cheyenne Hohman, the Director at the Archive, wrote that "by embracing the some rights reserved philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications"
+In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{blog post about this dataset}, Cheyenne Hohman, the Director at the Archive, wrote that "by embracing the some rights reserved philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications".
 We are grateful to SWITCH and EPFL for hosting the dataset within the context of the \href{https://projects.switch.ch/scale-up}{SCALE-UP} project, funded in part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC P-2 program}.
+% We are grateful to \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
 Xavier Bresson is supported by NRF Fellowship NRFF2017-10.
-% We are grateful to \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
 

### 2017-04-16 fma.tex
update

@@
 \newcommand{\todo}[1]{{\color{red} #1}}
+
+\newcommand{\ntracks}{106,574 }
+\newcommand{\nartists}{16,341 }
+\newcommand{\nalbums}{14,854 }
+\newcommand{\ngenres}{161 }
+\newcommand{\tduration}{343 }
+\newcommand{\aduration}{278 }
+\newcommand{\size}{900 }
+
+\newcommand{\tspace}{\hspace{0.7em}}
 
@@
 % as it happened in computer vision with the release of ImageNet.
-The released data is composed of 106,930 tracks, 16,401 artists, 14,873 albums, 161 genres, for a total of 344 days of audio and 900 GiB, all under permissive Creative Commons licenses.
-It features a collection of metadata such as the song title, artist name, album or genres, together with full-length and high-quality audio, user data such as play counts, as well as a set of pre-computed features.
-We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 4,000 songs and 10 genres, a medium genre-unbalanced dataset of 14,511 songs and 20 genres, and a 90 GiB version of the full set, with clips trimmed to 30s.
+The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
+It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography and tags; together with full-length and high-quality audio as well as some pre-computed features.
+We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 8,000 songs from 8 major genres, a medium genre-unbalanced dataset of 25,000 songs from 16 genres, and a 90 GiB version of the full set with clips trimmed to 30s.
 This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
@@
 		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
-		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf 106,930 & \bf 16,401 & \bf 2017 & \bf yes \\
+		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
 		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
@@
 		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
-			106,930 & 278\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & 900 & 343 \\
+			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
 		\midrule
@@
 		\hline
-		track\_id & album\_id & artist\_id & title \\
-		duration & bitrate & samplerate \\
-		\#listens & interest & \#comments & \#favorites \\
-		genres & play\_count & & \\
+		track\_id & title & number & \\
+		information & language\_code & license & \\
+		composer & publisher & lyricist & \\
+		genres & genres\_all & genre\_top & \\
+		duration & bit\_rate & interest & \\ %samplerate \\
+		\#listens & \#comments & \#favorites & tags \\
+		date\_created & date\_recorded & & \\
 		\hline
-		album\_id & title & \#tracks \\
-		date\_created & date\_released \\
-		\#listens & \#comments & \#favorites \\
+		album\_id & title & type & \#tracks \\
+		information & engineer & producer & \\
+		\#listens & \#comments & \#favorites & tags \\
+		date\_created & date\_released & & \\
 		\hline
-		artist\_id & name & location \\
-		\#comments & \#favorites \\
+		artist\_id & name & & \\
+		bio & members & \multicolumn{2}{l|}{associated\_labels} \\
+		website & wikipedia\_page & \multicolumn{2}{l|}{related\_projects} \\
+		location & longitude & latitude & \\
+		& \#comments & \#favorites & tags \\
+		date\_created & active\_year\_end & \multicolumn{2}{l|}{active\_year\_begin} \\
 		\hline
 	\end{tabular}
-	\caption{List of per-track, per-album and per-artist metadata contained in \texttt{tracks.csv}, \texttt{albums.csv} and \texttt{artists.csv}.}
+	\caption{List of per-track, per-album and per-artist metadata distributed in \texttt{tracks.csv}.}
+	%Not all of them are filled for each track.\footnote{See }
+	%\texttt{albums.csv} and \texttt{artists.csv}.}
 	% genres.csv ?
@@
 
-Song metadata, such as the title, album, name of the artist, genres as well as some usage statistics, was crawled using the available \href{https://freemusicarchive.org/api}{API}.
-See \tabref{tab:metadata} for a list and \texttt{webapi.ipynb} for how to query the API with our helpers.
-The collected meta-information is stored in the \texttt{tracks.csv} table, which \tabref{tab:tracks} shows an excerpt.
-As of April 1st 2017, the date of collection, the online archive largest track id was 155,321, of which 109,727 were valid (probably some tracks went deleted).
-Additionally, the used genre hierarchy, shown in \tabref{tab:genres}, was collected and stored in \texttt{genres.csv}.
-Audio was then downloaded over HTTPS.
-
-\todo{Top-level genres.} Then, we created a top genre by simply picking the first genre in the original list of track genres provided by the artists.
+% Collection
+As of April 1st 2017, the date of collection, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks).
+In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.}.
+Finally, mp3 audio was downloaded over HTTPS.
+% Cleaning
+Out of all collected track ids, 180 mp3s could not be downloaded, 286 could not be trimmed by ffmpeg, and features could not be extracted from 71.
+Finally, the license of 2,616 songs prohibited their redistribution, leaving us with \ntracks tracks.
 
@@
 
-\todo{Adjust when redone.}
-As a first filtering, we kept the data that has audio tracks, reducing the size to 86,886 music data. 
-We removed data without genre information, giving 79,947 song data.
-At this stage the number of genres is 138, but half have only a few samples.
+%At this stage the number of genres is 138, but half have only a few samples.
 Thus we decided to only retain the genres with more than 100 samples.
@@
 
-The FMA\_large dataset of 77,643 data contains songs of 30sec length, and artist name, track title, track genres, track listens as metadata.
-Our original motivation was to give the full audio tracks, not limited to 30sec, but the size of the whole music tracks is 806.8GB (for approximatively 256 days of music listening) making the hosting a challenging issue.
-
-See \texttt{creation.ipynb} for full details.
-
 \begin{table*}
@@
 	\centering
-	\begin{tabular}{lrlrllll}
+\begin{tabular}{@{}l@{\tspace}l@{\tspace}l@{\tspace}l@{\tspace}r@{\tspace}r@{\tspace}l@{\tspace}r@{\tspace}l@{\tspace}l@{\tspace}l@{}}
 	\toprule
-	 fma\_id &                                             artist &  play\_count &                                              title &                                             genres &            top\_genre  \\
-	 \toprule
-		10 &                                          Kurt Vile &       42936 &                                            Freeway &                                              [Pop] &                  Pop  \\
-		134 &                                               AWOL &         880 &                                       Street Music &                                          [Hip-Hop] &              Hip-Hop  \\
-		 142 &                    Alec K.
-Redfearn \& the Eyesores &         670 &                               Punjabi Watery Grave &                                             [Folk] &                 Folk  \\
-		 236 &                                       Banana Clipz &        6695 &                              Push Am (Left, Right) &                              [Electronic, African] &           Electronic  \\
-		 237 &                                          Barnacled &        1056 &                    Garbage and Fire &                                             [Jazz] &                 Jazz  \\
-		 461 &                              Cantonement Jazz Band &        3933 &                                           Bessemer &                               [Blues, Jazz: Vocal] &                Blues  \\
-		824 &                     Here Comes A Big Black Cloud!!
-&         391 &                                         Black Mold &          [Rock, Loud-Rock, Psych-Rock, Indie-Rock] &                 Rock  \\
-		 825 &                     Here Comes A Big Black Cloud!!
-&        1847 &                                        Death March &          [Rock, Loud-Rock, Psych-Rock, Indie-Rock] &                 Rock  \\
-		837 &                                          Heroin UK &         146 &                                           DopeSick &                                             [Rock] &                 Rock \\
-		 889 &                                 Illusion of Safety &         163 &                                          Wasteland &                                       [Electronic] &           Electronic  \\
-		 896 &                                        Impediments &        1542 &                                               2012 &                                  [Punk, Power-Pop] &                 Punk  \\
-			  992 &                                      Jason Willett &         560 &                   Beautiful Song w/ kick drum solo &                                             [Rock] &                 Rock  \\
+	\multicolumn{6}{@{}l}{track} & \multicolumn{3}{@{}l}{album} & \multicolumn{2}{@{}l}{artist} \\
+	track\_id & title & genres\_all & genre\_top & dur. & listens & title & listens & tags & name & location \\
+	\midrule
+	150073 & Welcome to Asia & [2, 79] & International & 81 & 683 & Reprise & 4091 & [world music, dubtronica, fusion] & DubRaJah & Russia \\
+	140943 & Sleepless Nights & [322, 5] & Classical & 246 & 1777 & Creative Commons Vol. 7 & 28900 & [classical, alternate, soundtrack, piano, ... & Dexter Britain & United Kingdom \\
+	64604 & i dont want to die alone & [32, 38, 456] & Experimental & 138 & 830 & Summer Gut String & 7408 & [improvised, minimalist, noise, ... & Buildings and Mountains & Oneonta, NY \\
+	23500 & A Life In A Day & [236, 286, 15] & Electronic & 264 & 1149 & A Life in a Day & 6691 & [idm, candlestick, romanian, candle, ... & Candlestickmaker & Romania \\
+	131150 & Yeti-Bo-Betty & [25, 12, 85] & Rock & 124 & 183 & No Life After Crypts & 3594 & [richmond, fredericksburg, trash rock, ... & The Crypts! & Fredericksburg \\
 	\bottomrule
 	\end{tabular}
-	\caption{Some rows and columns of the table stored in \texttt{tracks.csv}. \todo{representative and varied}}
+	\caption{Some rows and columns of the table stored in \texttt{tracks.csv}.}
 	\label{tab:tracks}
@@
 
-\subsection{Content} \label{sec:content}
+\subsection{Content}
 % Figures and tables from analysis.ipynb
@@
 
-All tracks come with the metadata listed in \tabref{tab:metadata}, although some values may be missing.
-All audio tracks are mp3-encoded with sampling rate of 44,100Hz, bit rate 128kb/s, and in stereo.
+The collected track, album and artist meta-informations\footnote{Raw data is available in \texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, and \texttt{raw\_artists.csv}.} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
+That file is a relational table where each rows represent a track.
+\tabref{tab:metadata} lists the collected song metadata and user data, although some values may be missing for some tracks.
+The audio for each track is stored in a file wich name is the track id.
+All songs are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo.
+See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.
 
-\todo{describe the metadata}
+%\todo{describe the metadata}
 
-%\subsubsection{Genres}
-
-While it is difficult to construct a hierarchy of genres, we follow that of our source.
-\tabref{tab:genres} and \figref{fig:genres} show the distribution of the \todo{??} most common genres.
-As often encountered with natural categories, it obeys a power law.
-See \texttt{analysis.ipynb} for more details.
-
-% not strictly genres (e.g. one of them is romantic dinner music)
-% one non-music category
+\subsection{Genres}
 
@@
 	\centering
-	\begin{tabular}{cclc}
+	\begin{tabular}{ccclr}
 		\toprule
-		ID & Parent & Genre & \#songs \\
+		id & parent & top\_level & title & \#tracks \\
 		\midrule
-		4 & None & Electronic & 50,000 \\
-		4 & None & Electronic & 50,000 \\
-		4 & None & Electronic & 50,000 \\
+		38 & None & 38 & Experimental & 38,154 \\
+		15 & None & 15 & Electronic & 34,413 \\
+		12 & None & 12 & Rock & 32,923 \\
+		1235 & None & 1235 & Instrumental & 14,938 \\
+		10 & None & 10 & Pop & 13,845 \\
+		17 & None & 17 & Folk & 12,706 \\
 		\midrule
-		45 & 2 & Psych-Rock & 500 \\
-		45 & 2 & Psych-Rock & 500 \\
-		45 & 2 & Psych-Rock & 500 \\
+		25 & 12 & 12 & Punk & 9,261 \\
+		89 & 25 & 12 & Post-Punk & 1,858 \\
+		1  & 38 & 38 & Avant-Garde & 8,693 \\
 		\bottomrule
 	\end{tabular}
-	\caption{Number of songs per genre. Top part is the 16 top-level genres, bottom part is the second- and third-level genres. The genre hierarchy is stored in \texttt{genres.csv}.}
+	\caption{An excerpt of the genre hierarchy stored in \texttt{genres.csv}. Top part is some of the 16 top-level genres, bottom part is some second- and third-level genres.}
 	\label{tab:genres}
@@
 	\centering
-	\includegraphics[width=\linewidth]{histo_large.pdf}
-	\caption{Power law distribution of the top 20 genres.}
-	\label{fig:genres}
+	\includegraphics[width=\linewidth]{genre_distribution.pdf}
+	\caption{Power law distribution of the genres.}
+	\label{fig:genre_distribution}
 \end{figure}
@@
 	\includegraphics[width=0.5\linewidth]{genre_hierarchy.pdf}
-	\caption{Genre hierarchy example from the top-level Soul-RnB genre, \texttt{genre\_id}=14.}
+	\caption{Example of genre hierarchy for the top-level Soul-RnB genre. Left number is the \texttt{genre\_id}, right number is the number of tracks.}
 	\label{fig:genre_hierarchy}
@@
 
+While it is difficult to devise a hierarchy of genres, we followed that of our source.
+Indeed, the website offers a parent genre for each of the 163 available genres.
+\tabref{tab:genres} shows an excerp of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree.
+\figref{fig:genre_hierarchy} shows an excerpt of the tree.
+In the per-track table, the \texttt{genres} column contain the genre ids indicated by the artist.
+Then, given such hierarchical information, we constructed a \texttt{genres\_all} column which contains all the genres encountered when traversing the tree from the indicated genres to the roots.
+Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
+\figref{fig:genre_distribution} shows the distribution of the most common genres. As often encountered with natural categories, it obeys a power law.
+
+% MSD:
+% not strictly genres (e.g. one of them is romantic dinner music)
+% one non-music category
+
 \subsection{Features} % Feature sets
 
-% good features according to \cite{msd_features}: Statistical Spectrum Descriptors from rp_extract
-
 The upper part of \tabref{tab:features} lists the (mostly spectral) features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0.
-Each feature set is extracted on windows of 2048 samples spaced by 512 samples.
-We then computed 7 statistics over all windows: the mean, standard deviation, skew, kurtosis, median, minimum and maximum.
-All those pre-computed features are distributed in \texttt{features.csv} for all tracks.\footnote{The code to compute them can be found in \texttt{features.ipynb}}
+Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples.
+Seven statistics were then computed over all windows: the mean, standard deviation, skew, kurtosis, median, minimum and maximum.
+Those 518 pre-computed features are distributed in \texttt{features.csv} for all tracks.\footnote{See \texttt{features.py} for the code which computed the features.}
 
@@
 		\toprule
-		dataset & clips & genres & length & \multicolumn{2}{c}{Size} \\
+		dataset & clips & genres & length & \multicolumn{2}{c}{size} \\
 		\cmidrule{5-6}
@@
 		medium &   25,000 &  16 &  30 & 12.2 & 8.7  \\
-		large  & 104,469 & 161 &  30 & 90   & 36.3 \\
-		full   & 106,930 & 161 & 278 & 900  & 344  \\
+		large  & \ntracks & 161 &  30 & 90   & 37 \\
+		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
 		\bottomrule
@@
 \begin{enumerate}
-	\item \textbf{Full}: the complete dataset, described in \secref{sec:content}.
+	\item \textbf{Full}: the complete dataset, described above.
 	\item \textbf{Large}: the full dataset with audio limited to 30 seconds clips extracted from the middle of the tracks (or entire song if shorter than 30 seconds).
-That reduces the size of the data by a factor 10.
-	\item \textbf{Medium}: all tracks which have Echonest features.
-	\item \textbf{Small}: tracks from 10 top-level genres, balanced with 400 tracks per genre, 1 genre per track.
-This subset is similar to the very popular GTZAN in terms of size and genre distribution with the added benefits of the FMA, that is metadata, pre-computed features and copyright-free audio.
+That trimming reduces the size of the data by a factor 10.
+	\item \textbf{Medium}: 25,000 popular tracks with good metadata, e.g. only one top-level genre, and quality audio.
+	\item \textbf{Small}: 8,000 tracks from 8 top-level genres, balanced with 1000 tracks per genre, 1 top genre per track.
+This subset is similar to the very popular GTZAN \cite{gtzan} with the added benefits of the FMA, that is metadata, pre-computed features and copyright-free audio.
 \end{enumerate}
 \tabref{tab:subsets} highlights the main differentiating factors between the proposed subsets.
-See \texttt{analysis.ipynb} for an analysis of the subsets' content similar to that shown in \secref{sec:content}.
+%See \texttt{analysis.ipynb} for an analysis of the subsets' content similar to that shown in \secref{sec:content}.
 
@@
 \begin{enumerate}
-	\item Stratified sampling was used to preserve the percentage of samples for each class. This is important for minority classes.
-	\item Stratified sampling with an artist filter, i.e. forcing an artist to appear in one set only. All songs from a given artist either appear in train, validation or test, never in two of them.
-	\item Stratified sampling with a time filter, i.e. for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
-	\item Finally, we alternatively filled the three sets in descending order of the count of artist songs.
-This forces the three sets to have the same distribution of songs per artist (in addition to songs per genre).
+	\item Stratified sampling was used to preserve the percentage of tracks per genre. This is important for minority genres.
+	\item An artist filter was used, i.e. an artist was forced to appear in one set only. %All songs from a given artist either appear in train, validation or test, never in two of them.
+	%\item Stratified sampling with a time filter, i.e. for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
+	\item The three sets were forced to have the same distribution of songs per artist. %The three sets were alternatively filled in descending order of the count of artist songs. 
 \end{enumerate}
-A track is assigned in the same split across all the subsets. The above constraints are satisfied for all subsets. See \texttt{creation.ipynb} for full details.
+The above constraints are satisfied for all subsets, and a track is assigned to the same split across all of them. See \texttt{creation.ipynb} for full details.
 

### 2017-04-17 fma.tex
more figures

@@
 \usepackage{enumitem} \setlist{itemsep=0pt,topsep=0pt,parsep=0pt,partopsep=0pt}
+
+\usepackage{caption}
+%\captionsetup[figure]{skip=0pt}
 
@@
 
+\begin{figure*}
+	\begin{minipage}{0.74\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{growth.pdf}
+	\end{minipage} \hfill
+	\begin{minipage}{0.24\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{album_release_year.pdf}
+	\end{minipage}
+	\vspace{-1em}
+	\caption{(Left) Growth of the archive, created in 2008. (Right) Number of albums released per year (1902-2017).}
+	\label{fig:growth}
+	\label{fig:album_release_year}
+\end{figure*}
+
+\begin{table*}
+	\tiny
+	\centering
+\begin{tabular}{@{}l@{\tspace}l@{\tspace}l@{\tspace}l@{\tspace}r@{\tspace}r@{\tspace}l@{\tspace}r@{\tspace}l@{\tspace}l@{\tspace}l@{}}
+	\toprule
+	\multicolumn{6}{@{}l}{track} & \multicolumn{3}{@{}l}{album} & \multicolumn{2}{@{}l}{artist} \\
+	track\_id & title & genres\_all & genre\_top & dur. & listens & title & listens & tags & name & location \\
+	\midrule
+	150073 & Welcome to Asia & [2, 79] & International & 81 & 683 & Reprise & 4091 & [world music, dubtronica, fusion] & DubRaJah & Russia \\
+	140943 & Sleepless Nights & [322, 5] & Classical & 246 & 1777 & Creative Commons Vol. 7 & 28900 & [classical, alternate, soundtrack, piano, ... & Dexter Britain & United Kingdom \\
+	64604 & i dont want to die alone & [32, 38, 456] & Experimental & 138 & 830 & Summer Gut String & 7408 & [improvised, minimalist, noise, ... & Buildings and Mountains & Oneonta, NY \\
+	23500 & A Life In A Day & [236, 286, 15] & Electronic & 264 & 1149 & A Life in a Day & 6691 & [idm, candlestick, romanian, candle, ... & Candlestickmaker & Romania \\
+	131150 & Yeti-Bo-Betty & [25, 12, 85] & Rock & 124 & 183 & No Life After Crypts & 3594 & [richmond, fredericksburg, trash rock, ... & The Crypts! & Fredericksburg \\
+	\bottomrule
+	\end{tabular}
+	\caption{Some rows and columns of the table stored in \texttt{tracks.csv}.}
+	\label{tab:tracks}
+\end{table*}
+
+\begin{table}
+	\centering
+	\begin{tabular}{|llll|}
+		\hline
+		track\_id & title & number & \\
+		information & language\_code & license & \\
+		composer & publisher & lyricist & \\
+		genres & genres\_all & genre\_top & \\
+		duration & bit\_rate & interest & \\ %samplerate \\
+		\#listens & \#comments & \#favorites & tags \\
+		date\_created & date\_recorded & & \\
+		\hline
+		album\_id & title & type & \#tracks \\
+		information & engineer & producer & \\
+		\#listens & \#comments & \#favorites & tags \\
+		date\_created & date\_released & & \\
+		\hline
+		artist\_id & name & & \\
+		bio & members & \multicolumn{2}{l|}{associated\_labels} \\
+		website & wikipedia\_page & \multicolumn{2}{l|}{related\_projects} \\
+		location & longitude & latitude & \\
+		& \#comments & \#favorites & tags \\
+		date\_created & active\_year\_end & \multicolumn{2}{l|}{active\_year\_begin} \\
+		\hline
+	\end{tabular}
+	\caption{List of per-track, per-album and per-artist metadata distributed in \texttt{tracks.csv}.}
+	%Not all of them are filled for each track.\footnote{See }
+	%\texttt{albums.csv} and \texttt{artists.csv}.}
+	% genres.csv ?
+	\label{tab:metadata}
+\end{table}
+
 % Besides, the most influential competition in the field of music analysis organized every year is the Music Information Retrieval Evaluation eXchange (MIREX).\href{http://www.music-ir.org/mirex/wiki/}{MIREX} proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc.
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{duration_distribution.pdf}
+	\caption{Gaussian distribution of track durations.}
+	\label{fig:duration_distribution}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{listens_distribution.pdf}
+	\caption{Power-law distribution of album play counts.}
+	\label{fig:listens_distribution}
+\end{figure}
+
 \subsection{The Free Music Archive}
@@
 
-\begin{table}
-	\centering
-	\begin{tabular}{|llll|}
-		\hline
-		track\_id & title & number & \\
-		information & language\_code & license & \\
-		composer & publisher & lyricist & \\
-		genres & genres\_all & genre\_top & \\
-		duration & bit\_rate & interest & \\ %samplerate \\
-		\#listens & \#comments & \#favorites & tags \\
-		date\_created & date\_recorded & & \\
-		\hline
-		album\_id & title & type & \#tracks \\
-		information & engineer & producer & \\
-		\#listens & \#comments & \#favorites & tags \\
-		date\_created & date\_released & & \\
-		\hline
-		artist\_id & name & & \\
-		bio & members & \multicolumn{2}{l|}{associated\_labels} \\
-		website & wikipedia\_page & \multicolumn{2}{l|}{related\_projects} \\
-		location & longitude & latitude & \\
-		& \#comments & \#favorites & tags \\
-		date\_created & active\_year\_end & \multicolumn{2}{l|}{active\_year\_begin} \\
-		\hline
-	\end{tabular}
-	\caption{List of per-track, per-album and per-artist metadata distributed in \texttt{tracks.csv}.}
-	%Not all of them are filled for each track.\footnote{See }
-	%\texttt{albums.csv} and \texttt{artists.csv}.}
-	% genres.csv ?
-	\label{tab:metadata}
-\end{table}
-
 % Collection
 As of April 1st 2017, the date of collection, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks).
+See \figref{fig:growth} to visualize the growth of the dataset.
 In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.}.
@@
 
-\begin{table*}
-	\tiny
-	\centering
-\begin{tabular}{@{}l@{\tspace}l@{\tspace}l@{\tspace}l@{\tspace}r@{\tspace}r@{\tspace}l@{\tspace}r@{\tspace}l@{\tspace}l@{\tspace}l@{}}
-	\toprule
-	\multicolumn{6}{@{}l}{track} & \multicolumn{3}{@{}l}{album} & \multicolumn{2}{@{}l}{artist} \\
-	track\_id & title & genres\_all & genre\_top & dur. & listens & title & listens & tags & name & location \\
-	\midrule
-	150073 & Welcome to Asia & [2, 79] & International & 81 & 683 & Reprise & 4091 & [world music, dubtronica, fusion] & DubRaJah & Russia \\
-	140943 & Sleepless Nights & [322, 5] & Classical & 246 & 1777 & Creative Commons Vol. 7 & 28900 & [classical, alternate, soundtrack, piano, ... & Dexter Britain & United Kingdom \\
-	64604 & i dont want to die alone & [32, 38, 456] & Experimental & 138 & 830 & Summer Gut String & 7408 & [improvised, minimalist, noise, ... & Buildings and Mountains & Oneonta, NY \\
-	23500 & A Life In A Day & [236, 286, 15] & Electronic & 264 & 1149 & A Life in a Day & 6691 & [idm, candlestick, romanian, candle, ... & Candlestickmaker & Romania \\
-	131150 & Yeti-Bo-Betty & [25, 12, 85] & Rock & 124 & 183 & No Life After Crypts & 3594 & [richmond, fredericksburg, trash rock, ... & The Crypts! & Fredericksburg \\
-	\bottomrule
-	\end{tabular}
-	\caption{Some rows and columns of the table stored in \texttt{tracks.csv}.}
-	\label{tab:tracks}
-\end{table*}
-
 \subsection{Content}
@@
 The collected track, album and artist meta-informations\footnote{Raw data is available in \texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, and \texttt{raw\_artists.csv}.} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
-That file is a relational table where each rows represent a track.
-\tabref{tab:metadata} lists the collected song metadata and user data, although some values may be missing for some tracks.
+That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks.
 The audio for each track is stored in a file wich name is the track id.
 All songs are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo.
+Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
 See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.
@@
 \begin{figure}
+	\begin{minipage}[t]{0.44\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{genre_hierarchy.pdf}
+	\end{minipage} \hfill
+	\begin{minipage}[t]{0.55\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{genres_per_track.pdf}
+	\end{minipage}
+	\vspace{-1em}
+	\caption{(Left) Example of genre hierarchy for the top-level Soul-RnB genre.
+Left number is the \texttt{genre\_id}, right number is the number of tracks per genre.
+(Right) Number of genres per track (\texttt{genres\_all}).}
+	\label{fig:genre_hierarchy}
+	\label{fig:genres_per_track}
+\end{figure}
+
+\begin{figure}[t]
 	\centering
 	\includegraphics[width=\linewidth]{genre_distribution.pdf}
-	\caption{Power law distribution of the genres.}
+	\caption{Power law distribution of genres.}
 	\label{fig:genre_distribution}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\linewidth]{genre_hierarchy.pdf}
-	\caption{Example of genre hierarchy for the top-level Soul-RnB genre. Left number is the \texttt{genre\_id}, right number is the number of tracks.}
-	\label{fig:genre_hierarchy}
 \end{figure}
@@
 
-\begin{table}
-	\centering
-	\begin{tabular}{lrrccc}
-		\toprule
-		dataset & clips & genres & length & \multicolumn{2}{c}{size} \\
-		\cmidrule{5-6}
-		        &       &        &  [s]   & [GB] & [days] \\
-		\midrule
-		small  &    8,000 &   8 &  30 & 3.4  & 2.8  \\
-		medium &   25,000 &  16 &  30 & 12.2 & 8.7  \\
-		large  & \ntracks & 161 &  30 & 90   & 37 \\
-		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
-		\bottomrule
-	\end{tabular}
-	\caption{Proposed subsets of the FMA.}
-	\label{tab:subsets}
-\end{table}
-
 Finally, for the dataset to be useful as a development set or for people with lower computational resources, we propose the following subsets, each of which is a subset or the larger set:
@@
 The above constraints are satisfied for all subsets, and a track is assigned to the same split across all of them. See \texttt{creation.ipynb} for full details.
+
+\begin{table}
+	\centering
+	\begin{tabular}{lrrccc}
+		\toprule
+		dataset & clips & genres & length & \multicolumn{2}{c}{size} \\
+		\cmidrule{5-6}
+		        &       &        &  [s]   & [GB] & [days] \\
+		\midrule
+		small  &    8,000 &   8 &  30 & 3.4  & 2.8  \\
+		medium &   25,000 &  16 &  30 & 12.2 & 8.7  \\
+		large  & \ntracks & 161 &  30 & 90   & 37 \\
+		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
+		\bottomrule
+	\end{tabular}
+	\caption{Proposed subsets of the FMA.}
+	\label{tab:subsets}
+\end{table}
 

### 2017-04-17 fma.tex
update

@@
 
-\usepackage{caption}
-%\captionsetup[figure]{skip=0pt}
-
 \newcommand{\todo}[1]{{\color{red} #1}}
@@
 
-\newcommand{\tspace}{\hspace{0.7em}}
+\newcommand{\tspacea}{\hspace{0.7em}}
+\newcommand{\tspaceb}{\hspace{0.9em}}
 
@@
   }
-
-%\thanks{Xavier Bresson is now with NTU, Singapore. This work was conducted when he was at EPFL.}
 
@@
 		\toprule
-		Dataset & \#samples & \multicolumn{3}{c}{Dimensionality} & \multicolumn{3}{c}{Size} \\
+		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
 		\cmidrule{6-8}
@@
 		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos.
-		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}. See \todo{ad} for FMA.
+		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
 		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
@@
 	\end{minipage}
-	\vspace{-1em}
-	\caption{(Left) Growth of the archive, created in 2008. (Right) Number of albums released per year (1902-2017).}
+	\caption{(left) Growth of the archive, created in 11/2008. (right) Number of albums released per year, from 1902 to 2017.}
 	\label{fig:growth}
@@
 	\centering
-\begin{tabular}{@{}l@{\tspace}l@{\tspace}l@{\tspace}l@{\tspace}r@{\tspace}r@{\tspace}l@{\tspace}r@{\tspace}l@{\tspace}l@{\tspace}l@{}}
+\begin{tabular}{@{}l@{\tspacea}l@{\tspacea}l@{\tspacea}l@{\tspacea}r@{\tspacea}r@{\tspacea}l@{\tspacea}r@{\tspacea}l@{\tspacea}l@{\tspacea}l@{}}
 	\toprule
@@
 	\end{tabular}
-	\caption{Some rows and columns of the table stored in \texttt{tracks.csv}.}
+	\caption{Some rows and columns of the metadata table, stored in \texttt{tracks.csv}.}
 	\label{tab:tracks}
@@
 	\end{tabular}
-	\caption{List of per-track, per-album and per-artist metadata distributed in \texttt{tracks.csv}.}
+	\caption{List of available per-track, per-album and per-artist metadata, i.e. the columns of \texttt{tracks.csv}.}
 	%Not all of them are filled for each track.\footnote{See }
@@
 Finally, released or available to download audio is usually clips of 10 to 30 seconds, sometimes of low quality.
-In comparison, FMA is distributed with legally redistributable full-length, high-quality and non-modified audio. \todo{See Table ? for a quantitative analysis of quality.}
+In comparison, FMA is distributed with legally redistributable full-length, high-quality and non-modified audio.
+% \todo{See Table ? for a quantitative analysis of quality.}
 % TODO table quality
@@
 
-\textbf{Future proof and reproducible.} Archives are checksummed, \todo{have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}}.
+\textbf{Future proof and reproducible.} All files and archives are checksummed and hosted in a long-term digital archive.
+%\todo{have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}}.
 Doing so alleviates the risks of tracks to become unavailable, to change without notice or for online services to shut down.
@@
 	\includegraphics[width=\linewidth]{duration_distribution.pdf}
-	\caption{Gaussian distribution of track durations.}
+	\caption{Track duration, from 0 to 3 hours.}
 	\label{fig:duration_distribution}
@@
 	\includegraphics[width=\linewidth]{listens_distribution.pdf}
-	\caption{Power-law distribution of album play counts.}
+	\caption{Album listens, from 0 to 3.6 millions.}
 	\label{fig:listens_distribution}
@@
 	\end{tabular}
-	\caption{An excerpt of the genre hierarchy stored in \texttt{genres.csv}. Top part is some of the 16 top-level genres, bottom part is some second- and third-level genres.}
+	\caption{An excerpt of the genre hierarchy, stored in \texttt{genres.csv}.
+Some of the 16 top-level genres appear in the top part, while some second- and third-level genres appear in the bottom part.}
 	\label{tab:genres}
 \end{table}
-
-\begin{figure}
-	\begin{minipage}[t]{0.44\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{genre_hierarchy.pdf}
-	\end{minipage} \hfill
-	\begin{minipage}[t]{0.55\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{genres_per_track.pdf}
-	\end{minipage}
-	\vspace{-1em}
-	\caption{(Left) Example of genre hierarchy for the top-level Soul-RnB genre.
-Left number is the \texttt{genre\_id}, right number is the number of tracks per genre.
-(Right) Number of genres per track (\texttt{genres\_all}).}
-	\label{fig:genre_hierarchy}
-	\label{fig:genres_per_track}
-\end{figure}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\linewidth]{genre_distribution.pdf}
-	\caption{Power law distribution of genres.}
-	\label{fig:genre_distribution}
-\end{figure}
 
@@
 % one non-music category
+
+\begin{figure}
+	\begin{minipage}[t]{0.44\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{genre_hierarchy.pdf}
+	\end{minipage} \hfill
+	\begin{minipage}[t]{0.55\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{genres_per_track.pdf}
+	\end{minipage}
+	\caption{(left) Example of genre hierarchy for the top-level Soul-RnB genre.
+Left number is the \texttt{genre\_id}, right number is the number of tracks per genre.
+(right) Number of genres per track, from 0 to 31 (\texttt{genres\_all}).}
+	\label{fig:genre_hierarchy}
+	\label{fig:genres_per_track}
+\end{figure}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\linewidth]{genre_distribution.pdf}
+	\caption{Tracks per genre, from 1 to 38,154.}
+	\label{fig:genre_distribution}
+\end{figure}
 
@@
 	\item \textbf{Full}: the complete dataset, described above.
+All \ngenres genres, unbalanced with 1 to 38,154 tracks per genre (\figref{fig:genre_distribution}) and 0 to 31 genres per track (\figref{fig:genres_per_track}).
 	\item \textbf{Large}: the full dataset with audio limited to 30 seconds clips extracted from the middle of the tracks (or entire song if shorter than 30 seconds).
 That trimming reduces the size of the data by a factor 10.
-	\item \textbf{Medium}: 25,000 popular tracks with good metadata, e.g. only one top-level genre, and quality audio.
-	\item \textbf{Small}: 8,000 tracks from 8 top-level genres, balanced with 1000 tracks per genre, 1 top genre per track.
-This subset is similar to the very popular GTZAN \cite{gtzan} with the added benefits of the FMA, that is metadata, pre-computed features and copyright-free audio.
+	\item \textbf{Medium}: 25,000 of the most popular tracks with good metadata and quality audio. Genre unbalanced with 21 to 7,103 tracks per top genre, but only one top genre per track.
+	\item \textbf{Small}: 8,000 tracks from 8 top genres, balanced with 1000 tracks per genre, 1 genre per track.
+This subset is similar to the very popular GTZAN \cite{gtzan} with the added benefits of the FMA, that is metadata, pre-computed features, and copyright-free audio.
 \end{enumerate}
@@
 % PTB 19/3/3, MNIST 60/10, ImageNet 1200/50/100 --> ratio ~6-12
-A 80/10/10\% split into training, validation and test sets is proposed to make research on the FMA reproducible. The training and validation sets should be merged if cross-validation is used instead.
+An 80/10/10\% split into training, validation and test sets is proposed to make research on the FMA reproducible. The training and validation sets should be merged if cross-validation is used instead.
 Below are the constraints we followed:
@@
 	%\item Stratified sampling with a time filter, i.e. for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
-	\item The three sets were forced to have the same distribution of songs per artist. %The three sets were alternatively filled in descending order of the count of artist songs. 
+	\item Artists with lots of songs were distributed between the sets. %The three sets were alternatively filled in descending order of the count of artist songs. 
 \end{enumerate}
@@
 
-% MSD: metadata analysis, artist recognition, automatic music tagging, recommendation, cover song recognition, lyrics, limitations
-% MSD: Year Prediction
+With its rich set of metadata, user data, audio and features, the FMA is amenable to many tasks in MIR.
+We share below some possible uses, which serves to illustrate the breadth of data available in the dataset.
 
-% MSD website typical MIR tasks:
+% Metadata analysis
+% Recommendation / Recommender systems, Music Similarity
+% Music classification
+	% Genre recognition / classification
+	% Artist identification / recognition
+	% Year prediction / recognition
+% Automatic (music) tagging / autotagging --> tag prediction
+
+%On an other hand, unsupervised learning methods could be used for music categorization. Does genre emerge? How?
+
 % Segmentation
-% Automatic tagging
-% Year recognition
-% Imputation of missing data
-% Artist name, release, song title analysis
-% Preview audio
-% Yahoo ratings dataset
 % Visualization
-% Artist recognition
-% Cover song recognition
-% Lyrics
-
-genre classification, mood classification, artist identification/recognition, instrument recognition, music annotation (tags),
-Artist Identification, Genre classification, Mood Classification, Instrument identification, Music Similarity, Autotagging, Automatic playlist generation
-
-Can be down: genre, artist, popularity prediction, tags, year prediction
-
 % Musicology
 
-% Use: unsupervised learning for categorization, does genre emerge how ?
-% Task: classify 3secs excerpt. Much more samples, voting is often used.
+%\subsection{Album covers}
+% image covers: generate covers per genre / artist
+% study the influence of covers on ratings
 
-% TODO
-% Problems
-% Genre classif / mutli-genre classif
-% Recommendation
-% Play count prediction
-% Covers: generate covers per genre / artist
+\subsection{Metadata Analysis}
 
-\subsection{Metadata}
-
-Meta-data analysis: which genres are artists contributing to, drift in time, curators.
-While this collection does obviously not contain hits from well-known artists, it would be interesting to study to what extent it is representative of mainstream music.
-%is representative of the commercial music that most users are interested in.
-
-% How similar to commercially available music / other datasets / common sets
-
-\subsection{Genre Recognition}
-% ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
-
-We propose three genre classification problems of varying difficulty. We'll establish some baselines for that specific problem in the next section.
-
-\begin{enumerate}
-	\item Easy, based on \texttt{fma\_small.zip}: 10 top-level genres, balanced with 400 tracks per genre, 1 genre per track
-	\item Medium, based on \texttt{fma\_medium.zip}: all 16 top-level genres, unbalanced with x-y tracks per genre, multiple genres per track
-	\item Hard, based on \texttt{fma\_large.zip}: all 164 genres, unbalanced with x-y tracks per genre, multiple genres per track
-\end{enumerate}
-
-* Single top\_genres prediction on medium and small (16 top-level genres, 1 top-level genre per track)
-* Multiple top\_genres on large and full (16 top-level genres, 0 to 8 top-level genres per track)
-* Multiple genres on all subsets (161 genres, 0 to 25 genres per track)
+While the original intention of the dataset was to release a large volume of audio for machine learning algorithms, analyzing metadata from such a collection is also extremely interesting.
+For instance, one could address questions like: How broads artists are, i.e.
+to which genres do they contribute to?
+Do we observe some drift in time for the vocabulary used in song titles, album names or artist names?
+To which extent tags correspond to genres?
+What other semantic categories can we found in them?
+Other lines of work are to test algorithms for the imputation of missing data, or to study to what extent is the FMA representative of mainstream music, or how similar it is to other datasets.
 
@@
 
-The availability of user data makes the FMA suitable to evaluate content-based recommender systems.
+Due to the potential commercial value of a working system, music recommendation and music similarity are ones of the best studied areas of MIR.
+So far, content-based systems have fallen short at predicting user ratings when compared to collaborative filtering methods.
+The availability of user data such as play counts for tracks and albums as well as number of comments and favorites for tracks, albums and artists opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features, with a clean ground truth.
+
+\subsection{Music Classification and Annotation}
+%\subsection{Automatic Music Tagging}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{tag_distribution.pdf}
+	\caption{Per-track, album and artist tags, from 0 to 150.}
+	\label{fig:tag_distribution}
+\end{figure}
+
+Music classification is a key problem in MIR with many potential applications.
+For one, a classification system enables end users to search for the types of music they are interested in.
+On the other hand, different music types are managed more effectively and efficiently once they are categorized into different groups \cite{mir_review_classif}.
+The classification tasks which can be evaluated on the available labels for the FMA include genre recognition, artist identification, year prediction, and music annotation.
+Automatic tagging is a classification problem which cover different semantic categories, where tags are labels which can be any musical term that describes the genre, mood, instrumentation, and style of the song.
+%The task of music annotation has recently gained much popularity in the MIR community.
+In the FMA, tags can be found at the track, album and artist level: 23,496 tracks (22\%), 2,659 albums (18\%) and 1,605 artists (10\%) have at least one tag.
+\figref{fig:tag_distribution} shows their distribution.\footnote{One of the artist tag is often the artist name, which we subtracted in our counts.}
 
 \subsection{Limitations}
+% ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
+% musicbrainz, playme, 7digital
+% Allmusic: album reviews, artist biographies, discographies as well as classification of albums according to genres, styles, moods and themes. Information is provided and curated by experts.
 
-\todo{what cannot be done currently}
-% Cannot: mood, style, annotations
+% Cover song recognition
+% Lyrics
+% Automatic playlist generation
+% style?
+% Music classification
+	% mood classification,
+	% instrument recognition / identification
+	% music annotation --> temporal labels, e.g. instrument / note
 
-% Assess ground truth quality or augment genres by independent annotators, e.g. mechanical turk or crowdflowers
-% Cross-verify on allmusic.com
 % metadata such as genre, mood, groove, composer, performer, lyrics, meter, chord progressions, instruments present
-% tags annotation
 
-% musicbrainz, playme, 7digital
-
+While the FMA can be used to evaluate many tasks, metadata is missing for e.g. cover song recognition, mood classification or instrument recognition.
 We can however expect people to cross-reference the dataset with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
 
-% Allmusic: The Web page offers a wide range of music information, including album reviews, artist biographies, discographies well as classification of albums according to genres, styles, moods and themes.
-Information is provided and curated by experts.
-% Genre information is very coarse, provided as a single tag for each album.
+Diversity is another issue: as suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental music, electronic and rock.
+Moreover, it probably does not contain hits from renown artists or commercial music.
 
@@
 % While there certainly is noise in our labels, we trust the artists to have selected the most appropriate set of genres for their own songs.
+Genre recognition problems of increasing difficulty are:
+\begin{enumerate}
+	\item Single top genre prediction on the small subset.
+	\item Single top genre prediction on the medium subset.
+	\item Multiple top genres prediction on the large / full set.
+	\item Multiple genres prediction on the large / full set.
+\end{enumerate}
 
 %\subsection{Methods}
-
 %\subsection{Evaluation and results}
 
-\tabref{tab:mgr} reports classification accuracies obtained on the \todo{small} subset with \todo{15} different mainstream feature sets and 5 standard classifiers using scikit-learn \cite{scikit-learn}, version 0.18.1.
-Specifically, we employed \todo{Naive Bayes, Support Vector Machines (polynomial kernel with exponent 1), k-nearest Neighbours with $k=1$, a J48 Decision Tree, and Random Forests}.
+\tabref{tab:mgr} reports classification accuracies for problem 2 with 9 different mainstream feature sets and some combinations as well as 4 standard classifiers using scikit-learn \cite{scikit-learn}, version 0.18.1.
+Specifically, we employed linear regression (LR) with an $L^2$ penalty, k-nearest neighbors (kNN) with $k=200$, support vector machines (SVM) with a radial basis function (RBF) kernel and a multilayer perceptron (MLP) with 100 hidden neurons.
 All classifiers were tested with mostly default settings.\footnote{See \texttt{baselines.ipynb} for all details.} Reported performances should not be taken as the state-art-of-the-art but rather as a lower-bound and an indication of the task's difficulty.
@@
 	\centering
-	\begin{tabular}{lrrrrrr}
+	\begin{tabular}{l@{\tspaceb}r@{\tspaceb}c@{\tspaceb}c@{\tspaceb}c@{\tspaceb}c}
 		\toprule
-		Feature set & Dim. & NB & SVM & kNN & DT & RF \\
+		feature set & dim. & LR & kNN & SVM & MLP \\
 		\midrule
-		1 Chroma \cite{chroma} & 84 \\
-		2 Tonnetz \cite{tonnetz} & 42 \\
-		3 MFCC \cite{mfcc} & 140 \\
-		4 Spec. centroid & 7 \\
-		5 Spec. bandwidth & 7 \\
-		6 Spec. contrast \cite{contrast} & 49 \\
-		7 Spec. rolloff & 7 \\
-		8 RMSE & 7 \\ % Root-mean-square energy
-		9 Zero-crossing rate & 7 \\
+		1 Chroma \cite{chroma} & 84 & 44 & 44 & 48 & 49 \\
+		2 Tonnetz \cite{tonnetz} & 42 & 40 & 37 & 42 & 41 \\
+		3 MFCC \cite{mfcc} & 140 & 58 & 55 & 61 & 53 \\
+		4 Spec. centroid & 7 & 42 & 45 & 46 & 48 \\
+		5 Spec. bandwidth & 7 & 41 & 45 & 44 & 45\\
+		6 Spec. contrast \cite{contrast} & 49 & 51 & 50 & 54 & 53 \\
+		7 Spec. rolloff & 7 & 42 & 46 & 48 & 48 \\
+		8 RMSE & 7 & 37 & 39 & 39 & 39 \\ % Root-mean-square energy
+		9 Zero-crossing rate & 7 & 42 & 45 & 45 & 46 \\
 		\midrule
-		Ech. audio & 8 \\
-		Ech. social & 5 \\
-		Ech. temporal \cite{echonest_temporal} & 224 \\
+		Ech. audio & 8 & - & - & - & - \\
+		Ech. social & 5 & - & - & - & - \\
+		Ech. temporal \cite{echonest_temporal} & 224 & - & - & - & - \\
 		\midrule
 		% Number and use them for aggregates
-		3 + 6 & 189 \\
-		3 + 6 + 1 & 273 \\
-		All & 150 \\
+		3 + 6 & 189 & 60 & 55 & 63 & 54 \\
+		3 + 6 + 4 & 273 & 60 & 55 & 63 & 53 \\
+		1 to 9 & 518 & 61 & 52 & 63 & 58 \\
 		\bottomrule
 	\end{tabular}
-	\caption{Genre recognition results for various features and classifiers on the FMA \todo{small} subset.}
+	\caption{Accuracies of various features and classifiers for top genre recognition on the medium subset.}
 	\label{tab:mgr}
@@
 
+% Future work
+In addition to the proposed usage and many others people will find, future work on the dataset itself may consider to scrap the website for information not available through the API, e.g.
+the number of downloads or the user comments.
+Moreover, cross-referencing with other resources or crowd-sourcing (with e.g.
+\href{https://www.mturk.com}{Mechanical Turk} or \href{https://www.crowdflower.com/}{CrowdFlower}) is desirable to obtain additional labels or measure agreement on the current ones by independent annotators.
+
+In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{blog post about this dataset}, Cheyenne Hohman, the Director at the Archive, wrote that "by embracing the some rights reserved philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications".
+% Let's hope the research community may contribute back and that we can engage in a mutually benefit.
 % Maybe in some future utopia, those kind of collaborations could shape a future where sharing is first and researchers feed open platforms with their algorithms while they feed us with data.
 That would especially be useful for recommendation systems.
+
+% Foster usage of CC and interplay between artists, platforms, users and researchers.
 
@@
 We want to thank the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
-In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{blog post about this dataset}, Cheyenne Hohman, the Director at the Archive, wrote that "by embracing the some rights reserved philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications".
 We are grateful to SWITCH and EPFL for hosting the dataset within the context of the \href{https://projects.switch.ch/scale-up}{SCALE-UP} project, funded in part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC P-2 program}.

### 2017-04-18 fma.tex
proof-reading, final version sent to co-authors

@@
 \usepackage{enumitem} \setlist{itemsep=0pt,topsep=0pt,parsep=0pt,partopsep=0pt}
+\usepackage{scrextend}
 
@@
 \begin{abstract}
-The Free Music Archive (FMA) is an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR).
-% TODO: which tasks
+The Free Music Archive (FMA) is an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
 While content-based MIR tasks have typically been solved with feature engineering,
@@
 % The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
-By releasing the FMA we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g. genre recognition (MGR).
+By releasing the FMA, we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
 % as it happened in computer vision with the release of ImageNet.
@@
 This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
-% Xavier: tasks, such as ...
 Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.
-%All the code to create,  as well as links to download the data and usage examples are available at \url{https://github.com/mdeff/fma}.
 
@@
 
-% which contains 77,643 songs and 68 genres spanning 26.9 days of song listening
 \end{abstract}
@@
 		% 103-Artists is also known as DB-L, Extended Ballroom
-		\item[1] Names are links to datasets' homepage.
+		\item[1] Names are clickable links to datasets' homepage.
 		\item[2] Audio can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
@@
 		\cmidrule{6-8}
-				&           & \multicolumn{3}{c}{}               & \text{scale} & [GB] & [days] \\
+				&           & \multicolumn{3}{c}{}               & \text{scale} & [GiB] & [days] \\
 		\midrule
@@
 % Value proposition.
-\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with genre.
+\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with musical categories.
 They are also more diverse and allows to average out annotation noise.
@@
 \textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets, which stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
-The sole solution to that problem is to aim for songs which license permits redistribution.
-Therfore all songs chosen for the FMA are licensed under Creative Commons by the artists.
+The FMA embraced the sole solution to that problem, that is to aim for tracks which license permits redistribution, such as Creative Commons.
 All metadata and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{
@@
 They either contain features derived from the audio or provide links to download the audio from an online service.
-The problem with the former is that researchers are stuck with the set of chosen features and prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later  makes no assurance that the files or services won't disappear or change without notice.
+The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later  makes no assurance that the files or services won't disappear or change without notice.
 Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
@@
 Each song is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio.
-Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 songs.
+Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
 Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing.
@@
 
-The collected track, album and artist meta-informations\footnote{Raw data is available in \texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, and \texttt{raw\_artists.csv}.} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
+The collected track, album and artist meta-informations\footnote{Raw data is available in \texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, and \texttt{raw\_artists.csv}.} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
 That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks.
@@
 Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
-\figref{fig:genre_distribution} shows the distribution of the most common genres. As often encountered with natural categories, it obeys a power law.
+\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre. As often encountered with natural categories, it obeys a power law.
 
@@
 
-The upper part of \tabref{tab:features} lists the (mostly spectral) features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0.
+The upper part of \tabref{tab:features} lists the (mostly spectral) features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
 Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples.
@@
 % a number of descriptive medium-level and high-level audio
-features extracted by \href{http://the.echonest.com/}{Echonest}, a music platform, are provided for some tracks.
-These features include high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and MFCC-like temporal features for a number of segments \cite{echonest_temporal}.
+features extracted by \href{http://the.echonest.com/}{Echonest} are provided for some tracks.
+These features (\tabref{tab:features}, lower part) include high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and MFCC-like temporal features for a number of segments \cite{echonest_temporal}.
 They are distributed in \texttt{echonest.csv}.
@@
 
-Finally, for the dataset to be useful as a development set or for people with lower computational resources, we propose the following subsets, each of which is a subset or the larger set:
+Finally, for the dataset to be useful as a development set or for people with lower computational resources, we propose the following subsets, each of which is a subset of the larger set:
 \begin{enumerate}
@@
 All \ngenres genres, unbalanced with 1 to 38,154 tracks per genre (\figref{fig:genre_distribution}) and 0 to 31 genres per track (\figref{fig:genres_per_track}).
-	\item \textbf{Large}: the full dataset with audio limited to 30 seconds clips extracted from the middle of the tracks (or entire song if shorter than 30 seconds).
+	\item \textbf{Large}: the full dataset with audio limited to 30 seconds clips extracted from the middle of the tracks (or entire track if shorter than 30 seconds).
 That trimming reduces the size of the data by a factor 10.
 	\item \textbf{Medium}: 25,000 of the most popular tracks with good metadata and quality audio. Genre unbalanced with 21 to 7,103 tracks per top genre, but only one top genre per track.
-	\item \textbf{Small}: 8,000 tracks from 8 top genres, balanced with 1000 tracks per genre, 1 genre per track.
+	\item \textbf{Small}: 8,000 tracks from 8 top genres, balanced with 1,000 tracks per genre, 1 genre per track.
 This subset is similar to the very popular GTZAN \cite{gtzan} with the added benefits of the FMA, that is metadata, pre-computed features, and copyright-free audio.
@@
 \end{enumerate}
-The above constraints are satisfied for all subsets, and a track is assigned to the same split across all of them. See \texttt{creation.ipynb} for full details.
+The above constraints are satisfied for all subsets, and a track is assigned to the same split across all of them.\footref{creation}
 
@@
 With its rich set of metadata, user data, audio and features, the FMA is amenable to many tasks in MIR.
-We share below some possible uses, which serves to illustrate the breadth of data available in the dataset.
+We share below some possible uses which serve to illustrate the breadth of data available in the dataset.
 
@@
 While the original intention of the dataset was to release a large volume of audio for machine learning algorithms, analyzing metadata from such a collection is also extremely interesting.
-For instance, one could address questions like: How broads artists are, i.e.
+For instance, one could address questions like: How broad artists are, i.e.
 to which genres do they contribute to?
-Do we observe some drift in time for the vocabulary used in song titles, album names or artist names?
+Do we observe some drift in time for the vocabulary used in song titles, album or artist names?
 To which extent tags correspond to genres?
@@
 So far, content-based systems have fallen short at predicting user ratings when compared to collaborative filtering methods.
-The availability of user data such as play counts for tracks and albums as well as number of comments and favorites for tracks, albums and artists opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features, with a clean ground truth.
+The availability of user data such as play counts for tracks and albums (\figref{fig:listens_distribution}) as well as number of comments and favorites for tracks, albums and artists opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features.
+%with a clean ground truth.
 
@@
 
-% Cover song recognition
+% Cover song recognition --> why not, we have track titles
 % Lyrics
@@
 
-While the FMA can be used to evaluate many tasks, metadata is missing for e.g. cover song recognition, mood classification or instrument recognition.
+While the FMA can be used to evaluate many tasks, metadata is missing for e.g.\ mood classification or instrument recognition.
 We can however expect people to cross-reference the dataset with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
 
-Diversity is another issue: as suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental music, electronic and rock.
+Diversity is another issue: as suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental, electronic, and rock music.
 Moreover, it probably does not contain hits from renown artists or commercial music.
@@
 
-\tabref{tab:mgr} reports classification accuracies for problem 2 with 9 different mainstream feature sets and some combinations as well as 4 standard classifiers using scikit-learn \cite{scikit-learn}, version 0.18.1.
+\tabref{tab:mgr} reports classification accuracies for problem 2 with 9 different mainstream feature sets and some combinations as well as 4 standard classifiers using scikit-learn, version 0.18.1 \cite{scikit-learn}.
 Specifically, we employed linear regression (LR) with an $L^2$ penalty, k-nearest neighbors (kNN) with $k=200$, support vector machines (SVM) with a radial basis function (RBF) kernel and a multilayer perceptron (MLP) with 100 hidden neurons.
-All classifiers were tested with mostly default settings.\footnote{See \texttt{baselines.ipynb} for all details.} Reported performances should not be taken as the state-art-of-the-art but rather as a lower-bound and an indication of the task's difficulty.
-Moreover, the produced code can serve as a reference and is easily modified to accommodate other features and classifiers.
+All classifiers were tested with otherwise default settings.\footnote{See \texttt{baselines.ipynb} for all details.} Reported accuracies should not be taken as the state-art-of-the-art but rather as a lower-bound and an indication of the task's difficulty.
+Moreover, the developed code can serve as a reference and is easily modified to accommodate other features and classifiers.
 
@@
 While MGR has been MIR's flagship application for a while, prediction accuracy reached a glass ceiling \cite{mgr_why}.
-A major motivation to construct this dataset was to enable the use of the powerful deep learning set of techniques to music analysis.
+A major motivation to construct this dataset was to enable the use of the powerful DL set of techniques to music analysis.
 %thanks to the advent of specialized DL architectures with testable performances which can be compared.
@@
 		\midrule
-		Ech. audio & 8 & - & - & - & - \\
-		Ech. social & 5 & - & - & - & - \\
-		Ech. temporal \cite{echonest_temporal} & 224 & - & - & - & - \\
-		\midrule
 		% Number and use them for aggregates
@@
 		1 to 9 & 518 & 61 & 52 & 63 & 58 \\
+		\midrule
+		Ech. audio & 8 & - & - & - & - \\
+		Ech. social & 5 & - & - & - & - \\
+		Ech. temporal \cite{echonest_temporal} & 224 & - & - & - & - \\
 		\bottomrule
@@
 The FMA enables researchers to test algorithms on a large-scale collection, closer to real-world environments.
-Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of songs,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset (see \tabref{tab:size}) which opened the door to dramatic performance improvements for many tasks in computer vision.
+Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of songs,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset (\tabref{tab:size}) which opened the door to dramatic performance improvements for many tasks in computer vision.
 %is also far behind the amount of tagged photos \href{https://www.facebook.com}{Facebook} has access to for example.
@@
 % Future work
-In addition to the proposed usage and many others people will find, future work on the dataset itself may consider to scrap the website for information not available through the API, e.g.
-the number of downloads or the user comments.
-Moreover, cross-referencing with other resources or crowd-sourcing (with e.g.
-\href{https://www.mturk.com}{Mechanical Turk} or \href{https://www.crowdflower.com/}{CrowdFlower}) is desirable to obtain additional labels or measure agreement on the current ones by independent annotators.
+In addition to the proposed usage and many others people will find, future work on the dataset itself may consider to scrap the website for information not available through the API, e.g.\ the number of downloads or the user comments.
+Moreover, cross-referencing with other resources or crowd-sourcing (with e.g.\ \href{https://www.mturk.com}{Mechanical Turk} or \href{https://www.crowdflower.com/}{CrowdFlower}) is desirable to obtain additional labels or measure agreement on the current ones by independent annotators.
 
-In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{blog post about this dataset}, Cheyenne Hohman, the Director at the Archive, wrote that "by embracing the some rights reserved philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications".
-% Let's hope the research community may contribute back and that we can engage in a mutually benefit.
-% Maybe in some future utopia, those kind of collaborations could shape a future where sharing is first and researchers feed open platforms with their algorithms while they feed us with data.
-That would especially be useful for recommendation systems.
-
-% Foster usage of CC and interplay between artists, platforms, users and researchers.
+In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{blog post about this dataset}, Cheyenne Hohman, the Director at the Archive, wrote that ''by embracing the some rights reserved philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications``.
+Let's hope for a future where sharing is first and researchers feed open platforms with their algorithms while they feed us with data.
+% Foster usage of CC and tighter interplay / integration between artists, platforms, users and researchers.
 

### 2017-04-18 fma.tex
update

@@
 The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
-It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography and tags; together with full-length and high-quality audio as well as some pre-computed features.
+It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography, tags, and comments; together with full-length and high-quality audio as well as some pre-computed features.
 We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 8,000 songs from 8 major genres, a medium genre-unbalanced dataset of 25,000 songs from 16 genres, and a 90 GiB version of the full set with clips trimmed to 30s.
@@
 		\cmidrule{6-8}
-				&           & \multicolumn{3}{c}{}               & \text{scale} & [GiB] & [days] \\
+				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & [days] \\
 		\midrule
@@
 		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
-		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
+		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a volume of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
 		\item[4] Object detection / classification task. It is the most used dataset for image classification.
@@
 Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images.
-Scale is the number of samples times the dimensionality.
-Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GB, an indication of the quantity of information affected by the size and diversity of the data.
+Volume is the number of samples times the dimensionality.
+Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data.
 The last column is the number of days necessary to listen to the whole available audio.}
@@
 They are also more diverse and allows to average out annotation noise.
-While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (see \tabref{tab:datasets}).
-For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (see \tabref{tab:size}) and shall thus be sufficient for the foreseable future.
+While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (\tabref{tab:datasets}).
+For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (\tabref{tab:size}) and shall thus be sufficient for the foreseable future.
 Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.
@@
 		\cmidrule{5-6}
-		        &       &        &  [s]   & [GB] & [days] \\
+		        &       &        &  [s]   & [GiB] & [days] \\
 		\midrule

### 2017-04-18 fma.tex
edit after Xavier's feedback

@@
 \multauthor
-{Michaël Defferrard \hspace{.6cm} Kirell Benzi \hspace{.6cm} Pierre Vandergheynst \hspace{.6cm} Xavier Bresson$^1$}
+{Michaël Defferrard$^*$ \hspace{.6cm} Kirell Benzi$^*$ \hspace{.6cm} Pierre Vandergheynst$^*$ \hspace{.6cm} Xavier Bresson$^\dagger$}
 {
-	LTS2, EPFL, Lausanne, Switzerland \hspace{1cm}
-	{\small$^1$XB is now at NTU, Singapore. Work conducted when he was at EPFL.}\\
+	$^*$LTS2, EPFL, Lausanne, Switzerland \hspace{0.4cm}
+	$^\dagger$SCSE, NTU, Singapore \hspace{0.4cm}
+	{\small Work conducted when XB was at EPFL.}\\
 	{\tt\small \{michael.defferrard,kirell.benzi,pierre.vandergheynst\}@epfl.ch, xbresson@ntu.edu.sg}
@@
 \begin{abstract}
-The Free Music Archive (FMA) is an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
+We introduce the Free Music Archive (FMA), an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
 While content-based MIR tasks have typically been solved with feature engineering,
@@
 		\item[1] Names are clickable links to datasets' homepage.
-		\item[2] Audio can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
+		\item[2] Audio not directly available, can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
 		\item[3] The 25,863 clips are cut from 5,405 songs.
@@
 		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
-		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a volume of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
+		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
 		\item[4] Object detection / classification task. It is the most used dataset for image classification.
@@
 Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images.
-Volume is the number of samples times the dimensionality.
+Scale is the number of samples times the dimensionality.
 Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data.
@@
 \textbf{High-quality audio.}
-\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger don't.
+\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger do not.
 They either contain features derived from the audio or provide links to download the audio from an online service.
-The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later  makes no assurance that the files or services won't disappear or change without notice.
+The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later  makes no assurance that the files or services will not disappear or change without notice.
 Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
@@
 Each song is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio.
-Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
+Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks \cite{art:MossFMA}.
 Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing.
@@
 		1235 & None & 1235 & Instrumental & 14,938 \\
-		10 & None & 10 & Pop & 13,845 \\
-		17 & None & 17 & Folk & 12,706 \\
+		%10 & None & 10 & Pop & 13,845 \\
+		%17 & None & 17 & Folk & 12,706 \\
 		\midrule
@@
 	\includegraphics[width=\linewidth]{genre_distribution.pdf}
-	\caption{Tracks per genre, from 1 to 38,154.}
+	\\ \vspace{1em}
+	\includegraphics[width=\linewidth]{genre_top_distribution.pdf}
+	\caption{(top) Tracks per genre, from 1 to 38,154.
+	(bottom) Tracks per top genre (medium), from 21 to 7,103.}
 	\label{fig:genre_distribution}
+	\label{fig:genre_top_distribution}
 \end{figure}
@@
 That trimming reduces the size of the data by a factor 10.
-	\item \textbf{Medium}: 25,000 of the most popular tracks with good metadata and quality audio. Genre unbalanced with 21 to 7,103 tracks per top genre, but only one top genre per track.
+	\item \textbf{Medium}: 25,000 of the most popular tracks with good metadata and quality audio.
+Genre unbalanced with 21 to 7,103 tracks per top genre (\figref{fig:genre_top_distribution}), but only one top genre per track.
 	\item \textbf{Small}: 8,000 tracks from 8 top genres, balanced with 1,000 tracks per genre, 1 genre per track.
@@
 % While there certainly is noise in our labels, we trust the artists to have selected the most appropriate set of genres for their own songs.
-Genre recognition problems of increasing difficulty are:
+We consider the following problems of increasing difficulty:
 \begin{enumerate}

### 2017-04-18 fma.tex
reduce size

@@
 	%\item Stratified sampling with a time filter, i.e. for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
-	\item Artists with lots of songs were distributed between the sets. %The three sets were alternatively filled in descending order of the count of artist songs. 
+	\item Artists with lots of tracks were allocated across sets. %The three sets were alternatively filled in descending order of the count of artist songs.
 \end{enumerate}
@@
 
-\tabref{tab:mgr} reports classification accuracies for problem 2 with 9 different mainstream feature sets and some combinations as well as 4 standard classifiers using scikit-learn, version 0.18.1 \cite{scikit-learn}.
+\tabref{tab:mgr} reports accuracies for problem 2 with 9 mainstream feature sets and some combinations as well as 4 standard classifiers using scikit-learn, version 0.18.1 \cite{scikit-learn}.
 Specifically, we employed linear regression (LR) with an $L^2$ penalty, k-nearest neighbors (kNN) with $k=200$, support vector machines (SVM) with a radial basis function (RBF) kernel and a multilayer perceptron (MLP) with 100 hidden neurons.
-All classifiers were tested with otherwise default settings.\footnote{See \texttt{baselines.ipynb} for all details.} Reported accuracies should not be taken as the state-art-of-the-art but rather as a lower-bound and an indication of the task's difficulty.
+All classifiers were tested with otherwise default settings.\footnote{See \texttt{baselines.ipynb} for all details.} Reported performance should not be taken as the state-of-the-art but rather as a lower-bound and an indication of the task's difficulty.
 Moreover, the developed code can serve as a reference and is easily modified to accommodate other features and classifiers.
@@
 
-In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{blog post about this dataset}, Cheyenne Hohman, the Director at the Archive, wrote that ''by embracing the some rights reserved philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications``.
-Let's hope for a future where sharing is first and researchers feed open platforms with their algorithms while they feed us with data.
+In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{post about the dataset}, Cheyenne Hohman, the Director at the Archive, wrote that ''by embracing the
+~\ldots~%some rights reserved
+philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications``.
+Let's hope for a future where sharing is first and researchers feed open platforms with algorithms while they feed us with data.
 % Foster usage of CC and tighter interplay / integration between artists, platforms, users and researchers.

### 2017-04-20 fma.tex
update

@@
 \newcommand{\aduration}{278 }
-\newcommand{\size}{900 }
+\newcommand{\size}{940 }
 
@@
 It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography, tags, and comments; together with full-length and high-quality audio as well as some pre-computed features.
-We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 8,000 songs from 8 major genres, a medium genre-unbalanced dataset of 25,000 songs from 16 genres, and a 90 GiB version of the full set with clips trimmed to 30s.
+We propose a train/validation/test split and three subsets: a small genre-balanced set of 8,000 songs from 8 major genres, a medium genre-unbalanced set of 25,000 songs from 16 genres, and a 94 GiB version of the full set with clips trimmed to 30s.
 This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
@@
 		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
-			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & - & 241 \\
+			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & 776 & 241 \\
 		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
-			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 541 \\
+			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
 		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
@@
 	\begin{tablenotes}
-		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos.
+		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
 		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
@@
 % Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
+% Robin: Alors il télécharge aussi des clips audios de 10 secondes, non-compressé et à 48 kHz pour un total de 4 Terabytes! Donc environ 8x la taille de ton dataset il semble.
+% uncompressed: 2084320*10*48e3*2*2 / 1024**4 = 3.6
+% 320 kbit/s mp3: 2084320*10*320e3 / 8 / 1024**3 = 776
 
@@
 
-\subsection{Content}
-% Figures and tables from analysis.ipynb
-% TODO audio properties: samplerate (% 44kHz), bitrate (% 128, VBR), channels (% mono, stereo)
-
-The collected track, album and artist meta-informations\footnote{Raw data is available in \texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, and \texttt{raw\_artists.csv}.} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
-That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks.
-The audio for each track is stored in a file wich name is the track id.
-All songs are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo.
-Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
-See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.
-
-%\todo{describe the metadata}
-
-\subsection{Genres}
-
 \begin{table}
@@
 
-While it is difficult to devise a hierarchy of genres, we followed that of our source.
-Indeed, the website offers a parent genre for each of the 163 available genres.
-\tabref{tab:genres} shows an excerp of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree.
-\figref{fig:genre_hierarchy} shows an excerpt of the tree.
-In the per-track table, the \texttt{genres} column contain the genre ids indicated by the artist.
-Then, given such hierarchical information, we constructed a \texttt{genres\_all} column which contains all the genres encountered when traversing the tree from the indicated genres to the roots.
-Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
-\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre. As often encountered with natural categories, it obeys a power law.
-
-% MSD:
-% not strictly genres (e.g. one of them is romantic dinner music)
-% one non-music category
-
 \begin{figure}
@@
 
+\subsection{Content}
+% Figures and tables from analysis.ipynb
+% TODO audio properties: samplerate (% 44kHz), bitrate (% 128, VBR), channels (% mono, stereo)
+
+The collected track, album and artist meta-informations\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
+That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks.
+The audio for each track is stored in a file wich name is the track id.
+All songs are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo.
+Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
+See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.
+
+%\todo{describe the metadata}
+
+\subsection{Genres}
+
+While it is difficult to devise a hierarchy of genres, we followed that of our source.
+Indeed, the website offers a parent genre for each of the 163 available genres.
+\tabref{tab:genres} shows an excerp of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree.
+\figref{fig:genre_hierarchy} shows an excerpt of the tree.
+In the per-track table, the \texttt{genres} column contain the genre ids indicated by the artist.
+Then, given such hierarchical information, we constructed a \texttt{genres\_all} column which contains all the genres encountered when traversing the tree from the indicated genres to the roots.
+Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
+\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre. As often encountered with natural categories, it obeys a power law.
+
+% MSD:
+% not strictly genres (e.g. one of them is romantic dinner music)
+% one non-music category
+
 \begin{figure}[t]
@@
 	\caption{(top) Tracks per genre, from 1 to 38,154.
-	(bottom) Tracks per top genre (medium), from 21 to 7,103.}
+	(bottom) Tracks per top genre (medium), from 21 to 7,103.
+	Experimental music is less represented in the curated set.}
 	\label{fig:genre_distribution}
@@
 % a number of descriptive medium-level and high-level audio
-features extracted by \href{http://the.echonest.com/}{Echonest} are provided for some tracks.
+features extracted by \href{http://the.echonest.com/}{Echonest} are provided for 13,129 tracks.
 These features (\tabref{tab:features}, lower part) include high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and MFCC-like temporal features for a number of segments \cite{echonest_temporal}.
@@
 
-Finally, for the dataset to be useful as a development set or for people with lower computational resources, we propose the following subsets, each of which is a subset of the larger set:
+For the dataset to be useful as a development set or for people with lower computational resources, we propose the following sets, each of which is a subset of the larger set:
 \begin{enumerate}
@@
 		\midrule
-		small  &    8,000 &   8 &  30 & 3.4  & 2.8  \\
-		medium &   25,000 &  16 &  30 & 12.2 & 8.7  \\
-		large  & \ntracks & 161 &  30 & 90   & 37 \\
+		small  &    8,000 &   8 &  30 &  7.2 & 2.8  \\
+		medium &   25,000 &  16 &  30 & 23   & 8.7  \\
+		large  & \ntracks & 161 &  30 & 94   & 37 \\
 		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
@@
 % Allmusic: album reviews, artist biographies, discographies as well as classification of albums according to genres, styles, moods and themes. Information is provided and curated by experts.
+% playlists: Art of the Mix playlist categories
 

### 2017-04-21 fma.tex
anonymous initial submission to ISMIR

@@
 \multauthor
-{Michaël Defferrard$^*$ \hspace{.6cm} Kirell Benzi$^*$ \hspace{.6cm} Pierre Vandergheynst$^*$ \hspace{.6cm} Xavier Bresson$^\dagger$}
-{
+{\phantom{Michaël Defferrard$^*$ \hspace{.6cm} Kirell Benzi$^*$ \hspace{.6cm} Pierre Vandergheynst$^*$ \hspace{.6cm} Xavier Bresson$^\dagger$}}
+{\phantom{
 	$^*$LTS2, EPFL, Lausanne, Switzerland \hspace{0.4cm}
 	$^\dagger$SCSE, NTU, Singapore \hspace{0.4cm}
-	{\small Work conducted when XB was at EPFL.}\\
-	{\tt\small \{michael.defferrard,kirell.benzi,pierre.vandergheynst\}@epfl.ch, xbresson@ntu.edu.sg}
+	{\small Work conducted when XB was at EPFL.}}\\
+	\phantom{\tt\small \{michael.defferrard,kirell.benzi,pierre.vandergheynst\}@epfl.ch, xbresson@ntu.edu.sg}
   }
@@
 %% To make customize author list in Creative Common license, uncomment and customize the next line
-\def\authorname{Kirell Benzi, Michaël Defferrard, Pierre Vandergheynst, Xavier Bresson}
+\def\authorname{\phantom{Kirell Benzi, Michaël Defferrard, Pierre Vandergheynst, Xavier Bresson}}
 
@@
 This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
-Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.
+Code, data, and usage examples are available at \url{https://github.com/anonymous}.
 
@@
 		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
-		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
+		\bf \href{https://github.com/anonymous/}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
 		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
@@
 			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
-		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
+		\href{https://github.com/anonymous/}{FMA} \tnote{2} &
 			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
@@
 
+\phantom{
 We want to thank the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
@@
 Xavier Bresson is supported by NRF Fellowship NRFF2017-10.
+}
 

### 2017-04-22 fma.tex
update

--- 2017-04-20 fma.tex
@@
 \newcommand{\aduration}{278 }
-\newcommand{\size}{940 }
+\newcommand{\size}{917 }
+\newcommand{\weblink}{https://github.com/mdeff/fma}
 
@@
 \begin{abstract}
-We introduce the Free Music Archive (FMA), an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
+We introduce the Free Music Archive (FMA), an open and easily accessible dataset which can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
 While content-based MIR tasks have typically been solved with feature engineering,
@@
 The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
-It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography, tags, and comments; together with full-length and high-quality audio as well as some pre-computed features.
-We propose a train/validation/test split and three subsets: a small genre-balanced set of 8,000 songs from 8 major genres, a medium genre-unbalanced set of 25,000 songs from 16 genres, and a 94 GiB version of the full set with clips trimmed to 30s.
+It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts, favorites, and comments; free-form text such as description, biography, and tags; together with full-length and high-quality audio as well as some pre-computed features.
+We propose a train/validation/test split and three subsets: a small genre-balanced set of 8,000 tracks from 8 major genres, a medium genre-unbalanced set of 25,000 tracks from 16 genres, and a 94 GiB version of the full set with clips trimmed to 30s.
 This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
-Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.
+Code, data, and usage examples are available at \url{\weblink}.
 
 % Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of metadata for artists analysis or song ratings for recommender systems.
-Existing datasets such as GTZAN, TagATune, and Million Song suffer from the previous limitations.
+Existing datasets such as GTZAN, MagnaTagATune, and Million Song suffer from the previous limitations.
 It is however essential to establish such benchmark datasets to advance the field of music analysis, like the ImageNet dataset which made possible the large success of deep learning techniques in computer vision.
@@
 		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
-		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 729 & - & 2004 & yes \\
 		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
@@
 		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
+		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 1,458 & - & 2004 & yes \\
 		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
@@
 		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
-		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
+		\bf \href{\weblink}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
 		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
@@
 		\item[3] The 25,863 clips are cut from 5,405 songs.
-		\item[4] Low quality 16 kHz, 32kbps, mono mp3.
+		\item[4] Low quality 16 kHz, 32 kbit/s, mono mp3.
 	\end{tablenotes}
@@
 		\cmidrule{6-8}
-				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & [days] \\
+				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & \#days \\
 		\midrule
-		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} &
-			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\ % MagnaTagATune
+		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} &
+			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\
 		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
@@
 			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
-		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
+		\href{\weblink}{FMA} \tnote{2} &
 			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
@@
 Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images.
-Scale is the number of samples times the dimensionality.
-Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data.
+Volume is the number of samples times the dimensionality.
+Size is for a directory containing all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data.
 The last column is the number of days necessary to listen to the whole available audio.}
@@
 Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
-Finally, released or available to download audio is usually clips of 10 to 30 seconds, sometimes of low quality.
-In comparison, FMA is distributed with legally redistributable full-length, high-quality and non-modified audio.
+Finally, distributed or downloadable audio are usually clips of 10 to 30 seconds, sometimes of low quality, e.g.\ 32 kbit/s for MagnaTagATune or an average of 104 kbit/s for MSD \cite{msd_features}.
+In comparison, FMA comes with legally redistributable full-length, high-quality (263 kbit/s on average) and non-modified audio.
 % \todo{See Table ? for a quantitative analysis of quality.}
-% TODO table quality
+% Todo table quality
 
 \textbf{Complete.} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}.
-Basically all the metadata available through the API has been archived.
-In addition to the basics like title, album or artist names, this includes per-song genres and user data such as per-track/album/artist favorites, play counts and comments.
+All the metadata available through the API has been archived.
+In addition to the basics like title, album or artist names, this includes per-track genres and user data such as per-track/album/artist favorites, play counts and comments.
 % which are necessary to evaluate content-based recommender systems.
-There is also free-form text such as per-track/album/artist tags, album description and artist biography.
+There is also free-form text such as per-track/album/artist tags, album description and artist biography. To compare, the MSD only provides artist-level metadata and tags \cite{msd}.
 % social such as artist location or website
 % images for tracks, albums and artists
-% tags are scarce.
 
@@
 %\todo{have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}}.
-Doing so alleviates the risks of tracks to become unavailable, to change without notice or for online services to shut down.
-Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) generate the subsets and splits, (iv) compute the features and (v) test the baselines, so that it can be reproduced or extended.
+Doing so alleviates the risks of songs to become unavailable, to change without notice or for online services to shut down.
+Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) generate the subsets and splits, (iv) compute the features and (v) test the baselines.
+The developed code can serve as a starting point for researchers to compute their own features or evaluate their methods.
 Finally, anybody can recreate or extend the collection, thanks to public songs and APIs.
@@
 The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
-The website provides a large catalog of artists and high-quality mp3-encoded songs, hand-picked by established audio curators.
-Each song is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio.
-Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks \cite{art:MossFMA}.
+The website provides a large catalog of artists and high-quality mp3-encoded tracks, hand-picked by established audio curators.
+Each track is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio.
+Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
 Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing.
@@
 Out of all collected track ids, 180 mp3s could not be downloaded, 286 could not be trimmed by ffmpeg, and features could not be extracted from 71.
-Finally, the license of 2,616 songs prohibited their redistribution, leaving us with \ntracks tracks.
+Finally, the license of 2,616 tracks prohibited their redistribution, leaving us with \ntracks tracks.
 
@@
 The audio for each track is stored in a file wich name is the track id.
-All songs are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo.
+All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo.
 Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
@@
 	\includegraphics[width=\linewidth]{genre_top_distribution.pdf}
-	\caption{(top) Tracks per genre, from 1 to 38,154.
-	(bottom) Tracks per top genre (medium), from 21 to 7,103.
-	Experimental music is less represented in the curated set.}
+	\caption{(top) Tracks per genre on the full set, from 1 to 38,154.
+	(bottom) Tracks per all top genre on the medium subset, from 21 to 7,103.
+	Note how experimental music is much less represented in the curated medium subset.}
 	\label{fig:genre_distribution}
@@
 
-The upper part of \tabref{tab:features} lists the (mostly spectral) features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
+\tabref{tab:features} (upper part) lists the features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
 Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples.
@@
 % PTB 19/3/3, MNIST 60/10, ImageNet 1200/50/100 --> ratio ~6-12
-An 80/10/10\% split into training, validation and test sets is proposed to make research on the FMA reproducible. The training and validation sets should be merged if cross-validation is used instead.
-Below are the constraints we followed:
+We propose an 80/10/10\% split into training, validation and test sets to make research on the FMA reproducible. Training and validation shall be merged if cross-validation is used instead.
+Below are the followed constraints:
 % Data was stratified according to the following criteria: ground truth, artist and times filters
@@
 		\cmidrule{5-6}
-		        &       &        &  [s]   & [GiB] & [days] \\
+		        &       &        &  [s]   & [GiB] & \#days \\
 		\midrule
-		small  &    8,000 &   8 &  30 &  7.2 & 2.8  \\
+		small  &    8,000 &   8 &  30 &  7.4 & 2.8  \\
 		medium &   25,000 &  16 &  30 & 23   & 8.7  \\
-		large  & \ntracks & 161 &  30 & 94   & 37 \\
+		large  & \ntracks & 161 &  30 & 98   & 37 \\
 		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
@@
 On the other hand, different music types are managed more effectively and efficiently once they are categorized into different groups \cite{mir_review_classif}.
-The classification tasks which can be evaluated on the available labels for the FMA include genre recognition, artist identification, year prediction, and music annotation.
-Automatic tagging is a classification problem which cover different semantic categories, where tags are labels which can be any musical term that describes the genre, mood, instrumentation, and style of the song.
+The classification tasks which can readily be evaluated on FMA include genre recognition, artist identification, year prediction, and music annotation.
+Automatic tagging is a classification problem which covers different semantic categories, where tags are labels which can be any musical term that describes the genre, mood, instrumentation, and style of the song.
 %The task of music annotation has recently gained much popularity in the MIR community.
@@
 \figref{fig:tag_distribution} shows their distribution.\footnote{One of the artist tag is often the artist name, which we subtracted in our counts.}
+A more thorough investigation of the available tags may reveal the feasibility of mood classification or instrument recognition.
+Similarly, cover song detection may be doable if multiple versions of many songs are featured.
 
@@
 
-% Cover song recognition --> why not, we have track titles
+% Cover song detection / recognition --> why not, we have track titles
 % Lyrics
@@
 % Music classification
-	% mood classification,
-	% instrument recognition / identification
+	% mood classification --> may be in tags
+	% instrument recognition / identification --> may be in tags
 	% music annotation --> temporal labels, e.g. instrument / note
@@
 % While there certainly is noise in our labels, we trust the artists to have selected the most appropriate set of genres for their own songs.
-We consider the following problems of increasing difficulty:
+We propose the following problems of increasing difficulty:
 \begin{enumerate}
@@
 
-\tabref{tab:mgr} reports accuracies for problem 2 with 9 mainstream feature sets and some combinations as well as 4 standard classifiers using scikit-learn, version 0.18.1 \cite{scikit-learn}.
+\tabref{tab:mgr} reports accuracies for problem 2 with nine mainstream feature sets and some combinations as well as four standard classifiers using scikit-learn, version 0.18.1 \cite{scikit-learn}.
 Specifically, we employed linear regression (LR) with an $L^2$ penalty, k-nearest neighbors (kNN) with $k=200$, support vector machines (SVM) with a radial basis function (RBF) kernel and a multilayer perceptron (MLP) with 100 hidden neurons.
@@
 		7 Spec. rolloff & 7 & 42 & 46 & 48 & 48 \\
-		8 RMSE & 7 & 37 & 39 & 39 & 39 \\ % Root-mean-square energy
+		8 RMS energy & 7 & 37 & 39 & 39 & 39 \\ % Root-mean-square energy
 		9 Zero-crossing rate & 7 & 42 & 45 & 45 & 46 \\
@@
 The FMA enables researchers to test algorithms on a large-scale collection, closer to real-world environments.
-Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of songs,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset (\tabref{tab:size}) which opened the door to dramatic performance improvements for many tasks in computer vision.
+Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of tracks,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset (\tabref{tab:size}) which opened the door to dramatic performance improvements for many tasks in computer vision.
 %is also far behind the amount of tagged photos \href{https://www.facebook.com}{Facebook} has access to for example.

### 2017-04-24 fma.tex
update

@@
 
+% ISMIR 199 words abstract:
+% We introduce the Free Music Archive (FMA), an open and easily accessible dataset which can be used to evaluate several tasks in music information retrieval (MIR).
+The community's growing interest in feature and end-to-end learning is however restrained by the limited availability of large audio datasets.
+By releasing the FMA, we hope to foster research which will improve the state-of-the-art and hopefully surpass the performance ceiling observed in e.g.
+genre recognition (MGR).
+The data is made of 106,574 tracks, 16,341 artists, 14,854 albums, arranged in a taxonomy of 161 genres, for a total of 343 days of audio and 917 GiB.
+It features metadata like song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags; together with full-length, high-quality audio, and some pre-computed features.
+We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with clips trimmed to 30s.
+This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
+Code, data, and usage examples are available at [anonymous].
+
 % about 150-200 words.
 \begin{abstract}
-We introduce the Free Music Archive (FMA), an open and easily accessible dataset which can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
-While content-based MIR tasks have typically been solved with feature engineering,
+We introduce the Free Music Archive (FMA), an open and easily accessible dataset which can be used to evaluate several tasks in music information retrieval (MIR), a field concerned with browsing, searching, and organizing large music collections.
+%While content-based MIR tasks have typically been solved with feature engineering,
 %with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
-the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
-That endeavor is however restrained by the limited availability of large audio datasets.
+%the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
+%That endeavor is however restrained by the limited availability of large audio datasets.
+The community's growing interest in feature and end-to-end learning is however restrained by the limited availability of large audio datasets.
 % The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
-By releasing the FMA, we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
+By releasing the FMA, we hope to foster research which will improve the state-of-the-art and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
 % as it happened in computer vision with the release of ImageNet.
-The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
-It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts, favorites, and comments; free-form text such as description, biography, and tags; together with full-length and high-quality audio as well as some pre-computed features.
-We propose a train/validation/test split and three subsets: a small genre-balanced set of 8,000 tracks from 8 major genres, a medium genre-unbalanced set of 25,000 tracks from 16 genres, and a 94 GiB version of the full set with clips trimmed to 30s.
-This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
+The data is made of \ntracks tracks, \nartists artists, \nalbums albums, arranged in a hierarchical taxonomy of \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
+It features metadata like song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags; together with full-length, high-quality audio, and some pre-computed features.
+We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 major genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with clips trimmed to 30s.
+This paper describes the dataset and how it was created, proposes some tasks like music classification and annotation or recommendation, and evaluates some baselines for MGR.
 Code, data, and usage examples are available at \url{\weblink}.
@@
 It is however essential to establish such benchmark datasets to advance the field of music analysis, like the ImageNet dataset which made possible the large success of deep learning techniques in computer vision.
-
 \end{abstract}
@@
 
-While it is difficult to devise a hierarchy of genres, we followed that of our source.
-Indeed, the website offers a parent genre for each of the 163 available genres.
-\tabref{tab:genres} shows an excerp of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree.
+While there is no agreement on a taxonomy of genres \cite{mir_review_genre}, we followed the hierarchy used by our source.
+Indeed, the website offers a parent genre for each of their 163 genres.
+\tabref{tab:genres} shows an excerpt of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree.
 \figref{fig:genre_hierarchy} shows an excerpt of the tree.
@@
 
-\tabref{tab:features} (upper part) lists the features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
+\tabref{tab:features} lists the features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
 Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples.
@@
 features extracted by \href{http://the.echonest.com/}{Echonest} are provided for 13,129 tracks.
-These features (\tabref{tab:features}, lower part) include high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and MFCC-like temporal features for a number of segments \cite{echonest_temporal}.
+These features include 8 high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), 5 social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and 224 MFCC-like temporal features for a number of segments \cite{echonest_temporal}.
 They are distributed in \texttt{echonest.csv}.
@@
 \begin{enumerate}
-	\item Stratified sampling was used to preserve the percentage of tracks per genre. This is important for minority genres.
-	\item An artist filter was used, i.e. an artist was forced to appear in one set only. %All songs from a given artist either appear in train, validation or test, never in two of them.
+	\item Stratified sampling to preserve the percentage of tracks per genre (important for minority genres).
+	\item An artist filter for artists to be part of one set only. %All songs from a given artist either appear in train, validation or test, never in two of them.
 	%\item Stratified sampling with a time filter, i.e. for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Proposed usage} % tasks, usage, problems
+\section{Proposed Usage} % tasks, usage, problems
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	% Artist identification / recognition
+	% Singer / composer recognition
 	% Year prediction / recognition
 % Automatic (music) tagging / autotagging --> tag prediction
-
-%On an other hand, unsupervised learning methods could be used for music categorization. Does genre emerge? How?
 
@@
 
-\subsection{Metadata Analysis}
+\subsection{Data Analysis}
+% Musicology?
 
@@
 Do we observe some drift in time for the vocabulary used in song titles, album or artist names?
-To which extent tags correspond to genres?
-What other semantic categories can we found in them?
-Other lines of work are to test algorithms for the imputation of missing data, or to study to what extent is the FMA representative of mainstream music, or how similar it is to other datasets.
+Which semantic categories can we found in tags and to what extent?
+How are tags correlated (e.g.
+rock \& guitar, disco \& violin)?
+Other lines of work are to test algorithms for the imputation of missing data, or to study to what extent is the FMA representative of mainstream music and its similarity to other datasets.
+Moreover, such vast amount of audio is certainly of interest to musicologists.
 
@@
 Automatic tagging is a classification problem which covers different semantic categories, where tags are labels which can be any musical term that describes the genre, mood, instrumentation, and style of the song.
+It helps to convert the music retrieval problem to text retrieval by substituting songs with tags.
 %The task of music annotation has recently gained much popularity in the MIR community.
 In the FMA, tags can be found at the track, album and artist level: 23,496 tracks (22\%), 2,659 albums (18\%) and 1,605 artists (10\%) have at least one tag.
-\figref{fig:tag_distribution} shows their distribution.\footnote{One of the artist tag is often the artist name, which we subtracted in our counts.}
-A more thorough investigation of the available tags may reveal the feasibility of mood classification or instrument recognition.
-Similarly, cover song detection may be doable if multiple versions of many songs are featured.
+\figref{fig:tag_distribution} shows their distribution.\footnote{One of the artist tag is often the artist name.
+We thus subtracted one.}
+In addition to supervised methods which classify music given an arbitrary taxonomy, another approach is to cluster data in an unsupervised way so that a categorization will emerge from the data itself based on objective similarity measures.
+Then, does genre or another taxonomy naturally come up?
 
@@
 % ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
+% genre taxonomy: AllMusic, Amazon, mp3.com
 % musicbrainz, playme, 7digital
@@
 % Music classification
-	% mood classification --> may be in tags
+	% mood classification (happy, sad, and angry) --> may be in tags
 	% instrument recognition / identification --> may be in tags
@@
 While the FMA can be used to evaluate many tasks, metadata is missing for e.g.\ mood classification or instrument recognition.
-We can however expect people to cross-reference the dataset with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
+However, a more thorough investigation of the available tags may reveal their feasibilities.
+Similarly, cover song detection may be doable if multiple versions of many songs are featured.
+Moreover, we can expect the dataset to be cross-referenced with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
 
@@
 Music genres are categories that have arisen through a complex interplay of cultures, artists, and market forces to characterize similarities between compositions and organize music collections.
-Yet, the boundaries between genres still remain fuzzy, making the problem of music genre recognition (MGR) a nontrivial task.
+Yet, the boundaries between genres still remain fuzzy, making the problem of music genre recognition (MGR) a nontrivial task \cite{mir_review_genre}.
 % genre not defined by audio content only
@@
 		1 to 9 & 518 & 61 & 52 & 63 & 58 \\
-		\midrule
-		Ech. audio & 8 & - & - & - & - \\
-		Ech. social & 5 & - & - & - & - \\
-		Ech. temporal \cite{echonest_temporal} & 224 & - & - & - & - \\
+		%\midrule
+		%Ech. audio & 8 & - & - & - & - \\
+		%Ech. social & 5 & - & - & - & - \\
+		%Ech. temporal \cite{echonest_temporal} & 224 & - & - & - & - \\
 		\bottomrule

### 2017-04-25 fma.tex
anonymous ISMIR submission update

@@
 \newcommand{\size}{917 }
-\newcommand{\weblink}{https://github.com/mdeff/fma}
+%\newcommand{\weblink}{https://github.com/mdeff/fma}
+\newcommand{\weblink}{http://164.132.40.229/fma}
 
@@
 This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
-Code, data, and usage examples are available at [anonymous].
+Code, data, and usage examples are available at http://164.132.40.229/fma/.
 

### 2017-04-25 fma.tex
non-anonymous output

@@
 \newcommand{\size}{917 }
-%\newcommand{\weblink}{https://github.com/mdeff/fma}
-\newcommand{\weblink}{http://164.132.40.229/fma}
+\newcommand{\weblink}{https://github.com/mdeff/fma}
+%\newcommand{\weblink}{http://164.132.40.229/fma}
 
@@
 \multauthor
-{\phantom{Michaël Defferrard$^*$ \hspace{.6cm} Kirell Benzi$^*$ \hspace{.6cm} Pierre Vandergheynst$^*$ \hspace{.6cm} Xavier Bresson$^\dagger$}}
-{\phantom{
+{Michaël Defferrard$^*$ \hspace{.6cm} Kirell Benzi$^*$ \hspace{.6cm} Pierre Vandergheynst$^*$ \hspace{.6cm} Xavier Bresson$^\dagger$}
+{
 	$^*$LTS2, EPFL, Lausanne, Switzerland \hspace{0.4cm}
 	$^\dagger$SCSE, NTU, Singapore \hspace{0.4cm}
-	{\small Work conducted when XB was at EPFL.}}\\
-	\phantom{\tt\small \{michael.defferrard,kirell.benzi,pierre.vandergheynst\}@epfl.ch, xbresson@ntu.edu.sg}
+	{\small Work conducted when XB was at EPFL.}\\
+	{\tt\small \{michael.defferrard,kirell.benzi,pierre.vandergheynst\}@epfl.ch, xbresson@ntu.edu.sg}
   }
@@
 
-\phantom{
 We want to thank the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
@@
 Xavier Bresson is supported by NRF Fellowship NRFF2017-10.
-}
 

### 2017-05-09 fma.tex
authorship in attribution
dagger instead of asterix

@@
 \multauthor
-{Michaël Defferrard$^*$ \hspace{.6cm} Kirell Benzi$^*$ \hspace{.6cm} Pierre Vandergheynst$^*$ \hspace{.6cm} Xavier Bresson$^\dagger$}
+{Michaël Defferrard$^\dagger$ \hspace{.6cm} Kirell Benzi$^\dagger$ \hspace{.6cm} Pierre Vandergheynst$^\dagger$ \hspace{.6cm} Xavier Bresson$^\ddagger$}
 {
-	$^*$LTS2, EPFL, Lausanne, Switzerland \hspace{0.4cm}
-	$^\dagger$SCSE, NTU, Singapore \hspace{0.4cm}
+	$^\dagger$LTS2, EPFL, Switzerland \hspace{0.4cm}
+	$^\ddagger$SCSE, NTU, Singapore \hspace{0.4cm}
 	{\small Work conducted when XB was at EPFL.}\\
@@
 %% To make customize author list in Creative Common license, uncomment and customize the next line
-\def\authorname{Kirell Benzi, Michaël Defferrard, Pierre Vandergheynst, Xavier Bresson}
+\def\authorname{Michaël Defferrard, Kirell Benzi, Pierre Vandergheynst, Xavier Bresson}
 

### 2017-07-10 fma.tex
update abstract

@@
 
-% ISMIR 199 words abstract:
-% We introduce the Free Music Archive (FMA), an open and easily accessible dataset which can be used to evaluate several tasks in music information retrieval (MIR).
-The community's growing interest in feature and end-to-end learning is however restrained by the limited availability of large audio datasets.
-By releasing the FMA, we hope to foster research which will improve the state-of-the-art and hopefully surpass the performance ceiling observed in e.g.
-genre recognition (MGR).
-The data is made of 106,574 tracks, 16,341 artists, 14,854 albums, arranged in a taxonomy of 161 genres, for a total of 343 days of audio and 917 GiB.
-It features metadata like song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags; together with full-length, high-quality audio, and some pre-computed features.
-We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with clips trimmed to 30s.
-This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
-Code, data, and usage examples are available at http://164.132.40.229/fma/.
-
 % about 150-200 words.
 \begin{abstract}
-We introduce the Free Music Archive (FMA), an open and easily accessible dataset which can be used to evaluate several tasks in music information retrieval (MIR), a field concerned with browsing, searching, and organizing large music collections.
-%While content-based MIR tasks have typically been solved with feature engineering,
-%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
-%the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
-%That endeavor is however restrained by the limited availability of large audio datasets.
+We introduce the Free Music Archive (FMA), an open and easily accessible dataset suitable for evaluating several tasks in MIR, a field concerned with browsing, searching, and organizing large music collections.
 The community's growing interest in feature and end-to-end learning is however restrained by the limited availability of large audio datasets.
-% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
-By releasing the FMA, we hope to foster research which will improve the state-of-the-art and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
-% as it happened in computer vision with the release of ImageNet.
-The data is made of \ntracks tracks, \nartists artists, \nalbums albums, arranged in a hierarchical taxonomy of \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
-It features metadata like song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags; together with full-length, high-quality audio, and some pre-computed features.
+The FMA aims to overcome this hurdle by providing \size GiB and \tduration days of Creative Commons-licensed audio from \ntracks tracks from \nartists artists and \nalbums albums, arranged in a hierarchical taxonomy of \ngenres genres.
+It provides metadata like song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags; together with full-length, high-quality audio, and some pre-computed features.
 We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 major genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with clips trimmed to 30s.
-This paper describes the dataset and how it was created, proposes some tasks like music classification and annotation or recommendation, and evaluates some baselines for MGR.
+This paper describes the dataset and how it was created, discuss some suitable MIR tasks, and evaluates some baselines for genre recognition.
 Code, data, and usage examples are available at \url{\weblink}.
-
-% Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of metadata for artists analysis or song ratings for recommender systems.
-Existing datasets such as GTZAN, MagnaTagATune, and Million Song suffer from the previous limitations.
-It is however essential to establish such benchmark datasets to advance the field of music analysis, like the ImageNet dataset which made possible the large success of deep learning techniques in computer vision.
 \end{abstract}

### 2017-07-12 fma.tex
update reviewer 1

@@
 \usepackage{graphicx} \graphicspath{{figs/}}
+\usepackage[font=small,skip=8pt]{caption}
 \usepackage{booktabs, threeparttable, tabularx}
@@
 }
-
-%\multauthor
-%{$^1$ \hspace{1cm} Second author$^1$ \hspace{1cm} Third author$^2$} { \bfseries{Fourth author$^3$ \hspace{1cm} Fifth author$^2$ \hspace{1cm} Sixth author$^1$}\\
-%  $^1$ Department of Computer Science, University , Country\\
-%$^2$ International Laboratories, City, Country\\
-%$^3$  Company, Address\\
-%{\tt\small CorrespondenceAuthor@ismir.edu, PossibleOtherAuthor@ismir.edu}
-%}
-%\def\authorname{First author, Second author, Third author, Fourth author, Fifth author, Sixth author}
-
-%\threeauthors
-%  {First Author} {Affiliation1 \\ {\tt author1@ismir.edu}}
-%  {Second Author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
-%  {Third Author} {Affiliation3 \\ {\tt author3@ismir.edu}}
-
-%% To make customize author list in Creative Common license, uncomment and customize the next line
 \def\authorname{Michaël Defferrard, Kirell Benzi, Pierre Vandergheynst, Xavier Bresson}
@@
 It provides metadata like song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags; together with full-length, high-quality audio, and some pre-computed features.
-We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 major genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with clips trimmed to 30s.
+We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 major genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with 30s clips.
 This paper describes the dataset and how it was created, discuss some suitable MIR tasks, and evaluates some baselines for genre recognition.
@@
 % to achieve their purpose, which are sometimes challenging to collect and share.
-In computer vision, the community has developed over the years established benchmark datasets such as \href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist}, \href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR} \cite{cifar}, or \href{http://www.image-net.org}{ImageNet} \cite{imagenet}.
-These datasets are free, legal, easily available, and have proved essential to advance the field.
-The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.3M images \cite{imagenet_challenge}, demonstrated the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best \cite{convnet_imagenet}, and unlocked incredible achievements in both fields \cite{dl}.
+In computer vision, the community has developed established benchmark datasets such as \href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist}, \href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR} \cite{cifar}, or \href{http://www.image-net.org}{ImageNet} \cite{imagenet}, which have proved essential to advance the field.
+The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.3M images \cite{imagenet_challenge}, demonstrated the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best \cite{convnet_imagenet}, and enabled incredible achievements in both fields \cite{dl}.
 
 \begin{table}[t]
+	\small
 	\centering
@@
 		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
-		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
+		\href{https://marsyasweb.appspot.com/download/data_sets/}{GTZAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
 		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
@@
 		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
-		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
+		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
 		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
@@
 		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
+		\href{https://acousticbrainz.org}{AcousticBrainz} \cite{acousticbrainz} & 2,508,627\tnote{5} & - & 2017 & no \\
 		\bottomrule
@@
 		\item[1] Names are clickable links to datasets' homepage.
-		\item[2] Audio not directly available, can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
+		\item[2] Audio not directly available, can be downloaded from \\ \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
 		\item[3] The 25,863 clips are cut from 5,405 songs.
 		\item[4] Low quality 16 kHz, 32 kbit/s, mono mp3.
+		\item[5] As of 2017-07-10, of which a subset has been linked to genre labels for the \href{http://www.multimediaeval.org/mediaeval2017}{MediaEval 2017} \href{https://multimediaeval.github.io/2017-AcousticBrainz-Genre-Task}{genre task}.
 	\end{tablenotes}
 	\end{threeparttable}
-	\caption{Comparison with some other datasets.} % Genre classification datasets.
+	\caption{Comparison between FMA and alternative datasets which can be used for genre recognition.}
 	\label{tab:datasets}
@@
 		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
-		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
+		\item[2] Average duration and sample rate. Most clips from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
 		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
@@
 \begin{table}
+	\small
 	\centering
@@
 	\end{tabular}
-	\caption{List of available per-track, per-album and per-artist metadata, i.e. the columns of \texttt{tracks.csv}.}
+	\caption{List of available per-track, per-album and per-artist metadata, i.e.\ the columns of \texttt{tracks.csv}.}
 	%Not all of them are filled for each track.\footnote{See }
@@
 
-% Besides, the most influential competition in the field of music analysis organized every year is the Music Information Retrieval Evaluation eXchange (MIREX).\href{http://www.music-ir.org/mirex/wiki/}{MIREX} proposes several important music analysis challenges such as song identification, tag classification, music similarity and retrieval, etc.
-However, participants do not have access to neither the test set, nor the important train set.
-They must upload their code to a website that will be evaluated by the organizers.
-In other words, participants cannot train their analysis models on any part of the dataset.
+An alternative to open benchmarking is the approach taken by evaluation challenges such as the music information retrieval evaluation exchange (MIREX) \cite{mirex}.
+The imposed constraint that input data cannot be made available to participants however incurs an approximately linear cost with the number of submissions as all algorithms have to be run by the organizers.
+By releasing this open dataset, we realize part of the vision of McFee et al.\ in ''a distributed, community-centric paradigm for system evaluation, built upon the principles of openness, transparency, and reproducibility`` \cite{mirex_critic}.
 
@@
 \tabref{tab:datasets} lists the most common datasets used for content-based MIR.
-GTZAN \cite{gtzan}, a collection of 1000 songs from 10 genres, was the first publicly available benchmark MGR dataset.
-As a result, despites its flaws \cite{gtzan_critic_1}, it continues to be the most used dataset for MGR \cite{mgr_eval}.
-The main limitations of GTZAN is its illegality, small size, and missing metadata (no user ratings or even artist name and song title).
+GTZAN \cite{gtzan}, a collection of 1000 clips from 10 genres, was the first publicly available benchmark MGR dataset.
+As a result, despites its flaws, it continues to be the most used dataset for MGR \cite{mgr_eval_2}.
+In addition to its small size, the main problems of GTZAN is mislabelling, repetitions, and distortions, as well as missing metadata (no artist name and song title) which prevent researchers to control for artists or album effects.
 % Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
-Looking at \tabref{tab:datasets}, the well-known MagnaTagATune and the Million Song Dataset (MSD) as well as the newer AudioSet appear as contenders for a large-scale reference dataset.
+Looking at \tabref{tab:datasets}, the well-known MagnaTagATune \cite{magnatagatune} and the Million Song Dataset (MSD) \cite{msd} as well as the newer AudioSet \cite{audioset} appear as contenders for a large-scale reference dataset.
 % They however each have limitations.
@@
 The poor audio quality and limited number of songs does however limit its usage.
-MSD and AudioSet, while legal and very large, force researchers to download audio excerpts from online services.
+MSD and AudioSet, while legal and very large, force researchers to download audio clips from online services.
 % Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
@@
 % Value proposition.
+\textbf{Avoiding biases.} sampling bias TODO: reviewer 1 comment 15 \cite{mgr_eval_2}
+
 \textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with musical categories.
@@
 %(4) Free and easy access of the dataset,\\
-%(5) Legality of the dataset, no copyright issue with e.g. open access licence.\\
+%(5) Legality of the dataset, no copyright issue with e.g.\ open access licence.\\
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Dataset} %Introducing the FMA dataset}
+\section{Dataset} % The dataset, dataset, Introducing the FMA dataset
 % Describe what it is, what are the metadata and features.
@@
 	\includegraphics[width=\linewidth]{duration_distribution.pdf}
-	\caption{Track duration, from 0 to 3 hours.}
+	\caption{Track duration, from 0 to 3 hours (zoomed).}
 	\label{fig:duration_distribution}
@@
 	\includegraphics[width=\linewidth]{listens_distribution.pdf}
-	\caption{Album listens, from 0 to 3.6 millions.}
+	\caption{Album listens, from 0 to 3.6 millions (zoomed).}
 	\label{fig:listens_distribution}
@@
 Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing.
-While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis \cite{art:MossFMA}.
+While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}
 
@@
 That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks.
-The audio for each track is stored in a file wich name is the track id.
+For ease of use, we kept all the metadata in a single table despite the redundancy incurred by e.g.\ the fact that all tracks from a given artist share all artist related columns.
+From a practical point of view, the problem is mitigated by compression for storage and those columns can be treated as \textit{categorical variables} to minimize memory usage.
+The audio for each track is stored in a file which name is the track id.
 All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo.
@@
 Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
-\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre. As often encountered with natural categories, it obeys a power law.
+\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre. As often encountered in natural categories, it obeys a power law.
 
 % MSD:
-% not strictly genres (e.g. one of them is romantic dinner music)
+% not strictly genres (e.g.\ one of them is romantic dinner music)
 % one non-music category
@@
 \tabref{tab:features} lists the features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
+These are all the features the library is able to extract.
 Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples.
@@
 %Echonest audio features are described in Kirell's paper
-%Those features can be obtained either by uploading the songs to their website, which are then analyzed and sent back to the user, or simply retrieved from their database if the songs are known to them, i.e.
-they were analyzed before.
+%Those features can be obtained either by uploading the songs to their website, which are then analyzed and sent back to the user, or simply retrieved from their database if the songs are known to them, i.e.\ they were analyzed before.
 All of which can be done through an \href{http://developer.echonest.com}{API}.
@@
 Genre unbalanced with 21 to 7,103 tracks per top genre (\figref{fig:genre_top_distribution}), but only one top genre per track.
+		\todo{Purpose: single genre prediction.}
 	\item \textbf{Small}: 8,000 tracks from 8 top genres, balanced with 1,000 tracks per genre, 1 genre per track.
@@
 	\item Stratified sampling to preserve the percentage of tracks per genre (important for minority genres).
-	\item An artist filter for artists to be part of one set only. %All songs from a given artist either appear in train, validation or test, never in two of them.
-	%\item Stratified sampling with a time filter, i.e. for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
-	\item Artists with lots of tracks were allocated across sets. %The three sets were alternatively filled in descending order of the count of artist songs.
+	\item An artist filter for artists to be part of one set only.
+It has been shown that the use of songs from the same artist in both training and test sets leads to over-optimistic accuracy and may favor some approaches \cite{artist_filters_1, artist_filters_2}.
+%All songs from a given artist either appear in train, validation or test, never in two of them.
+	%\item Stratified sampling with a time filter, i.e.\ for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
+	\item \todo{Artists with lots of tracks were allocated across sets, i.e.\ the three sets were alternatively filled in descending order of the count of artist songs.}
+	\item The 2231 tracks without genre label are assigned to the training set as they might be useful as additional training samples for semi-supervised algorithms.\footnote{These are only found in the full and large sets.}
 \end{enumerate}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Proposed Usage} % tasks, usage, problems
+\section{Usage} % proposed usage, tasks, usage, problems
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{Data Analysis}
-% Musicology?
-
-While the original intention of the dataset was to release a large volume of audio for machine learning algorithms, analyzing metadata from such a collection is also extremely interesting.
-For instance, one could address questions like: How broad artists are, i.e.
-to which genres do they contribute to?
-Do we observe some drift in time for the vocabulary used in song titles, album or artist names?
-Which semantic categories can we found in tags and to what extent?
-How are tags correlated (e.g.
-rock \& guitar, disco \& violin)?
-Other lines of work are to test algorithms for the imputation of missing data, or to study to what extent is the FMA representative of mainstream music and its similarity to other datasets.
-Moreover, such vast amount of audio is certainly of interest to musicologists.
-
-\subsection{Recommendation} % Recommender Systems}
-% Cite Kirell's paper: Song recommendation with non-negative matrix factorization and graph total variation
-
-Due to the potential commercial value of a working system, music recommendation and music similarity are ones of the best studied areas of MIR.
-So far, content-based systems have fallen short at predicting user ratings when compared to collaborative filtering methods.
-The availability of user data such as play counts for tracks and albums (\figref{fig:listens_distribution}) as well as number of comments and favorites for tracks, albums and artists opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features.
-%with a clean ground truth.
-
 \subsection{Music Classification and Annotation}
@@
 On the other hand, different music types are managed more effectively and efficiently once they are categorized into different groups \cite{mir_review_classif}.
-The classification tasks which can readily be evaluated on FMA include genre recognition, artist identification, year prediction, and music annotation.
-Automatic tagging is a classification problem which covers different semantic categories, where tags are labels which can be any musical term that describes the genre, mood, instrumentation, and style of the song.
+The classification tasks which can readily be evaluated on FMA include genre recognition, artist identification, year prediction, and automatic tagging.
+
+Automatic tagging \cite{autotagging} is a classification problem which covers different semantic categories, where tags are labels which can be any musical term that describes the genre, mood, instrumentation, and style of the song.
 It helps to convert the music retrieval problem to text retrieval by substituting songs with tags.
@@
 We thus subtracted one.}
+
 In addition to supervised methods which classify music given an arbitrary taxonomy, another approach is to cluster data in an unsupervised way so that a categorization will emerge from the data itself based on objective similarity measures.
@@
 
-\subsection{Limitations}
-% ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
-% genre taxonomy: AllMusic, Amazon, mp3.com
-% musicbrainz, playme, 7digital
-% Allmusic: album reviews, artist biographies, discographies as well as classification of albums according to genres, styles, moods and themes. Information is provided and curated by experts.
-% playlists: Art of the Mix playlist categories
-
-% Cover song detection / recognition --> why not, we have track titles
-% Lyrics
-% Automatic playlist generation
-% style?
-% Music classification
-	% mood classification (happy, sad, and angry) --> may be in tags
-	% instrument recognition / identification --> may be in tags
-	% music annotation --> temporal labels, e.g. instrument / note
-
-% metadata such as genre, mood, groove, composer, performer, lyrics, meter, chord progressions, instruments present
-
-While the FMA can be used to evaluate many tasks, metadata is missing for e.g.\ mood classification or instrument recognition.
-However, a more thorough investigation of the available tags may reveal their feasibilities.
-Similarly, cover song detection may be doable if multiple versions of many songs are featured.
-Moreover, we can expect the dataset to be cross-referenced with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
-
-Diversity is another issue: as suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental, electronic, and rock music.
-Moreover, it probably does not contain hits from renown artists or commercial music.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Genre Recognition} % MGR, Baselines
+\subsection{Genre Recognition} % MGR, Baselines
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 As such, it is one of the most researched areas of MIR.
-%\subsection{Data}
-The FMA is especially suited for MGR as it features fine genre information, i.e.
-multiple genres associated to individual songs, has a built-in genre hierarchy (\tabref{tab:genres}), and is annotated by the artists themselves.
+
+The FMA is especially suited for MGR as it features fine genre information, i.e.\ multiple (sub-)genres associated to individual songs, has a built-in genre hierarchy (\tabref{tab:genres}), and is annotated by the artists themselves.
+While the artists are the best placed to judge the positioning of their creations, they might be inconsistent and motivated by factors not necessarily objective, such as achieving a higher play count.\footnote{While the content is curated, the curators focus on the musical content not the metadata.} Ideally those labels should be one of many ground truths, to be complemented by crowd-sourcing and experts (from different music metadata websites).
 % While there certainly is noise in our labels, we trust the artists to have selected the most appropriate set of genres for their own songs.
@@
 
-%\subsection{Methods}
-%\subsection{Evaluation and results}
-
 \tabref{tab:mgr} reports accuracies for problem 2 with nine mainstream feature sets and some combinations as well as four standard classifiers using scikit-learn, version 0.18.1 \cite{scikit-learn}.
@@
 
-% As a preliminary result, we trained a convolutional neural network on 3sec-spectrograms, and we select the class by majority voting.
-
-While MGR has been MIR's flagship application for a while, prediction accuracy reached a glass ceiling \cite{mgr_why}.
-A major motivation to construct this dataset was to enable the use of the powerful DL set of techniques to music analysis.
-%thanks to the advent of specialized DL architectures with testable performances which can be compared.
+A major motivation to construct this dataset was to enable the use of the powerful DL set of techniques to music analysis, an hypothesized cause of stagnation on MIREX tasks \cite{mirex_stagnation}.
 With availability of audio, DL architectures such as convolutional neural networks and recurrent neural networks can be applied to the waveform to avoid any feature engineering.
@@
 
+\subsection{Data Analysis}
+% Musicology?
+
+While the original intention of the dataset was to release a large volume of audio for machine learning algorithms, analyzing metadata from such a collection is also extremely interesting.
+For instance, one could address questions like: How broad artists are, i.e.\ to which genres do they contribute to?
+Do we observe some drift in time for the vocabulary used in song titles, album or artist names?
+Which semantic categories can we found in tags and to what extent?
+How are tags correlated (e.g.\ rock \& guitar, disco \& violin)?
+Other lines of work are to test algorithms for the imputation of missing data, or to study to what extent is the FMA representative of mainstream music and its similarity to other datasets.
+Moreover, such vast amount of audio is certainly of interest to musicologists.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\subsection{Limitations}
+% ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
+% genre taxonomy: AllMusic, Amazon, mp3.com
+% musicbrainz, playme, 7digital
+% Allmusic: album reviews, artist biographies, discographies as well as classification of albums according to genres, styles, moods and themes. Information is provided and curated by experts.
+% playlists: Art of the Mix playlist categories
+% AllMusic and Discogs are based on editorial metadata databases maintained by music experts and enthusiasts 
+% Lastfm and Tagtraum are based on collaborative music tagging platforms with large amounts of genre labels provided by their users for music recordings.
+
+% Cover song detection / recognition --> why not, we have track titles
+% Lyrics
+% Automatic playlist generation
+% style?
+% Music classification
+	% mood classification (happy, sad, and angry) --> may be in tags
+	% instrument recognition / identification --> may be in tags
+	% music annotation --> temporal labels, e.g.\ instrument / note
+
+% metadata such as genre, mood, groove, composer, performer, lyrics, meter, chord progressions, instruments present
+
+Music classification is the most straightforward use case for FMA, with MGR being particularly suitable.
+The inclusion of a genre hierarchy makes it specially interesting, as it offers possibilities rarely found in alternative collections.
+
+While the FMA can be used to evaluate many tasks, metadata is missing for e.g.\ mood classification or instrument recognition.
+However, a more thorough investigation of the available tags may reveal their feasibilities.
+Similarly, cover song detection may be doable if multiple versions of many songs are featured.
+Moreover, we can expect the dataset to be cross-referenced with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
+
+Diversity is another issue: as suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental, electronic, and rock music.
+Moreover, it probably does not contain hits from renown artists or commercial music.
+\todo{more discussion after reading Sturm}
+% features some renown artists (wikipedia list)
+
+%The full potential of the FMA has yet to be unlashed.
+
+\subsection{Recommendation} % Recommender Systems}
+% Cite Kirell's paper: Song recommendation with non-negative matrix factorization and graph total variation
+%Cheyenne: We do have lists for each user of which songs, albums and artists a user marked as 'favorites' though.
+%friends form a social networks
+% fans and favorites
+% user mixes
+
+Due to the potential commercial value of a working system, music recommendation and music similarity are among the best studied areas of MIR.
+%So far, content-based systems have fallen short at predicting user ratings when compared to collaborative filtering methods. \todo{ref}
+While not scraped in the present dump, the lists of which songs, albums and artists a user marked as favorites or commented are public.
+Listening and downloading counts are however only available in aggregates\footnote{That information can still be useful to e.g.\ predict hits.} (\figref{fig:listens_distribution}) as most of that activity happens by users that are not logged in.
+Moreover, users can send \textit{friend requests} to each other forming a social network.
+%The wealth of available data opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features.
+%with a clean ground truth.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Future work
+\todo{If the community finds interest in the dataset and validate its use, future work should focus on validating the ground truth and collecting additional information.}
 In addition to the proposed usage and many others people will find, future work on the dataset itself may consider to scrap the website for information not available through the API, e.g.\ the number of downloads or the user comments.
@@
 We want to thank the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
+We want to thank the anonymous ISMIR reviewers for their thorough reviews and many constructive comments which have improved the quality of this work.
 We are grateful to SWITCH and EPFL for hosting the dataset within the context of the \href{https://projects.switch.ch/scale-up}{SCALE-UP} project, funded in part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC P-2 program}.

### 2017-07-13 fma.tex
update reviewer 2

@@
 The FMA aims to overcome this hurdle by providing \size GiB and \tduration days of Creative Commons-licensed audio from \ntracks tracks from \nartists artists and \nalbums albums, arranged in a hierarchical taxonomy of \ngenres genres.
-It provides metadata like song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags; together with full-length, high-quality audio, and some pre-computed features.
+It provides full-length and high-quality audio, pre-computed features, together with track- and user-level metadata, tags, and free-form text such as biographies.
 We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 major genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with 30s clips.
@@
 	\caption{Comparison between the largest public audio and image datasets.
-Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images.
+Dimensionality is (length [s] $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images.
 Volume is the number of samples times the dimensionality.
@@
 	\end{minipage}
-	\caption{(left) Growth of the archive, created in 11/2008. (right) Number of albums released per year, from 1902 to 2017.}
+	\caption{(left) Growth of the archive, created in 11/2008. (right) Number of albums released per year (min 1902, max 2017).}
 	\label{fig:growth}
@@
 	\centering
-	\begin{tabular}{|llll|}
+	\begin{tabular}{|r@{ }l@{\hspace{.8em}}r@{ }l@{\hspace{.8em}}r@{ }l|}
 		\hline
-		track\_id & title & number & \\
-		information & language\_code & license & \\
-		composer & publisher & lyricist & \\
-		genres & genres\_all & genre\_top & \\
-		duration & bit\_rate & interest & \\ %samplerate \\
-		\#listens & \#comments & \#favorites & tags \\
-		date\_created & date\_recorded & & \\
+		100\% & track\_id     & 100\% & title          &  93\% & number \\
+		  2\% & information   &  14\% & language\_code & 100\% & license \\
+		  4\% & composer      &   1\% & publisher      &   1\% & lyricist \\
+		 98\% & genres        &  98\% & genres\_all    &  47\% & genre\_top \\
+		100\% & duration      & 100\% & bit\_rate      & 100\% & interest \\ %samplerate \\
+		100\% & \#listens     &   2\% & \#comments     &  61\% & \#favorites \\
+		100\% & date\_created &   6\% & date\_recorded &  22\% & tags\\
 		\hline
-		album\_id & title & type & \#tracks \\
-		information & engineer & producer & \\
-		\#listens & \#comments & \#favorites & tags \\
-		date\_created & date\_released & & \\
+		100\% & album\_id     & 100\% & title          &     &  \\
+		 94\% & type          &  96\% & \#tracks       &     &  \\
+		 76\% & information   &  16\% & engineer       &  18\% & producer \\
+		 97\% & \#listens     &  12\% & \#comments     &  38\% & \#favorites \\
+		 97\% & date\_created &  64\% & date\_released &  18\% & tags \\
 		\hline
-		artist\_id & name & & \\
-		bio & members & \multicolumn{2}{l|}{associated\_labels} \\
-		website & wikipedia\_page & \multicolumn{2}{l|}{related\_projects} \\
-		location & longitude & latitude & \\
-		& \#comments & \#favorites & tags \\
-		date\_created & active\_year\_end & \multicolumn{2}{l|}{active\_year\_begin} \\
+		100\% & artist\_id    & 100\% & name           &  25\% & members \\
+		 38\% & bio           &   5\% & \multicolumn{2}{@{}l}{associated\_labels} &  \\
+		 43\% & website       &   2\% & wikipedia\_page & &  \\
+		& & 5\% & \multicolumn{2}{@{}l}{related\_projects} & \\
+		 37\% & location      &  23\% & longitude      &  23\% & latitude \\
+		 11\% & \#comments    &  48\% & \#favorites    &  10\% & tags \\
+		 99\% & date\_created &   8\% & active\_year\_begin & & \\
+		& & 2\% & \multicolumn{2}{@{}l}{active\_year\_end} & \\
 		\hline
 	\end{tabular}
-	\caption{List of available per-track, per-album and per-artist metadata, i.e.\ the columns of \texttt{tracks.csv}.}
+	\caption{List of available per-track, per-album and per-artist metadata, i.e.\ the columns of \texttt{tracks.csv}. Percentages indicate coverage over all tracks, albums, and artists.}
 	%Not all of them are filled for each track.\footnote{See }
@@
 % Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
-Looking at \tabref{tab:datasets}, the well-known MagnaTagATune \cite{magnatagatune} and the Million Song Dataset (MSD) \cite{msd} as well as the newer AudioSet \cite{audioset} appear as contenders for a large-scale reference dataset.
+Looking at \tabref{tab:datasets}, the well-known MagnaTagATune \cite{magnatagatune} and the Million Song Dataset (MSD) \cite{msd} as well as the newer AudioSet \cite{audioset} and AcousticBrainz \cite{acousticbrainz} appear as contenders for a large-scale reference dataset.
 % They however each have limitations.
@@
 The poor audio quality and limited number of songs does however limit its usage.
-MSD and AudioSet, while legal and very large, force researchers to download audio clips from online services.
+MSD and AudioSet, while very large, force researchers to download audio clips from online services.
+AcousticBrainz's approach to the copyright issue is to ask the community to upload music descriptors of their tracks.
+Although it is the largest database to date, it will never distribute audio.
 % Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
@@
 The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later makes no assurance that the files or services will not disappear or change without notice.
-Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
+Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest}.
+They have e.g.\ extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
 Finally, distributed or downloadable audio are usually clips of 10 to 30 seconds, sometimes of low quality, e.g.\ 32 kbit/s for MagnaTagATune or an average of 104 kbit/s for MSD \cite{msd_features}.
-In comparison, FMA comes with legally redistributable full-length, high-quality (263 kbit/s on average) and non-modified audio.
+In comparison, FMA comes with CC-licensed, full-length, high-quality and non-modified audio.
 % \todo{See Table ? for a quantitative analysis of quality.}
-% Todo table quality
 
-\textbf{Complete.} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}.
+\textbf{Metadata rich.} \todo{redundant with content} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}.
 All the metadata available through the API has been archived.
@@
 % which are necessary to evaluate content-based recommender systems.
-There is also free-form text such as per-track/album/artist tags, album description and artist biography. To compare, the MSD only provides artist-level metadata and tags \cite{msd}.
+There is also free-form text such as per-track/album/artist tags, album description and artist biography.
+To compare, the MSD only provides artist-level metadata and tags \cite{msd}, while GTZAN offers none at all.
 % social such as artist location or website
@@
 % Todo an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and
-some usage examples in the \texttt{usage.ipynb} Jupyter notebook to quick start using the data.
+some usage examples in the \texttt{usage.ipynb} Jupyter notebook to start using the data quickly.
 
@@
 	\includegraphics[width=\linewidth]{duration_distribution.pdf}
-	\caption{Track duration, from 0 to 3 hours (zoomed).}
+	\caption{Track duration (min 0, max 3 hours).}
 	\label{fig:duration_distribution}
@@
 	\includegraphics[width=\linewidth]{listens_distribution.pdf}
-	\caption{Album listens, from 0 to 3.6 millions (zoomed).}
+	\caption{Album listens (min 0, max 3.6 millions).}
 	\label{fig:listens_distribution}
@@
 
-The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library of high-quality, legal audio downloads directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
-The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
-The website provides a large catalog of artists and high-quality mp3-encoded tracks, hand-picked by established audio curators.
-Each track is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio.
-Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
-Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing.
+The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
+%The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
+The website provides a large catalog of artists and high-quality tracks, hand-picked by established audio curators.
+Each track is legally free to download as artists decided to release their works under a permissive license.
+%Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
+Inspired by \href{https://creativecommons.org/}{Creative Commons} and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing.
 While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}
@@
 % Collection
-As of April 1st 2017, the date of collection, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks).
-See \figref{fig:growth} to visualize the growth of the dataset.
-In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.}.
-Finally, mp3 audio was downloaded over HTTPS.
+As of April 1st 2017, when the dataset was gathered, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks).
+\figref{fig:growth} shows the growth of the dataset.
+In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.} Finally, mp3 audio was downloaded over HTTPS.
 % Cleaning
@@
 Finally, the license of 2,616 tracks prohibited their redistribution, leaving us with \ntracks tracks.
+
+While it may be argued that the dataset should be cleaned, we wanted it to resemble real world data.
+As such, we did not remove tracks which have too many genres, are too long, belong to rare genres, etc.
+Moreover, it is hard to set a threshold and the small number of outliers will not impact much the prediction accuracy.
+Researchers are obviously free to discard any track for training.
 
@@
 Left number is the \texttt{genre\_id}, right number is the number of tracks per genre.
-(right) Number of genres per track, from 0 to 31 (\texttt{genres\_all}).}
+(right) Number of genres per track.}
 	\label{fig:genre_hierarchy}
@@
 
-The collected track, album and artist meta-informations\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
-That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks.
+The metadata collected from the archive\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} was cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
+That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}.
+All the metadata has been produced by the artists when uploading their music.
+While the content is curated, the curators focus on the musical content not the metadata.
+Metadata includes song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags.
+Coverage varies across fields, see \tabref{tab:metadata}.
 For ease of use, we kept all the metadata in a single table despite the redundancy incurred by e.g.\ the fact that all tracks from a given artist share all artist related columns.
 From a practical point of view, the problem is mitigated by compression for storage and those columns can be treated as \textit{categorical variables} to minimize memory usage.
+
+\todo{describe the metadata}
+In the FMA, tags can be found at the track, album and artist level: 23,496 tracks (22\%), 2,659 albums (18\%) and 1,605 artists (10\%) have at least one tag.
+\figref{fig:tag_distribution} shows their distribution.\footnote{One of the artist tag is often the artist name.
+We thus subtracted one.}
+
 The audio for each track is stored in a file which name is the track id.
-All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo.
+All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kbit/s (263 kbit/s on average), and in stereo.
 Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
 See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.
-
-%\todo{describe the metadata}
+%todo: graph for audio quality
 
@@
 
-While there is no agreement on a taxonomy of genres \cite{mir_review_genre}, we followed the hierarchy used by our source.
-Indeed, the website offers a parent genre for each of their 163 genres.
+\todo{The high coverage of genre information and fine-graine... makes the FMA well suited for MGR.}
+Labeling noise is unavoidable
+The FMA is especially suited for MGR as it features fine genre information, i.e.\ multiple (sub-)genres associated to individual tracks, has a built-in genre hierarchy (\tabref{tab:genres}), and is annotated by the artists themselves.
+While the artists are the best placed to judge the positioning of their creations, they might be inconsistent and motivated by factors not necessarily objective, such as achieving a higher play count.
+Ideally those labels should be one of many ground truths, to be complemented by crowd-sourcing and experts (from different music metadata websites).
+
+While there is no agreement on a taxonomy of genres \cite{mir_review_genre}, we followed the hierarchy used by the archive, which is the one the authors had in mind when annotating their tracks.
+That hierarchy is composed of \ngenres genres of which 16 are roots, the others being sub-genres.
 \tabref{tab:genres} shows an excerpt of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree.
 \figref{fig:genre_hierarchy} shows an excerpt of the tree.
+\todo{3 genres limit}
+
 In the per-track table, the \texttt{genres} column contain the genre ids indicated by the artist.
@@
 Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
-\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre. As often encountered in natural categories, it obeys a power law.
+\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre.
+%As often encountered in natural categories, it obeys a power law.
+
+While root genre recognition should be treated as a multi-label problem in general, \figref{fig:genres_per_track} shows that half the tracks have a single root genre.
 
@@
 	\includegraphics[width=\linewidth]{genre_top_distribution.pdf}
-	\caption{(top) Tracks per genre on the full set, from 1 to 38,154.
-	(bottom) Tracks per all top genre on the medium subset, from 21 to 7,103.
+	\caption{(top) Tracks per genre on the full set (min 1, max 38,154).
+(bottom) Tracks per all top genre on the medium subset (min 21, max 7,103).
 Note how experimental music is much less represented in the curated medium subset.}
@@
 That trimming reduces the size of the data by a factor 10.
-	\item \textbf{Medium}: 25,000 of the most popular tracks with good metadata and quality audio.
-Genre unbalanced with 21 to 7,103 tracks per top genre (\figref{fig:genre_top_distribution}), but only one top genre per track.
-		\todo{Purpose: single genre prediction.}
-	\item \textbf{Small}: 8,000 tracks from 8 top genres, balanced with 1,000 tracks per genre, 1 genre per track.
+	\item \textbf{Medium}: 25,000 of the most popular 30s clips with good metadata and quality audio.
+Genre unbalanced with 21 to 7,103 clips per top genre (\figref{fig:genre_top_distribution}), but only one of the 16 top genres per clip.
+		\todo{This subset was constructed for single root genre prediction. The clips were selected based on ... Non-musical genres, experimental?}
+	\item \textbf{Small}: 8,000 30s clips from 8 top genres, balanced with 1,000 clips per genre, 1 genre per clip.
 This subset is similar to the very popular GTZAN \cite{gtzan} with the added benefits of the FMA, that is metadata, pre-computed features, and copyright-free audio.
+\todo{The same process was used to select the top 1000 tracks per genre.
+The selected genres are those with more than 1000 tracks in the medium subset.}
 \end{enumerate}
@@
 	\item Stratified sampling to preserve the percentage of tracks per genre (important for minority genres).
-	\item An artist filter for artists to be part of one set only.
+	\item An artist filter for artists to be part of one set only, thus avoiding any artist and album effect.
 It has been shown that the use of songs from the same artist in both training and test sets leads to over-optimistic accuracy and may favor some approaches \cite{artist_filters_1, artist_filters_2}.
@@
 	%\item Stratified sampling with a time filter, i.e.\ for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
-	\item \todo{Artists with lots of tracks were allocated across sets, i.e.\ the three sets were alternatively filled in descending order of the count of artist songs.}
+	%\item \todo{Artists with lots of tracks are balanced across sets, i.e.\ the three sets were alternatively filled in descending order of the count of artist songs.
+This does not break the above constraint.}
 	\item The 2231 tracks without genre label are assigned to the training set as they might be useful as additional training samples for semi-supervised algorithms.\footnote{These are only found in the full and large sets.}
 \end{enumerate}
-The above constraints are satisfied for all subsets, and a track is assigned to the same split across all of them.\footref{creation}
+The above constraints are satisfied for all subsets, and a track is assigned to the same split across all of them.\footref{creation} Each root genre is guaranteed to be represented in all splits, but the ratio is only exact for the small subset (800/100/100).
+The seven smallest sub-genres (less than 20 tracks) are however not guaranteed to appear in all splits of the full and large sets.
 
@@
 	\includegraphics[width=\linewidth]{tag_distribution.pdf}
-	\caption{Per-track, album and artist tags, from 0 to 150.}
+	\caption{Per-track, album and artist tags (min 0, max 150).}
 	\label{fig:tag_distribution}
@@
 %The task of music annotation has recently gained much popularity in the MIR community.
-In the FMA, tags can be found at the track, album and artist level: 23,496 tracks (22\%), 2,659 albums (18\%) and 1,605 artists (10\%) have at least one tag.
-\figref{fig:tag_distribution} shows their distribution.\footnote{One of the artist tag is often the artist name.
-We thus subtracted one.}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Genre Recognition} % MGR, Baselines
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 As such, it is one of the most researched areas of MIR.
-
-The FMA is especially suited for MGR as it features fine genre information, i.e.\ multiple (sub-)genres associated to individual songs, has a built-in genre hierarchy (\tabref{tab:genres}), and is annotated by the artists themselves.
-While the artists are the best placed to judge the positioning of their creations, they might be inconsistent and motivated by factors not necessarily objective, such as achieving a higher play count.\footnote{While the content is curated, the curators focus on the musical content not the metadata.} Ideally those labels should be one of many ground truths, to be complemented by crowd-sourcing and experts (from different music metadata websites).
-% While there certainly is noise in our labels, we trust the artists to have selected the most appropriate set of genres for their own songs.
 We propose the following problems of increasing difficulty:
@@
 	\end{tabular}
-	\caption{Accuracies of various features and classifiers for top genre recognition on the medium subset.}
+	\caption{Test set accuracies of various features and classifiers for top genre recognition on the medium subset.}
 	\label{tab:mgr}
@@
 % Musicology?
+% metadata: better suited datasets. Focus on audio.
 
-While the original intention of the dataset was to release a large volume of audio for machine learning algorithms, analyzing metadata from such a collection is also extremely interesting.
-For instance, one could address questions like: How broad artists are, i.e.\ to which genres do they contribute to?
+While our intention was to release a large volume of audio for machine learning algorithms, analyzing audio is certainly of interest to musicologists and researchers who want to study relations with higher-level representations.
+Moreover, the metadata is surely a valuable addition to existing datasets\footnote{Like MusicBrainz, AllMusic, Discogs and Last.fm.} for metadata analysis.
+%For instance, one could address questions like: How broad artists are, i.e.\ to which genres do they contribute to?
 Do we observe some drift in time for the vocabulary used in song titles, album or artist names?
-Which semantic categories can we found in tags and to what extent?
+Which semantic categories can we find in tags and to what extent?
 How are tags correlated (e.g.\ rock \& guitar, disco \& violin)?
-Other lines of work are to test algorithms for the imputation of missing data, or to study to what extent is the FMA representative of mainstream music and its similarity to other datasets.
-Moreover, such vast amount of audio is certainly of interest to musicologists.
+%Other lines of work are to test algorithms for the imputation of missing data.
 
@@
 \todo{more discussion after reading Sturm}
+study to what extent is the FMA representative of mainstream music and its similarity to other datasets.
+\todo{McFee}
 % features some renown artists (wikipedia list)
@@
 
-\subsection{Recommendation} % Recommender Systems}
+\subsection{Music Recommendation} % Recommendation, Recommender Systems
 % Cite Kirell's paper: Song recommendation with non-negative matrix factorization and graph total variation
@@
 %So far, content-based systems have fallen short at predicting user ratings when compared to collaborative filtering methods. \todo{ref}
-While not scraped in the present dump, the lists of which songs, albums and artists a user marked as favorites or commented are public.
-Listening and downloading counts are however only available in aggregates\footnote{That information can still be useful to e.g.\ predict hits.} (\figref{fig:listens_distribution}) as most of that activity happens by users that are not logged in.
-Moreover, users can send \textit{friend requests} to each other forming a social network.
+While the present dump only captures listening and downloading counts in aggregates (\figref{fig:listens_distribution})\footnote{That information can be useful to e.g.\ predict hits.}, the lists of which songs, albums and artists a user marked as favorites or commented are public.
+While not public, listening and downloading activities are logged and might be shared after anonymization.\footnote{Private discussion with the administrators.} Moreover, users form a public social network via \textit{friend requests}.
 %The wealth of available data opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features.

### 2017-07-13 fma.tex
update reviewer 3

@@
 % {\bf Available music datasets and their limitations.}
-Unlike other modalities such as images or text for which a wealth of content is available, the lack of a large-scale, complete and easily available dataset for MIR has hindered research on data-heavy models such as DL.
+Unlike the wealth of available visual or textual content, the lack of a large-scale, complete and easily available dataset for MIR has hindered research on data-heavy models such as DL.
 \tabref{tab:datasets} lists the most common datasets used for content-based MIR.
@@
 As a result, despites its flaws, it continues to be the most used dataset for MGR \cite{mgr_eval_2}.
-In addition to its small size, the main problems of GTZAN is mislabelling, repetitions, and distortions, as well as missing metadata (no artist name and song title) which prevent researchers to control for artists or album effects.
+The main problems are mislabeling, repetitions, and distortions.
+Moreover, it is small and misses metadata which prevents researchers to control for artists or album effects.
 % Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
@@
 % They however each have limitations.
-MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes metadata, audio features and raw audio.
+MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes metadata, features and audio.
 The poor audio quality and limited number of songs does however limit its usage.
@@
 % Value proposition.
-\textbf{Avoiding biases.} sampling bias TODO: reviewer 1 comment 15 \cite{mgr_eval_2}
+\textbf{Avoiding biases.} sampling bias TODO: reviewer 1 comment 15 \cite{mgr_eval_2} Also McFee 2016
 
 \textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with musical categories.
-They are also more diverse and allows to average out annotation noise.
+They are also more diverse and allows to average out annotation noise as well as characteristics who might be confounded with the ground truth and exploited by learning algorithms.
 While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (\tabref{tab:datasets}).
-For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (\tabref{tab:size}) and shall thus be sufficient for the foreseable future.
-Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.
+\todo{For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (\tabref{tab:size}). % and shall thus be sufficient for the foreseable future.
+Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.}
 
 \textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets, which stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
-The FMA embraced the sole solution to that problem, that is to aim for tracks which license permits redistribution, such as Creative Commons.
+The FMA's solution is to aim for tracks which license permits redistribution.
 All metadata and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{
@@
 \textbf{High-quality audio.}
-\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger do not.
-They either contain features derived from the audio or provide links to download the audio from an online service.
+\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, most of the larger do not.
+% often regardless of copyrights
+They either only contain features derived from the audio, or provide links to download the audio from an online service.
 The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later makes no assurance that the files or services will not disappear or change without notice.
-Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest}.
-They have e.g.\ extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
+Moreover, researchers should be wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest}.
+% They have e.g.\ extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
 Finally, distributed or downloadable audio are usually clips of 10 to 30 seconds, sometimes of low quality, e.g.\ 32 kbit/s for MagnaTagATune or an average of 104 kbit/s for MSD \cite{msd_features}.
+The problem with clips is that the beginning 30 seconds of tracks may yield different results than the middle or the final 30 seconds, and that researchers may not have control over which part they get.
 In comparison, FMA comes with CC-licensed, full-length, high-quality and non-modified audio.
@@
 
-\textbf{Metadata rich.} \todo{redundant with content} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}.
-All the metadata available through the API has been archived.
-In addition to the basics like title, album or artist names, this includes per-track genres and user data such as per-track/album/artist favorites, play counts and comments.
-% which are necessary to evaluate content-based recommender systems.
-There is also free-form text such as per-track/album/artist tags, album description and artist biography.
-To compare, the MSD only provides artist-level metadata and tags \cite{msd}, while GTZAN offers none at all.
+\textbf{Metadata rich.} In addition to audio, the dataset comes with rich metadata, shown in \tabref{tab:metadata}.
+While not complete, it compares favorably with the MSD which only provides artist-level metadata \cite{msd} or GTZAN which offers none.
 % social such as artist location or website
@@
 %The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
-The website provides a large catalog of artists and high-quality tracks, hand-picked by established audio curators.
-Each track is legally free to download as artists decided to release their works under a permissive license.
 %Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
-Inspired by \href{https://creativecommons.org/}{Creative Commons} and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing.
+Inspired by \href{https://creativecommons.org/}{Creative Commons} and the open-source software movement, the FMA provides a platform for curators, artists, and listeners to harness the potential of music sharing.
+The website provides a large catalog of artists and tracks, hand-picked by established audio curators.
+Each track is legally free to download as artists decided to release their works under permissive licenses.
 While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}
@@
 
+All the metadata available through the API has been archived.
+In addition to the basics like title, album or artist names, this includes per-track genres and user data such as per-track/album/artist favorites, play counts and comments.
+% which are necessary to evaluate content-based recommender systems.
+There is also free-form text such as per-track/album/artist tags, album description and artist biography. 
 \todo{describe the metadata}
@@
 
-\tabref{tab:features} lists the features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
-These are all the features the library is able to extract.
+To allow researchers to experiment without dealing with feature extraction, we pre-computed the features listed in \tabref{tab:features}.
+These are all the features the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}, was able to extract.
 Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples.
@@
 	\item \textbf{Full}: the complete dataset, described above.
-All \ngenres genres, unbalanced with 1 to 38,154 tracks per genre (\figref{fig:genre_distribution}) and 0 to 31 genres per track (\figref{fig:genres_per_track}).
+All \ngenres genres, unbalanced with 1 to 38,154 tracks per genre (\figref{fig:genre_distribution}) and up to 31 genres per track (\figref{fig:genres_per_track}).
 	\item \textbf{Large}: the full dataset with audio limited to 30 seconds clips extracted from the middle of the tracks (or entire track if shorter than 30 seconds).
@@
 Genre unbalanced with 21 to 7,103 clips per top genre (\figref{fig:genre_top_distribution}), but only one of the 16 top genres per clip.
-		\todo{This subset was constructed for single root genre prediction. The clips were selected based on ... Non-musical genres, experimental?}
+		\todo{This subset was constructed for single root genre prediction.
+As such, we selected those tracks with only one top genre.
+Sampling.
+The clips were selected based on ...
+Non-musical genres, experimental?}
 	\item \textbf{Small}: 8,000 30s clips from 8 top genres, balanced with 1,000 clips per genre, 1 genre per clip.
@@
 While our intention was to release a large volume of audio for machine learning algorithms, analyzing audio is certainly of interest to musicologists and researchers who want to study relations with higher-level representations.
-Moreover, the metadata is surely a valuable addition to existing datasets\footnote{Like MusicBrainz, AllMusic, Discogs and Last.fm.} for metadata analysis.
+Moreover, the availability of complete tracks allows proper study of music properties, for example music structure analysis.
+Finally, the metadata is surely a valuable addition to existing datasets\footnote{Like MusicBrainz, AllMusic, Discogs and Last.fm.} for metadata analysis.
 %For instance, one could address questions like: How broad artists are, i.e.\ to which genres do they contribute to?
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
+\section{Discussion} % Limitations
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%\subsection{Limitations}
 % ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
@@
 Diversity is another issue: as suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental, electronic, and rock music.
-Moreover, it probably does not contain hits from renown artists or commercial music.
+Moreover, it probably does not contain mainstream music.
+%hits from renown artists or commercial music.
+It is unknown whether datasets made up of mainstream or non-mainstream music have the same properties and if algorithms tailored on one perform similarly on the other.
+Nevertheless, algorithms should ideally be evaluated on multiple datasets.
 \todo{more discussion after reading Sturm}

### 2017-07-13 fma.tex
remove comparison with computer vision datasets (metareview)

@@
 % * A Survey of Evaluation in Music Genre Recognition
-
-\begin{table*}[t]
-	\centering
-	\begin{threeparttable}
-		\begin{tabular}{l@{\hspace{2.7em}}r
-			@{\hspace{2.7em}}>{$}r<{$}
-			@{}>{${}={}$}c@{}
-			>{$}l<{$}
-			@{\hspace{2.7em}}>{$}c<{$}
-			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
-		\toprule
-		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
-		\cmidrule{6-8}
-				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & \#days \\
-		\midrule
-		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} &
-			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\
-		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
-			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & 776 & 241 \\
-		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
-			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
-		\href{\weblink}{FMA} \tnote{2} &
-			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
-		\midrule
-		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
-			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
-		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
-			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
-		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
-			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
-		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
-			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
-		\bottomrule
-	\end{tabular}
-	\begin{tablenotes}
-		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
-		\item[2] Average duration and sample rate. Most clips from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
-		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
-		\item[4] Object detection / classification task. It is the most used dataset for image classification.
-	\end{tablenotes}
-	\end{threeparttable}
-	\caption{Comparison between the largest public audio and image datasets.
-Dimensionality is (length [s] $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images.
-Volume is the number of samples times the dimensionality.
-Size is for a directory containing all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data.
-The last column is the number of days necessary to listen to the whole available audio.}
-	\label{tab:size}
-\end{table*}
 % Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
 % Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
-% Robin: Alors il télécharge aussi des clips audios de 10 secondes, non-compressé et à 48 kHz pour un total de 4 Terabytes! Donc environ 8x la taille de ton dataset il semble.
-% uncompressed: 2084320*10*48e3*2*2 / 1024**4 = 3.6
-% 320 kbit/s mp3: 2084320*10*320e3 / 8 / 1024**3 = 776
-
-% Xavier: De meme pour Table 2, c'est interessant bien sur, mais encore une fois, les datasets de MSD et autres n'ont pas disponible (directement et facilement) les raw audio data.
-Cela rend confusing le main message.
-Aussi, je ne parlerais pas des MNIST, CIFAR et autres image datasets, ce n'est pas le point de ce papier, targeted music, et cela peut rendre notre message confusing.
 
@@
 While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (\tabref{tab:datasets}).
-\todo{For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (\tabref{tab:size}). % and shall thus be sufficient for the foreseable future.
-Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.}
 
@@
 The FMA enables researchers to test algorithms on a large-scale collection, closer to real-world environments.
-Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of tracks,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset (\tabref{tab:size}) which opened the door to dramatic performance improvements for many tasks in computer vision.
+Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of tracks,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset which opened the door to dramatic performance improvements for many tasks in computer vision.
 %is also far behind the amount of tagged photos \href{https://www.facebook.com}{Facebook} has access to for example.

### 2017-07-14 fma.tex
update

@@
 It provides full-length and high-quality audio, pre-computed features, together with track- and user-level metadata, tags, and free-form text such as biographies.
-We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 major genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with 30s clips.
-This paper describes the dataset and how it was created, discuss some suitable MIR tasks, and evaluates some baselines for genre recognition.
+We here describe the dataset and how it was created, propose a train/validation/test split and three subsets, discuss some suitable MIR tasks, and evaluate some
+baselines for genre recognition.
 Code, data, and usage examples are available at \url{\weblink}.
@@
 The imposed constraint that input data cannot be made available to participants however incurs an approximately linear cost with the number of submissions as all algorithms have to be run by the organizers.
-By releasing this open dataset, we realize part of the vision of McFee et al.\ in ''a distributed, community-centric paradigm for system evaluation, built upon the principles of openness, transparency, and reproducibility`` \cite{mirex_critic}.
+By releasing this open dataset, we realize part of the vision of McFee et al.\ in ``a distributed, community-centric paradigm for system evaluation, built upon the principles of openness, transparency, and reproducibility'' \cite{mirex_critic}.
 
@@
 Each track is legally free to download as artists decided to release their works under permissive licenses.
-While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}
+While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis.
+While the FMA is not the only source of CC-licensed music, see e.g.\ \href{https://www.jamendo.com}{Jamendo}, it is unique as it combines user-generated content with the curatorial role that WFMU and others have always played.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}
 
@@
 
-\todo{The high coverage of genre information and fine-graine... makes the FMA well suited for MGR.}
-Labeling noise is unavoidable
+\todo{Labeling noise is unavoidable}
 The FMA is especially suited for MGR as it features fine genre information, i.e.\ multiple (sub-)genres associated to individual tracks, has a built-in genre hierarchy (\tabref{tab:genres}), and is annotated by the artists themselves.
@@
 
-In addition,
+%In addition,
 % a number of descriptive medium-level and high-level audio
-features extracted by \href{http://the.echonest.com/}{Echonest} are provided for 13,129 tracks.
+%features extracted by \href{http://the.echonest.com/}{Echonest} are provided for 13,129 tracks.
 These features include 8 high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), 5 social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and 224 MFCC-like temporal features for a number of segments \cite{echonest_temporal}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%The full potential of the FMA has yet to be unlashed.
 
@@
 
-Music classification is the most straightforward use case for FMA, with MGR being particularly suitable.
-The inclusion of a genre hierarchy makes it specially interesting, as it offers possibilities rarely found in alternative collections.
-
 While the FMA can be used to evaluate many tasks, metadata is missing for e.g.\ mood classification or instrument recognition.
@@
 Similarly, cover song detection may be doable if multiple versions of many songs are featured.
-Moreover, we can expect the dataset to be cross-referenced with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
+% Due to the potential commercial value of a working system, music recommendation and similarity are among the best studied areas of MIR.
+% So far, content-based systems have fallen short at predicting user ratings when compared to collaborative filtering methods. \todo{ref}
+While the present dump only captures listening and downloading counts in aggregates\footnote{That information can be useful to e.g.\ analyze and predict hits.}, the lists of which songs, albums and artists a user marked as favorites or commented are public, as well as \textit{user mixes}.
+While not public, listening and downloading activities are logged and might be shared after anonymization.\footnote{Private discussion with the administrators.} Moreover, users form a public social network via \textit{friend requests}.
+Collecting this information would opens the possibility of a large-scale evaluation of content-based recommender systems.
+%The wealth of available data opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features.
+Finally, we can expect the dataset to be cross-referenced with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
 
 Diversity is another issue: as suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental, electronic, and rock music.
-Moreover, it probably does not contain mainstream music.
-%hits from renown artists or commercial music.
-It is unknown whether datasets made up of mainstream or non-mainstream music have the same properties and if algorithms tailored on one perform similarly on the other.
-Nevertheless, algorithms should ideally be evaluated on multiple datasets.
-\todo{more discussion after reading Sturm}
-study to what extent is the FMA representative of mainstream music and its similarity to other datasets.
-\todo{McFee}
+Moreover, it does not contain mainstream music and few commercially successful artists.
+A common criticism of basing research on CC-licensed music is that the music is of substantially lower ``quality''.
+Moreover, it is unknown whether datasets made up of mainstream or non-mainstream music have similar properties and if algorithms tailored on one perform similarly on the other.
+While those points are valid for high-level tasks such as recommendation (which depend on a variety of factors beyond the acoustic content), this is a much more tenuous case for the majority of tasks, in particular perceptual tasks.
+% \cite{mirex_critic}
 % features some renown artists (wikipedia list)
-
-%The full potential of the FMA has yet to be unlashed.
-
-\subsection{Music Recommendation} % Recommendation, Recommender Systems
-% Cite Kirell's paper: Song recommendation with non-negative matrix factorization and graph total variation
-%Cheyenne: We do have lists for each user of which songs, albums and artists a user marked as 'favorites' though.
-%friends form a social networks
-% fans and favorites
-% user mixes
-
-Due to the potential commercial value of a working system, music recommendation and music similarity are among the best studied areas of MIR.
-%So far, content-based systems have fallen short at predicting user ratings when compared to collaborative filtering methods. \todo{ref}
-While the present dump only captures listening and downloading counts in aggregates (\figref{fig:listens_distribution})\footnote{That information can be useful to e.g.\ predict hits.}, the lists of which songs, albums and artists a user marked as favorites or commented are public.
-While not public, listening and downloading activities are logged and might be shared after anonymization.\footnote{Private discussion with the administrators.} Moreover, users form a public social network via \textit{friend requests}.
-%The wealth of available data opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features.
-%with a clean ground truth.
+Nevertheless, algorithms should ideally be evaluated on multiple datasets, which will help answer such questions.
 
@@
 
-Benchmarking is an important aspect in experimental sciences -- results reported by individual research groups need to be comparable.
+Benchmarking is an important aspect in experimental sciences --- results reported by individual research groups need to be comparable.
 Important aspects of these are datasets that can be easily shared among researchers, together with a set of defined tasks and splits.
@@
 % Kirell: less DL oriented, some people don't like it. Insist on available audio.
+For now, music classification, and MGR in particular, is the most straightforward use case for FMA.
+The inclusion of a genre hierarchy makes it specially interesting, as it offers possibilities rarely found in alternative collections.
 
@@
 % Future work
-\todo{If the community finds interest in the dataset and validate its use, future work should focus on validating the ground truth and collecting additional information.}
+\todo{If the community finds interest in the dataset and validate its use, future work should focus on validating the ground truth and collecting additional information by cross-referencing the dataset with other databases.}
 In addition to the proposed usage and many others people will find, future work on the dataset itself may consider to scrap the website for information not available through the API, e.g.\ the number of downloads or the user comments.
@@
 
-In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{post about the dataset}, Cheyenne Hohman, the Director at the Archive, wrote that ''by embracing the
+In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{post about the dataset}, Cheyenne Hohman, the Director at the Archive, wrote that ``by embracing the
 ~\ldots~%some rights reserved
-philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications``.
+philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications''.
 Let's hope for a future where sharing is first and researchers feed open platforms with algorithms while they feed us with data.

### 2017-07-14 fma.tex
update camera-ready

@@
 % to achieve their purpose, which are sometimes challenging to collect and share.
-In computer vision, the community has developed established benchmark datasets such as \href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist}, \href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR} \cite{cifar}, or \href{http://www.image-net.org}{ImageNet} \cite{imagenet}, which have proved essential to advance the field.
-The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.3M images \cite{imagenet_challenge}, demonstrated the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best \cite{convnet_imagenet}, and enabled incredible achievements in both fields \cite{dl}.
+In computer vision, the community has developed established benchmark datasets such as \href{http://yann.lecun.com/exdb/mnist/}{MNIST}~\cite{mnist}, \href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR}~\cite{cifar}, or \href{http://www.image-net.org}{ImageNet}~\cite{imagenet}, which have proved essential to advance the field.
+The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.3M images~\cite{imagenet_challenge}, demonstrated the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best~\cite{convnet_imagenet}, and enabled incredible achievements in both fields~\cite{dl}.
 
@@
 		\midrule
-		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC} \cite{RWC} & 465 & - & 2001 & yes \\
-		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500} \cite{cal500} & 500 & 500 & 2007 & yes \\
-		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
-		\href{https://marsyasweb.appspot.com/download/data_sets/}{GTZAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
-		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
-		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
+		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC}~\cite{RWC} & 465 & - & 2001 & yes \\
+		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500}~\cite{cal500} & 500 & 500 & 2007 & yes \\
+		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom}~\cite{ballroom} & 698 & - & 2004 & yes \\
+		\href{https://marsyasweb.appspot.com/download/data_sets/}{GTZAN}~\cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
+		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef}~\cite{musiclef} & 1,355 & 218 & 2012 & yes \\
+		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20}~\cite{artist20} & 1,413 & 20 & 2007 & yes \\
 		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 1,458 & - & 2004 & yes \\
-		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
-		103-Artists \cite{103artists} & 2,445 & 103 & 2005 & yes \\
-		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} \cite{unique} & 3,115 & 3,115 & 2010 & yes \\
-		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} \cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
-		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
-		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
-		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
-		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
-		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
-		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
+		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg}~\cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
+		103-Artists~\cite{103artists} & 2,445 & 103 & 2005 & yes \\
+		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique}~\cite{unique} & 3,115 & 3,115 & 2010 & yes \\
+		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists}~\cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
+		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD}~\cite{lmd} & 3,227 & - & 2007 & no \\
+		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom}~\cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
+		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP}~\cite{uspop} & 8,752 & 400 & 2003 & no \\
+		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k}~\cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
+		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune}~\cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
+		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich}~\cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
 		\bf \href{\weblink}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
-		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
-		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
-		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
-		\href{https://acousticbrainz.org}{AcousticBrainz} \cite{acousticbrainz} & 2,508,627\tnote{5} & - & 2017 & no \\
+		\href{http://www.omras2.org/}{OMRAS2}~\cite{omras} & 152,410 & 6,938 & 2009 & no \\
+		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD}~\cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
+		\href{https://research.google.com/audioset/}{AudioSet}~\cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
+		\href{https://acousticbrainz.org}{AcousticBrainz}~\cite{acousticbrainz} & 2,524,739\tnote{5} & - & 2017 & no \\
 		\bottomrule
@@
 		\item[4] Low quality 16 kHz, 32 kbit/s, mono mp3.
-		\item[5] As of 2017-07-10, of which a subset has been linked to genre labels for the \href{http://www.multimediaeval.org/mediaeval2017}{MediaEval 2017} \href{https://multimediaeval.github.io/2017-AcousticBrainz-Genre-Task}{genre task}.
+		\item[5] As of 2017-07-14, of which a subset has been linked to genre labels for the \href{http://www.multimediaeval.org/mediaeval2017}{MediaEval 2017} \href{https://multimediaeval.github.io/2017-AcousticBrainz-Genre-Task}{genre task}.
 	\end{tablenotes}
 	\end{threeparttable}
-	\caption{Comparison between FMA and alternative datasets which can be used for genre recognition.}
+	\caption{Comparison between FMA and alternative datasets.}
 	\label{tab:datasets}
@@
 % Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
+
+\begin{table}
+	\small
+	\centering
+	\begin{threeparttable}
+	\begin{tabular}{|r@{ }l@{\hspace{.8em}}r@{ }l@{\hspace{.8em}}r@{ }l|}
+		\hline
+		100\% & track\_id     & 100\% & title          &  93\% & number \\
+		  2\% & information   &  14\% & language\_code & 100\% & license \\
+		  4\% & composer      &   1\% & publisher      &   1\% & lyricist \\
+		 98\% & genres        &  98\% & genres\_all    &  47\% & genre\_top \\
+		100\% & duration      & 100\% & bit\_rate      & 100\% & interest \\ %samplerate \\
+		100\% & \#listens     &   2\% & \#comments     &  61\% & \#favorites \\
+		100\% & date\_created &   6\% & date\_recorded &  22\% & tags\\
+		\hline
+		100\% & album\_id     & 100\% & title          &     &  \\
+		 94\% & type          &  96\% & \#tracks       &     &  \\
+		 76\% & information   &  16\% & engineer       &  18\% & producer \\
+		 97\% & \#listens     &  12\% & \#comments     &  38\% & \#favorites \\
+		 97\% & date\_created &  64\% & date\_released &  18\% & tags \\
+		\hline
+		100\% & artist\_id    & 100\% & name           &  25\% & members \\
+		 38\% & bio           &   5\% & \multicolumn{2}{@{}l}{associated\_labels} &  \\
+		 43\% & website       &   2\% & wikipedia\_page & &  \\
+		& & 5\% & \multicolumn{2}{@{}l}{related\_projects} & \\
+		 37\% & location      &  23\% & longitude      &  23\% & latitude \\
+		 11\% & \#comments    &  48\% & \#favorites    &  10\% & tags\tnote{1} \\
+		 99\% & date\_created &   8\% & active\_year\_begin & & \\
+		& & 2\% & \multicolumn{2}{@{}l}{active\_year\_end} & \\
+		\hline
+	\end{tabular}
+	\begin{tablenotes}
+		\item[1] One of the tags is often the artist name. It has been subtracted.
+	\end{tablenotes}
+	\end{threeparttable}
+	\caption{List of available per-track, per-album and per-artist metadata, i.e.\ the columns of \texttt{tracks.csv}. Percentages indicate coverage over all tracks, albums, and artists.}
+	%Not all of them are filled for each track.\footnote{See }
+	%\texttt{albums.csv} and \texttt{artists.csv}.}
+	% genres.csv ?
+	\label{tab:metadata}
+\end{table}
+
+% {\bf Available music datasets and their limitations.}
+Unlike the wealth of available visual or textual content, the lack of a large, complete and easily available dataset for MIR has hindered research on data-heavy models such as DL.
+\tabref{tab:datasets} lists the most common datasets used for content-based MIR.
+GTZAN~\cite{gtzan}, a collection of 1000 clips from 10 genres, was the first publicly available benchmark dataset for genre recognition (MGR).
+As a result, despites its flaws (mislabeling, repetitions, and distortions), it continues to be the most used dataset for MGR~\cite{mgr_eval_2}.
+Moreover, it is small and misses metadata which e.g.\ prevents researchers to control for artists or album effects.
+% Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
+Looking at \tabref{tab:datasets}, the well-known MagnaTagATune~\cite{magnatagatune} and the Million Song Dataset (MSD)~\cite{msd} as well as the newer AudioSet~\cite{audioset} and AcousticBrainz~\cite{acousticbrainz} appear as contenders for a large-scale reference dataset.
+MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes metadata, features and audio.
+The poor audio quality and limited number of songs does however limit its usage.
+MSD and AudioSet, while very large, force researchers to download audio clips from online services.
+AcousticBrainz's approach to the copyright issue is to ask the community to upload music descriptors of their tracks.
+Although it is the largest database to date, it will never distribute audio.
+On the other hand, the proposed dataset offers the following qualities, which in our view are essential for a reference benchmark.
+
+% Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
+% MSD: Each song data is composed of high-level and medium-level audio features provided by Echonest service and metadata like artist name, song title, track genre, etc.
+It has its own dedicated website with related code and additional datasets.
+% AudioSet: 2.1 million annotated 10-second sound clips associated with 527 classes. The ontology is quite broad and covers sounds from music to cars, engines or animals.
+
+% Value proposition.
+%\textbf{\todo{Bias free}.} sampling bias (reviewer 1 comment 15)~\cite{mgr_eval_2}
+
+\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with musical categories.
+They are also more diverse and allows to average out annotation noise as well as characteristics who might be confounded with the ground truth and exploited by learning algorithms.
+While FMA features less clips than MSD or AudioSet, every other dataset with available quality audio are two orders of magnitude smaller (\tabref{tab:datasets}).
+
+\textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets, which stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
+The FMA's solution is to aim for tracks which license permits redistribution.
+All data and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{CC BY 4.0} and MIT licenses.
+
+\textbf{Available audio.}
+\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, most of the larger do not.
+They either (i) only contain features derived from the audio, or (ii) provide links to download the audio from an online service.\footnote{Going to the source distributor is a way to adhere with copyright.}
+The problem with (i) is that researchers are stuck with the chosen features and are prevented to leverage feature learning or end-to-end learning systems like DL.
+Moreover, we should be wary of proprietary features like those computed by commercial services such as \href{http://the.echonest.com/}{Echonest}.
+The problem with (ii) is that researchers have no control, i.e.\ we have no assurance that the files or services will not disappear or change without notice.
+
+\textbf{Quality audio.} % high-quality
+Distributed or downloadable audio are usually clips of 10 to 30 seconds and sometimes of low quality, e.g.\ 32 kbit/s for MagnaTagATune or an average of 104 kbit/s for MSD~\cite{msd_features}.
+The problem with clips is that the beginning 30 seconds of tracks may yield different results than the middle or final 30 seconds, and that researchers may not have control over which part they get.
+In comparison, FMA comes with full-length and high-quality audio.
+
+\textbf{Metadata rich.} The dataset comes with rich metadata, shown in \tabref{tab:metadata}.
+While not complete in any means, it compares favorably with the MSD which only provides artist-level metadata~\cite{msd} or GTZAN which offers none.
+
+\textbf{Easily accessible.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.csv} metadata and \texttt{.mp3} audio.
+No need to crawl the web and circumvent rate limits or access restrictions.
+Besides, we provide
+% todo an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and
+some usage examples in the \texttt{usage.ipynb} Jupyter notebook to start using the data quickly.
+
+\textbf{Future proof and reproducible.} All files and archives are checksummed and hosted in a long-term digital archive.
+%\todo{have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}}.
+Doing so alleviates the risks of songs to become unavailable.
+Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) generate the subsets and splits, (iv) compute the features and (v) test the baselines.
+The developed code can serve as a starting point for researchers to compute their own features or evaluate their methods.
+Finally, anybody can recreate or extend the collection, thanks to public songs and APIs.
+
+Note that an alternative to open benchmarking is the approach taken by the MIREX evaluation challenges: the evaluation (by the organizers) of submitted algorithms on private datasets~\cite{mirex}.
+This practice however incurs an approximately linear cost with the number of submissions, which put the long-term sustainability of MIREX at risk~\cite{mirex_critic}.
+By releasing this open dataset, we realize part of the vision of McFee \textit{et al}.\ in ``a distributed, community-centric paradigm for system evaluation, built upon the principles of openness, transparency, and reproducibility''.
 
@@
 
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{|r@{ }l@{\hspace{.8em}}r@{ }l@{\hspace{.8em}}r@{ }l|}
-		\hline
-		100\% & track\_id     & 100\% & title          &  93\% & number \\
-		  2\% & information   &  14\% & language\_code & 100\% & license \\
-		  4\% & composer      &   1\% & publisher      &   1\% & lyricist \\
-		 98\% & genres        &  98\% & genres\_all    &  47\% & genre\_top \\
-		100\% & duration      & 100\% & bit\_rate      & 100\% & interest \\ %samplerate \\
-		100\% & \#listens     &   2\% & \#comments     &  61\% & \#favorites \\
-		100\% & date\_created &   6\% & date\_recorded &  22\% & tags\\
-		\hline
-		100\% & album\_id     & 100\% & title          &     &  \\
-		 94\% & type          &  96\% & \#tracks       &     &  \\
-		 76\% & information   &  16\% & engineer       &  18\% & producer \\
-		 97\% & \#listens     &  12\% & \#comments     &  38\% & \#favorites \\
-		 97\% & date\_created &  64\% & date\_released &  18\% & tags \\
-		\hline
-		100\% & artist\_id    & 100\% & name           &  25\% & members \\
-		 38\% & bio           &   5\% & \multicolumn{2}{@{}l}{associated\_labels} &  \\
-		 43\% & website       &   2\% & wikipedia\_page & &  \\
-		& & 5\% & \multicolumn{2}{@{}l}{related\_projects} & \\
-		 37\% & location      &  23\% & longitude      &  23\% & latitude \\
-		 11\% & \#comments    &  48\% & \#favorites    &  10\% & tags \\
-		 99\% & date\_created &   8\% & active\_year\_begin & & \\
-		& & 2\% & \multicolumn{2}{@{}l}{active\_year\_end} & \\
-		\hline
-	\end{tabular}
-	\caption{List of available per-track, per-album and per-artist metadata, i.e.\ the columns of \texttt{tracks.csv}. Percentages indicate coverage over all tracks, albums, and artists.}
-	%Not all of them are filled for each track.\footnote{See }
-	%\texttt{albums.csv} and \texttt{artists.csv}.}
-	% genres.csv ?
-	\label{tab:metadata}
-\end{table}
-
-An alternative to open benchmarking is the approach taken by evaluation challenges such as the music information retrieval evaluation exchange (MIREX) \cite{mirex}.
-The imposed constraint that input data cannot be made available to participants however incurs an approximately linear cost with the number of submissions as all algorithms have to be run by the organizers.
-By releasing this open dataset, we realize part of the vision of McFee et al.\ in ``a distributed, community-centric paradigm for system evaluation, built upon the principles of openness, transparency, and reproducibility'' \cite{mirex_critic}.
-
-% {\bf Available music datasets and their limitations.}
-Unlike the wealth of available visual or textual content, the lack of a large-scale, complete and easily available dataset for MIR has hindered research on data-heavy models such as DL.
-\tabref{tab:datasets} lists the most common datasets used for content-based MIR.
-GTZAN \cite{gtzan}, a collection of 1000 clips from 10 genres, was the first publicly available benchmark MGR dataset.
-As a result, despites its flaws, it continues to be the most used dataset for MGR \cite{mgr_eval_2}.
-The main problems are mislabeling, repetitions, and distortions.
-Moreover, it is small and misses metadata which prevents researchers to control for artists or album effects.
-% Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
-Looking at \tabref{tab:datasets}, the well-known MagnaTagATune \cite{magnatagatune} and the Million Song Dataset (MSD) \cite{msd} as well as the newer AudioSet \cite{audioset} and AcousticBrainz \cite{acousticbrainz} appear as contenders for a large-scale reference dataset.
-% They however each have limitations.
-MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes metadata, features and audio.
-The poor audio quality and limited number of songs does however limit its usage.
-MSD and AudioSet, while very large, force researchers to download audio clips from online services.
-AcousticBrainz's approach to the copyright issue is to ask the community to upload music descriptors of their tracks.
-Although it is the largest database to date, it will never distribute audio.
-% Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
-% Each song data is composed of high-level and medium-level audio features provided by Echonest service and metadata like artist name, song title, track genre, etc.
-It has its own dedicated website with related code and additional datasets.
-% 2.1 million annotated 10-second sound clips associated with 527 classes. The ontology is quite broad and covers sounds from music to cars, engines or animals. 
-On the other hand, the proposed dataset offers the following qualities, which in our view are essential for a reference benchmark.
-
-% Value proposition.
-\textbf{Avoiding biases.} sampling bias TODO: reviewer 1 comment 15 \cite{mgr_eval_2} Also McFee 2016
-
-\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with musical categories.
-They are also more diverse and allows to average out annotation noise as well as characteristics who might be confounded with the ground truth and exploited by learning algorithms.
-While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (\tabref{tab:datasets}).
-
-\textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets, which stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
-The FMA's solution is to aim for tracks which license permits redistribution.
-All metadata and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{
-%Creative Commons Attribution 4.0 International License (CC BY 4.0)
-CC BY 4.0} and MIT licenses.
-% TODO paper CC BY 4.0 if published at ISMIR
-% TODO licensing
-% Creative Commons: MagnaTagATune
-% Proprietary, non-redistributable: MSD, AudioSet?
-
-\textbf{High-quality audio.}
-\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, most of the larger do not.
-% often regardless of copyrights
-They either only contain features derived from the audio, or provide links to download the audio from an online service.
-The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later makes no assurance that the files or services will not disappear or change without notice.
-Moreover, researchers should be wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest}.
-% They have e.g.\ extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
-Finally, distributed or downloadable audio are usually clips of 10 to 30 seconds, sometimes of low quality, e.g.\ 32 kbit/s for MagnaTagATune or an average of 104 kbit/s for MSD \cite{msd_features}.
-The problem with clips is that the beginning 30 seconds of tracks may yield different results than the middle or the final 30 seconds, and that researchers may not have control over which part they get.
-In comparison, FMA comes with CC-licensed, full-length, high-quality and non-modified audio.
-% \todo{See Table ? for a quantitative analysis of quality.}
-
-\textbf{Metadata rich.} In addition to audio, the dataset comes with rich metadata, shown in \tabref{tab:metadata}.
-While not complete, it compares favorably with the MSD which only provides artist-level metadata \cite{msd} or GTZAN which offers none.
-% social such as artist location or website
-% images for tracks, albums and artists
-
-\textbf{Easily accessible.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.csv} metadata and \texttt{.mp3} audio.
-No need to crawl the web and circumvent rate limits or access restrictions.
-Besides, we provide
-% Todo an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and
-some usage examples in the \texttt{usage.ipynb} Jupyter notebook to start using the data quickly.
-
-\textbf{Future proof and reproducible.} All files and archives are checksummed and hosted in a long-term digital archive.
-%\todo{have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}}.
-Doing so alleviates the risks of songs to become unavailable, to change without notice or for online services to shut down.
-Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) generate the subsets and splits, (iv) compute the features and (v) test the baselines.
-The developed code can serve as a starting point for researchers to compute their own features or evaluate their methods.
-Finally, anybody can recreate or extend the collection, thanks to public songs and APIs.
-
-% Western / commercial music, how to qualify broadness?
-
-% Do not present like this because we might want representativeness. Which would necessarily include commercial music.
-%{\bf What would be the requirements for an ideal music dataset?} \\
-%(1) Availability of raw audio tracks like 30sec or more,\\
-%(2) Availability of metadata such as genres, artist, title, year, lyrics, users' ratings, users' comments, etc,\\
-%(3) Large-scale dataset to be able to sample the high-dimensional distribution of music genres\\
-%(4) Free and easy access of the dataset,\\
-%(5) Legality of the dataset, no copyright issue with e.g.\ open access licence.\\
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{The Free Music Archive}
+
+The dataset, both the audio and metadata, is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, a free and open library directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
+%The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
+%Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
+Inspired by \href{https://creativecommons.org/}{Creative Commons} and the open-source software movement, the FMA provides a platform for curators, artists, and listeners to harness the potential of music sharing.
+The website provides a large catalog of artists and tracks, hand-picked by established audio curators.
+Each track is legally free to download as artists decided to release their works under permissive licenses.
+While there exists other sources of CC-licensed music, notably \href{https://www.jamendo.com}{Jamendo}, FMA is unique as it combines user-generated content with the curatorial role that WFMU and others have always played.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}
+% An assurance of quality.
+
+\subsection{Creation} % Collection, Creation, Generation
+
+As of April 1st 2017, when the dataset was gathered, the online archive largest track id was 155,320, of which 109,727 were valid.
+The missing 45,594 ids probably correspond to deleted tracks.
+\figref{fig:growth} illustrates the growth of the dataset.
+In addition to per-track metadata, the used hierarchy of 161 genres and extended per-album (480 not found) and per-artist (250 not found) metadata were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.} Finally, mp3 audio was downloaded over HTTPS.
+Out of all collected track ids, 180 mp3s could not be downloaded, 286 could not be trimmed by ffmpeg, and features could not be extracted from 71.
+Finally, the license of 2,616 tracks prohibited their redistribution, leaving us with \ntracks tracks.
+
+%\subsection{Data Cleaning}
+While it may be argued that the dataset should be cleaned, we wanted it to resemble real world data.
+As such, we did not remove tracks which have too many genres, are too long, belong to rare genres, etc.
+Moreover, it is hard to set a threshold, algorithms shall handle outliers, and the small number of outliers will not impact performance much anyway.
+Researchers are obviously free to discard any track for training.
+
+% classes with less than 1% of the number of tracks of the biggest class Pop/Rock are removed
+%At this stage the number of genres is 138, but half have only a few samples.
+Thus we decided to only retain the genres with more than 100 samples.
+This defines a dataset of 77,643 music data with 68 genres.
+We further cleaned the dataset by removing tracks which belong to ``non-standard'' music genres like 'Noise', 'Garage', 'Sound Collage', 'Singer', 'Audio Collage', 'Glitch', 'Unclassifiable', 'Lo-Fi', 'Spoken', 'Poetry', 'Talk Radio', 'Avant-Garde', 'Experimental', 'Ambient', and 'Field Recording'.
+We also suppressed the songs with the title 'Untitled'.
+Besides, we decided to cut the long tail of the power law distribution by removing songs of genres with less than 300 songs.
 
@@
 
-\subsection{The Free Music Archive}
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{tag_distribution.pdf}
+	\caption{Per-track, album and artist tags (min 0, max 150).}
+	\label{fig:tag_distribution}
+\end{figure}
 
-The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
-%The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
-%Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
-Inspired by \href{https://creativecommons.org/}{Creative Commons} and the open-source software movement, the FMA provides a platform for curators, artists, and listeners to harness the potential of music sharing.
-The website provides a large catalog of artists and tracks, hand-picked by established audio curators.
-Each track is legally free to download as artists decided to release their works under permissive licenses.
-While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis.
-While the FMA is not the only source of CC-licensed music, see e.g.\ \href{https://www.jamendo.com}{Jamendo}, it is unique as it combines user-generated content with the curatorial role that WFMU and others have always played.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}
+\subsection{Content}
+% Figures and tables from analysis.ipynb
+% TODO audio properties: samplerate (% 44kHz), bitrate (% 128, VBR), channels (% mono, stereo)
 
-\subsection{Creation} % Collection, Creation, Generation
-% How it was collected.
+The collected metadata\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} was cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
+That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}.
+For ease of use, we kept all the metadata in a single table despite the redundancy incurred by the fact that all tracks from a given artist share all artist related columns.
+The problem is mitigated in practice by compression for storage and by \textit{categorical variables} for memory usage.
 
-% Collection
-As of April 1st 2017, when the dataset was gathered, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks).
-\figref{fig:growth} shows the growth of the dataset.
-In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.} Finally, mp3 audio was downloaded over HTTPS.
-% Cleaning
-Out of all collected track ids, 180 mp3s could not be downloaded, 286 could not be trimmed by ffmpeg, and features could not be extracted from 71.
-Finally, the license of 2,616 tracks prohibited their redistribution, leaving us with \ntracks tracks.
+All the metadata available through the API has been archived.
+It includes song title, album, artist, and per-track genres; user data such as per-track/album/artist favorites, play counts, and comments; free-form text such as per-track/album/artist tags, album description and artist biography.
+Coverage varies across fields and is reported in \tabref{tab:metadata}.
+Note that all that metadata has been produced by the artists when uploading their music and that while the content is curated, the curators focus on the musical content not the metadata.
+Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
+See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.
 
-While it may be argued that the dataset should be cleaned, we wanted it to resemble real world data.
-As such, we did not remove tracks which have too many genres, are too long, belong to rare genres, etc.
-Moreover, it is hard to set a threshold and the small number of outliers will not impact much the prediction accuracy.
-Researchers are obviously free to discard any track for training.
-
-%\subsection{Data Cleaning}
-% classes with less than 1% of the number of tracks of the biggest class Pop/Rock are removed
-
-%At this stage the number of genres is 138, but half have only a few samples.
-Thus we decided to only retain the genres with more than 100 samples.
-This defines a dataset of 77,643 music data with 68 genres.
-We further cleaned the dataset by removing tracks which belong to ``non-standard'' music genres like 'Noise', 'Garage', 'Sound Collage', 'Singer', 'Audio Collage', 'Glitch', 'Unclassifiable', 'Lo-Fi', 'Spoken', 'Poetry', 'Talk Radio', 'Avant-Garde', 'Experimental', 'Ambient', and 'Field Recording'.
-We also suppressed the songs with the title 'Untitled'.
-Besides, we decided to cut the long tail of the power law distribution by removing songs of genres with less than 300 songs.
+The audio for each track is stored in a file which name is the track id.
+All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kbit/s (263 kbit/s on average), and in stereo.
+%todo: graph for audio quality
+%\todo{See Table ? for a quantitative analysis of quality.}
 
@@
 	\caption{(left) Example of genre hierarchy for the top-level Soul-RnB genre.
-Left number is the \texttt{genre\_id}, right number is the number of tracks per genre.
-(right) Number of genres per track.}
+Left number is the \texttt{genre\_id}, right is the number of tracks per genre.
+(right) Number of genres per track.
+A 3 genres limit has been introduced early on by the administrators.}
 	\label{fig:genre_hierarchy}
@@
 
-\subsection{Content}
-% Figures and tables from analysis.ipynb
-% TODO audio properties: samplerate (% 44kHz), bitrate (% 128, VBR), channels (% mono, stereo)
-
-The metadata collected from the archive\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} was cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt.
-That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}.
-All the metadata has been produced by the artists when uploading their music.
-While the content is curated, the curators focus on the musical content not the metadata.
-Metadata includes song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags.
-Coverage varies across fields, see \tabref{tab:metadata}.
-For ease of use, we kept all the metadata in a single table despite the redundancy incurred by e.g.\ the fact that all tracks from a given artist share all artist related columns.
-From a practical point of view, the problem is mitigated by compression for storage and those columns can be treated as \textit{categorical variables} to minimize memory usage.
-
-All the metadata available through the API has been archived.
-In addition to the basics like title, album or artist names, this includes per-track genres and user data such as per-track/album/artist favorites, play counts and comments.
-% which are necessary to evaluate content-based recommender systems.
-There is also free-form text such as per-track/album/artist tags, album description and artist biography. 
-\todo{describe the metadata}
-In the FMA, tags can be found at the track, album and artist level: 23,496 tracks (22\%), 2,659 albums (18\%) and 1,605 artists (10\%) have at least one tag.
-\figref{fig:tag_distribution} shows their distribution.\footnote{One of the artist tag is often the artist name.
-We thus subtracted one.}
-
-The audio for each track is stored in a file which name is the track id.
-All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kbit/s (263 kbit/s on average), and in stereo.
-Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
-See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.
-%todo: graph for audio quality
-
 \subsection{Genres}
 
-\todo{Labeling noise is unavoidable}
 The FMA is especially suited for MGR as it features fine genre information, i.e.\ multiple (sub-)genres associated to individual tracks, has a built-in genre hierarchy (\tabref{tab:genres}), and is annotated by the artists themselves.
 While the artists are the best placed to judge the positioning of their creations, they might be inconsistent and motivated by factors not necessarily objective, such as achieving a higher play count.
-Ideally those labels should be one of many ground truths, to be complemented by crowd-sourcing and experts (from different music metadata websites).
+As labeling noise is unavoidable, those labels should ideally be one of many ground truths, to be complemented by crowd-sourcing and experts (from different music metadata websites).
 
-While there is no agreement on a taxonomy of genres \cite{mir_review_genre}, we followed the hierarchy used by the archive, which is the one the authors had in mind when annotating their tracks.
+While there is no agreement on a taxonomy of genres~\cite{mir_review_genre}, we followed the hierarchy used by the archive, which is the one the authors had in mind when annotating their tracks.
 That hierarchy is composed of \ngenres genres of which 16 are roots, the others being sub-genres.
@@
 \figref{fig:genre_hierarchy} shows an excerpt of the tree.
-\todo{3 genres limit}
 
@@
 Then, given such hierarchical information, we constructed a \texttt{genres\_all} column which contains all the genres encountered when traversing the tree from the indicated genres to the roots.
-Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
+%if the traversal yields only one root
+The root genres are stored in the \texttt{genres\_top} column.
 \figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre.
 %As often encountered in natural categories, it obeys a power law.
-
-While root genre recognition should be treated as a multi-label problem in general, \figref{fig:genres_per_track} shows that half the tracks have a single root genre.
 
@@
 	\includegraphics[width=\linewidth]{genre_top_distribution.pdf}
-	\caption{(top) Tracks per genre on the full set (min 1, max 38,154).
-(bottom) Tracks per all top genre on the medium subset (min 21, max 7,103).
+	\caption{(top) Tracks per (sub-)genre on the full set (min 1, max 38,154).
+(bottom) Tracks per all 16 root genres on the medium subset (min 21, max 7,103).
 Note how experimental music is much less represented in the curated medium subset.}
@@
 To allow researchers to experiment without dealing with feature extraction, we pre-computed the features listed in \tabref{tab:features}.
-These are all the features the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}, was able to extract.
+These are all the features the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0~\cite{librosa}, was able to extract.
 Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples.
@@
 %features extracted by \href{http://the.echonest.com/}{Echonest} are provided for 13,129 tracks.
-These features include 8 high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), 5 social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and 224 MFCC-like temporal features for a number of segments \cite{echonest_temporal}.
+These features include 8 high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), 5 social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and 224 MFCC-like temporal features for a number of segments~\cite{echonest_temporal}.
 They are distributed in \texttt{echonest.csv}.
@@
 That trimming reduces the size of the data by a factor 10.
-	\item \textbf{Medium}: 25,000 of the most popular 30s clips with good metadata and quality audio.
-Genre unbalanced with 21 to 7,103 clips per top genre (\figref{fig:genre_top_distribution}), but only one of the 16 top genres per clip.
-		\todo{This subset was constructed for single root genre prediction.
-As such, we selected those tracks with only one top genre.
-Sampling.
-The clips were selected based on ...
-Non-musical genres, experimental?}
-	\item \textbf{Small}: 8,000 30s clips from 8 top genres, balanced with 1,000 clips per genre, 1 genre per clip.
-This subset is similar to the very popular GTZAN \cite{gtzan} with the added benefits of the FMA, that is metadata, pre-computed features, and copyright-free audio.
-\todo{The same process was used to select the top 1000 tracks per genre.
-The selected genres are those with more than 1000 tracks in the medium subset.}
+	\item \textbf{Medium}: while root genre recognition should be treated as a multi-label problem in general, we constructed this subset for the simpler problem of single-label prediction.
+It makes sense as half the tracks have a single root genre (\figref{fig:genres_per_track}).
+As such, we selected those tracks with only one top genre and sampled the clips according to the completeness of their metadata and their popularity, hoping to select tracks of higher quality.
+That selection left us with 25,000 30s clips, genre unbalanced with 21 to 7,103 clips per top genre (\figref{fig:genre_top_distribution}), but only one of the 16 top genres per clip.
+		%\todo{Non-musical genres, experimental?}
+		\todo{}
+	\item \textbf{Small}: to construct a balanced subset, we selected with the same process the top 1,000 clips from the 8 most popular genres of the medium set.
+The subset is thus composed of 8,000 30s clips from 8 top genres, balanced with 1,000 clips per genre, 1 root genre per clip.
+This subset is similar to the very popular GTZAN~\cite{gtzan} with the added benefits of the FMA, that is metadata, pre-computed features, and copyright-free audio.
 \end{enumerate}
@@
 	\item Stratified sampling to preserve the percentage of tracks per genre (important for minority genres).
+Each root genre is guaranteed to be represented in all splits, but the ratio is only exact for the small subset (800/100/100).
+The seven smallest sub-genres (less than 20 tracks in total) are however not guaranteed to appear in all splits of the full and large sets.
 	\item An artist filter for artists to be part of one set only, thus avoiding any artist and album effect.
-It has been shown that the use of songs from the same artist in both training and test sets leads to over-optimistic accuracy and may favor some approaches \cite{artist_filters_1, artist_filters_2}.
+It has been shown that the use of songs from the same artist in both training and test sets leads to over-optimistic accuracy and may favor some approaches~\cite{artist_filters_1, artist_filters_2}.
 %All songs from a given artist either appear in train, validation or test, never in two of them.
@@
 This does not break the above constraint.}
-	\item The 2231 tracks without genre label are assigned to the training set as they might be useful as additional training samples for semi-supervised algorithms.\footnote{These are only found in the full and large sets.}
 \end{enumerate}
-The above constraints are satisfied for all subsets, and a track is assigned to the same split across all of them.\footref{creation} Each root genre is guaranteed to be represented in all splits, but the ratio is only exact for the small subset (800/100/100).
-The seven smallest sub-genres (less than 20 tracks) are however not guaranteed to appear in all splits of the full and large sets.
-
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{lrrccc}
-		\toprule
-		dataset & clips & genres & length & \multicolumn{2}{c}{size} \\
-		\cmidrule{5-6}
-		        &       &        &  [s]   & [GiB] & \#days \\
-		\midrule
-		small  &    8,000 &   8 &  30 &  7.4 & 2.8  \\
-		medium &   25,000 &  16 &  30 & 23   & 8.7  \\
-		large  & \ntracks & 161 &  30 & 98   & 37 \\
-		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
-		\bottomrule
-	\end{tabular}
-	\caption{Proposed subsets of the FMA.}
-	\label{tab:subsets}
-\end{table}
+The above constraints are satisfied for all subsets, and a track is assigned to the same split across all of them.\footref{creation} The 2,231 tracks without genre label are assigned to the training set (full and large sets) as they might be useful as additional training samples for semi-supervised algorithms.
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{tag_distribution.pdf}
-	\caption{Per-track, album and artist tags (min 0, max 150).}
-	\label{fig:tag_distribution}
-\end{figure}
-
 Music classification is a key problem in MIR with many potential applications.
 For one, a classification system enables end users to search for the types of music they are interested in.
-On the other hand, different music types are managed more effectively and efficiently once they are categorized into different groups \cite{mir_review_classif}.
+On the other hand, different music types are managed more effectively and efficiently once they are categorized into different groups~\cite{mir_review_classif}.
 The classification tasks which can readily be evaluated on FMA include genre recognition, artist identification, year prediction, and automatic tagging.
-
-Automatic tagging \cite{autotagging} is a classification problem which covers different semantic categories, where tags are labels which can be any musical term that describes the genre, mood, instrumentation, and style of the song.
+Automatic tagging~\cite{autotagging} is a classification problem which covers different semantic categories, where tags are labels which can be any musical term that describes the genre, mood, instrumentation, and style of the song.
 It helps to convert the music retrieval problem to text retrieval by substituting songs with tags.
 %The task of music annotation has recently gained much popularity in the MIR community.
-
 In addition to supervised methods which classify music given an arbitrary taxonomy, another approach is to cluster data in an unsupervised way so that a categorization will emerge from the data itself based on objective similarity measures.
@@
 
-\subsection{Genre Recognition} % MGR, Baselines
-
-Music genres are categories that have arisen through a complex interplay of cultures, artists, and market forces to characterize similarities between compositions and organize music collections.
-Yet, the boundaries between genres still remain fuzzy, making the problem of music genre recognition (MGR) a nontrivial task \cite{mir_review_genre}.
-% genre not defined by audio content only
-While its utility has been debated, mostly because of its ambiguity and cultural definition, it is widely used and understood by end-users who find it useful to discuss musical categories \cite{mgr_why}.
-As such, it is one of the most researched areas of MIR.
-We propose the following problems of increasing difficulty:
-\begin{enumerate}
-	\item Single top genre prediction on the small subset.
-	\item Single top genre prediction on the medium subset.
-	\item Multiple top genres prediction on the large / full set.
-	\item Multiple genres prediction on the large / full set.
-\end{enumerate}
-
-\tabref{tab:mgr} reports accuracies for problem 2 with nine mainstream feature sets and some combinations as well as four standard classifiers using scikit-learn, version 0.18.1 \cite{scikit-learn}.
-Specifically, we employed linear regression (LR) with an $L^2$ penalty, k-nearest neighbors (kNN) with $k=200$, support vector machines (SVM) with a radial basis function (RBF) kernel and a multilayer perceptron (MLP) with 100 hidden neurons.
-All classifiers were tested with otherwise default settings.\footnote{See \texttt{baselines.ipynb} for all details.} Reported performance should not be taken as the state-of-the-art but rather as a lower-bound and an indication of the task's difficulty.
-Moreover, the developed code can serve as a reference and is easily modified to accommodate other features and classifiers.
-
-A major motivation to construct this dataset was to enable the use of the powerful DL set of techniques to music analysis, an hypothesized cause of stagnation on MIREX tasks \cite{mirex_stagnation}.
-With availability of audio, DL architectures such as convolutional neural networks and recurrent neural networks can be applied to the waveform to avoid any feature engineering.
-While those approaches have fallen behind learning from higher-level representations such as spectrograms \cite{dieleman_endtoend}, a greater exploration of the design space will hopefully provide alternatives to solving MIR challenges \cite{mir_dl_feature_learning}.
-
-%Some very recent advances are indeed promising Chuo
+\begin{table}
+	\small
+	\centering
+	\begin{tabular}{lrrccc}
+		\toprule
+		dataset & clips & genres & length & \multicolumn{2}{c}{size} \\
+		\cmidrule{5-6}
+		        &       &        &  [s]   & [GiB] & \#days \\
+		\midrule
+		small  &    8,000 &   8 &  30 &  7.4 & 2.8  \\
+		medium &   25,000 &  16 &  30 & 23   & 8.7  \\
+		large  & \ntracks & 161 &  30 & 98   & 37 \\
+		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
+		\bottomrule
+	\end{tabular}
+	\caption{Proposed subsets of the FMA.}
+	\label{tab:subsets}
+\end{table}
 
@@
 		\midrule
-		1 Chroma \cite{chroma} & 84 & 44 & 44 & 48 & 49 \\
-		2 Tonnetz \cite{tonnetz} & 42 & 40 & 37 & 42 & 41 \\
-		3 MFCC \cite{mfcc} & 140 & 58 & 55 & 61 & 53 \\
+		1 Chroma~\cite{chroma} & 84 & 44 & 44 & 48 & 49 \\
+		2 Tonnetz~\cite{tonnetz} & 42 & 40 & 37 & 42 & 41 \\
+		3 MFCC~\cite{mfcc} & 140 & 58 & 55 & 61 & 53 \\
 		4 Spec. centroid & 7 & 42 & 45 & 46 & 48 \\
 		5 Spec. bandwidth & 7 & 41 & 45 & 44 & 45\\
-		6 Spec. contrast \cite{contrast} & 49 & 51 & 50 & 54 & 53 \\
+		6 Spec. contrast~\cite{contrast} & 49 & 51 & 50 & 54 & 53 \\
 		7 Spec. rolloff & 7 & 42 & 46 & 48 & 48 \\
@@
 		%Ech. social & 5 & - & - & - & - \\
-		%Ech. temporal \cite{echonest_temporal} & 224 & - & - & - & - \\
+		%Ech. temporal~\cite{echonest_temporal} & 224 & - & - & - & - \\
 		\bottomrule
@@
 
+\subsection{Genre Recognition} % MGR, Baselines
+
+Music genres are categories that have arisen through a complex interplay of cultures, artists, and market forces to characterize similarities between compositions and organize music collections.
+Yet, the boundaries between genres still remain fuzzy, making the problem of music genre recognition (MGR) a nontrivial task~\cite{mir_review_genre}.
+% genre not defined by audio content only
+While its utility has been debated, mostly because of its ambiguity and cultural definition, it is widely used and understood by end-users who find it useful to discuss musical categories~\cite{mgr_why}.
+As such, it is one of the most researched areas of MIR.
+We propose the following prediction problems of increasing difficulty:
+\begin{enumerate}
+	\item Single top genre on the balanced small subset.
+	\item Single top genre on the unbalanced medium subset.
+	\item Multiple top genres on the large / full set.
+	\item Multiple (sub-)genres on the large / full set.
+\end{enumerate}
+
+\tabref{tab:mgr} reports accuracies for problem 2 with nine mainstream feature sets and some combinations as well as four standard classifiers using scikit-learn, version 0.18.1~\cite{scikit-learn}.
+Specifically, we employed linear regression (LR) with an $L^2$ penalty, k-nearest neighbors (kNN) with $k=200$, support vector machines (SVM) with a radial basis function (RBF) kernel and a multilayer perceptron (MLP) with 100 hidden neurons.
+All classifiers were tested with otherwise default settings.\footnote{See \texttt{baselines.ipynb} for all details.} Reported performance should not be taken as the state-of-the-art but rather as a lower-bound and an indication of the task's difficulty.
+Moreover, the developed code can serve as a reference and is easily modified to accommodate other features and classifiers.
+
+A major motivation to construct this dataset was to enable the use of the powerful DL set of techniques to music analysis, an hypothesized cause of stagnation on MIREX tasks~\cite{mirex_stagnation}.
+With availability of audio, DL architectures such as convolutional neural networks and recurrent neural networks can be applied to the waveform to avoid any feature engineering.
+While those approaches have fallen behind learning from higher-level representations such as spectrograms~\cite{dieleman_endtoend}, a greater exploration of the design space will hopefully provide alternatives to solving MIR challenges~\cite{mir_dl_feature_learning}.
+
+%Some very recent advances are indeed promising Chuo
+
 \subsection{Data Analysis}
@@
 Moreover, the availability of complete tracks allows proper study of music properties, for example music structure analysis.
-Finally, the metadata is surely a valuable addition to existing datasets\footnote{Like MusicBrainz, AllMusic, Discogs and Last.fm.} for metadata analysis.
+Finally, the metadata is surely a valuable addition to existing datasets (e.g. MusicBrainz, AllMusic, Discogs, Last.fm) for metadata analysis.
 %For instance, one could address questions like: How broad artists are, i.e.\ to which genres do they contribute to?
@@
 % So far, content-based systems have fallen short at predicting user ratings when compared to collaborative filtering methods. \todo{ref}
-While the present dump only captures listening and downloading counts in aggregates\footnote{That information can be useful to e.g.\ analyze and predict hits.}, the lists of which songs, albums and artists a user marked as favorites or commented are public, as well as \textit{user mixes}.
-While not public, listening and downloading activities are logged and might be shared after anonymization.\footnote{Private discussion with the administrators.} Moreover, users form a public social network via \textit{friend requests}.
-Collecting this information would opens the possibility of a large-scale evaluation of content-based recommender systems.
+While the present dump only captures listening and downloading counts in aggregates,\footnote{That information can be useful to e.g.\ analyze and predict hits.} the lists of which songs, albums and artists a user marked as favorites or commented are public, as well as \textit{user mixes}.
+While not public, listening and downloading activities are logged and might be shared after anonymization.\footnote{Private discussion with the website administrators.} Moreover, users form a public social network via \textit{friend requests}.
+Collecting this information would open the possibility of a large-scale evaluation of content-based recommender systems.
 %The wealth of available data opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features.
-Finally, we can expect the dataset to be cross-referenced with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
+Cover images for tracks, albums, and artists are another public asset which may be of interest.
+Finally, we can expect the dataset to be cross-referenced with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition~\cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.
 
-Diversity is another issue: as suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental, electronic, and rock music.
+Diversity is another issue.
+As suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental, electronic, and rock music.
 Moreover, it does not contain mainstream music and few commercially successful artists.
@@
 While those points are valid for high-level tasks such as recommendation (which depend on a variety of factors beyond the acoustic content), this is a much more tenuous case for the majority of tasks, in particular perceptual tasks.
-% \cite{mirex_critic}
+%~\cite{mirex_critic}
 % features some renown artists (wikipedia list)
@@
 % Future work
-\todo{If the community finds interest in the dataset and validate its use, future work should focus on validating the ground truth and collecting additional information by cross-referencing the dataset with other databases.}
-In addition to the proposed usage and many others people will find, future work on the dataset itself may consider to scrap the website for information not available through the API, e.g.\ the number of downloads or the user comments.
-Moreover, cross-referencing with other resources or crowd-sourcing (with e.g.\ \href{https://www.mturk.com}{Mechanical Turk} or \href{https://www.crowdflower.com/}{CrowdFlower}) is desirable to obtain additional labels or measure agreement on the current ones by independent annotators.
+In addition to the proposed usage and many others people will find, future work on the dataset itself should focus on (i) validating the ground truth by measuring agreement by independent annotators and (ii) obtaining additional metadata and labels.
+If the community finds interest in the dataset and validate its use, that can be achieved by scraping the website for information not available through the API,
+%e.g.\ the number of downloads or the user comments.
+cross-referencing with other resources, or crowd-sourcing (with e.g.\ \href{https://www.mturk.com}{Mechanical Turk} or \href{https://www.crowdflower.com/}{CrowdFlower}).
 
